---
title: "EM-Tenure Model: Progress Update & Discussion Points"
author: "Zander Prinsloo"
date: "2026-04-07"
format:
  html:
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: true
    theme: cosmo
---

## Purpose of This Document

Prepared for supervisor meeting on **7 April 2026**. Summarises:

1. Where the project stands
2. What works and what doesn't
3. The core technical problem (with toy calculations)
4. Proposed path forward and where I need input

---

## Project Status

### Data

We use the South African Quarterly Labour Force Survey (QLFS), a rotating panel
where individuals are observed at 3 quarterly waves.

| Statistic | Value |
|:---|---:|
| Raw observations (post-ingest) | 402,923 |
| After excluding never-worked | ~261,000 |
| Employment rate (wave 1) | 47.0% |
| Employment rate (wave 2) | 47.0% |
| Employment rate (wave 3) | 47.2% |

### Observed (naïve) transition rates

These ignore misclassification — they are the raw quarter-to-quarter flows:

| From → To | Count | % of total |
|:---|---:|---:|
| Nonemployed → Nonemployed | 195,525 | 48.5% |
| Nonemployed → Employed | 17,910 | 4.5% |
| Employed → Nonemployed | 18,173 | 4.5% |
| Employed → Employed | 171,315 | 42.5% |

Implied naïve transition probabilities:

- $\hat\theta_1^{\text{naïve}} = P(\text{emp}_{t+1} \mid \text{emp}_t) = 0.904$
- $\hat\theta_0^{\text{naïve}} = P(\text{emp}_{t+1} \mid \text{nonemp}_t) = 0.084$
- Exit rate: $1 - \hat\theta_1 = 9.6\%$
- Entry rate: $\hat\theta_0 = 8.4\%$

### Duration data

For employed individuals at wave 1:

| Statistic | Tenure (years) |
|:---|---:|
| Min | 0.02 |
| 1st Quartile | 1.17 |
| Median | 3.83 |
| Mean | 6.52 |
| 3rd Quartile | 9.33 |
| Max | 69.92 |

For nonemployed individuals at wave 1:

| Statistic | Timegap (years) |
|:---|---:|
| Min | 0.00 |
| 1st Quartile | 0.33 |
| Median | 0.88 |
| Mean | 2.33 |
| 3rd Quartile | 3.25 |
| Max | 7.50 |

::: {.callout-note}
Timegap is observed as **categorical** (7 bins: [0,3), [3,6), [6,9), [9,12),
[12,36), [36,60), [60,∞) months) and mapped to midpoints.
55% of wave-1 nonemployed have **never worked** — these are excluded from the
EM-tenure estimation sample.
:::

---

## What Works

### Simple binary-sequence MLE (no duration data)

The existing model that uses **only the 3-wave binary employment sequences**
$(y_1, y_2, y_3)$ — without any tenure or timegap information — gives plausible
results:

| Parameter | Estimate | Interpretation |
|:---|---:|:---|
| $\pi$ | ~3% | Misclassification rate |
| $\theta_0$ | ~3–4% | Entry rate (corrected) |
| $\theta_1$ | ~96–97% | Retention rate (corrected) |

These are consistent with the literature and pass sense-checks.

### Synthetic data validation of the EM-tenure code

The EM-tenure algorithm **correctly recovers known true parameters** on
synthetic data generated from the model:

| Parameter | True | Estimated (n=5000) |
|:---|---:|---:|
| $\alpha$ | 0.55 | 0.551 |
| $\theta_1$ | 0.97 | 0.969 |
| $\theta_0$ | 0.03 | 0.048 |
| $\pi$ | 0.03 | 0.026 |
| $\sigma^2_g$ | 0.005 | 0.005 |

The code is correct. Log-likelihood is monotonically non-decreasing. The problem
is not a bug — it is a model–data mismatch.

---

## The Current Problem {#sec-problem}

### EM-tenure diverges on real data

Running the EM-tenure model on the QLFS data (with misclassification,
non-stationary, discrete timegap mode) produces **implausible estimates**:

```
EM iter 1 | ll = -1.87e9 | theta1=0.628 theta0=0.419 pi=0.076 sig2g=3.08
EM iter 2 | ll = -1.60e9 | theta1=0.568 theta0=0.457 pi=0.086 sig2g=1.31
EM iter 3 | ll = -1.53e9 | theta1=0.540 theta0=0.456 pi=0.103 sig2g=0.56
EM iter 4 | ll = -1.47e9 | theta1=0.522 theta0=0.448 pi=0.121 sig2g=0.25
EM iter 5 | ll = -1.41e9 | theta1=0.508 theta0=0.439 pi=0.139 sig2g=0.12
```

::: {.callout-important}
**$\theta_1 \approx 0.50$ and $\theta_0 \approx 0.45$ are nonsensical.**

They imply employed and nonemployed individuals have roughly equal probability
of being employed next quarter — contradicting the strong persistence we see in
the raw data (90.4% vs 8.4%).

The simple model without tenure gives $\theta_1 \approx 0.97$, $\theta_0 \approx 0.03$. The EM-tenure model is converging to a completely different (and wrong) part of the parameter space.
:::

---

## The CTMC Link: Why We Used It & Why It Fails {#sec-ctmc}

### What is the CTMC link?

The model assumes a **continuous-time Markov chain** (CTMC) underlies
employment dynamics. This creates a deterministic mapping between transition
probabilities and duration distribution rates:

$$
\lambda_g = \frac{-\log(1 - \theta_1)}{3}, \qquad
\lambda_d = \frac{-\log(1 - \theta_0)}{3}
$$

where $\lambda_g$ is the exponential rate for employment spells (tenure)
and $\lambda_d$ is the rate for nonemployment spells (timegap). The spacing
$\Delta = 3$ months is one quarter.

This means $\theta_1$ simultaneously controls:

1. The **Markov prior**: $P(\text{employed}_t \mid \text{employed}_{t-1})$
2. The **tenure distribution**: $G \sim \text{Exp}(\lambda_g(\theta_1))$, i.e., $E[G] = 1/\lambda_g$

And $\theta_0$ simultaneously controls:

1. The **Markov prior**: $P(\text{employed}_t \mid \text{nonemployed}_{t-1})$
2. The **timegap distribution**: $D \sim \text{Exp}(\lambda_d(\theta_0))$, i.e., $E[D] = 1/\lambda_d$


### Reasons for using the CTMC link

::: {.callout-tip title="Arguments in favour of the linked specification"}

**1. Parsimony (6 vs 8 parameters)**

The linked model has parameters $(\alpha, \theta_0, \theta_1, \pi, \sigma^2_g, \sigma^2_d)$ — just 6.
The free model adds $\lambda_g$ and $\lambda_d$ as independent parameters — 8 total.
With only 3 waves (2 transitions per person), fewer parameters is better.

**2. Identification of $\lambda$ and $\sigma^2$**

The EMG density (exponential + Gaussian) for left-censored spells confounds
$\lambda$ and $\sigma^2$. A toy example:

> **Toy: EMG is under-determined when λ is free**
>
> An EMG$(λ, σ^2)$ distribution has mean $1/λ$ and variance $1/λ^2 + σ^2$.
> Suppose we observe spells with sample mean $\bar{x} = 5$ years and sample
> variance $s^2 = 30$ years².
>
> - Solution A: $λ = 0.20 \Rightarrow σ^2 = 30 - 25 = 5$
> - Solution B: $λ = 0.50 \Rightarrow σ^2 = 30 - 4 = 26$
> - Solution C: $λ = 0.17 \Rightarrow σ^2 = 30 - 35 < 0$ (infeasible)
>
> Infinitely many $(λ, σ^2)$ pairs fit the same moments. The link **pins** $λ$
> via the transition data, resolving the ambiguity.

```{r}
#| label: toy-emg-identification
#| code-summary: "Verify: EMG moment-matching is under-determined"

# EMG(lambda, sigma2) has mean = 1/lambda, var = 1/lambda^2 + sigma2
xbar <- 5   # observed mean (years)
s2   <- 30  # observed variance (years^2)

lambdas <- c(0.20, 0.50, 0.17)
results <- data.frame(
  lambda     = lambdas,
  implied_mean = 1 / lambdas,
  sigma2     = s2 - 1 / lambdas^2,
  feasible   = (s2 - 1 / lambdas^2) > 0
)
results
# Multiple (lambda, sigma2) pairs yield the same moments — not identified
```

**3. Structural consistency**

If the true process is a two-state CTMC, then spell durations *must* be
exponential and $\lambda$ *must* equal the CTMC intensity. The link enforces
this consistency rather than allowing the two data sources to tell contradictory
stories.

**4. Computational simplicity**

The free specification requires nested numerical optimisation of the EMG
log-likelihood at each M-step. The linked specification avoids this — $\lambda$
is deterministic from $\theta$, and the joint FOC is a simple 1D root-finding
problem (Brent's method, 2–5 iterations).
:::


### Why the CTMC link fails on real data

::: {.callout-warning title="The fundamental tension"}
The CTMC link forces $\theta$ to satisfy **two conflicting signals simultaneously**:

- Binary transition data says: $\theta_1 \approx 0.97$
- Tenure distribution data says: $\theta_1$ should be much lower (to match
  the long observed spells)

The M-step FOC is:

$$
\underbrace{\frac{T_{11}}{\theta_1} - \frac{D_1 - T_{11}}{1-\theta_1}}_{\text{Markov prior: wants } \theta_1 \approx 0.97}
\;+\;
\underbrace{\frac{1}{3(1-\theta_1)} \cdot E_g}_{\text{EMG emission: pulls } \theta_1 \text{ down}}
\;=\; 0
$$

The EM "compromises" at $\theta_1 \approx 0.5$, which satisfies neither signal.
:::

Here is the core conflict, demonstrated with arithmetic:

::: {.callout-important title="Toy Calculation: The 60× Discrepancy"}

**What $\theta_1 = 0.97$ implies for tenure:**

$$\lambda_g = \frac{-\log(1 - 0.97)}{3} = \frac{-\log(0.03)}{3} = \frac{3.51}{3} = 1.17 \text{ /month}$$

$$E[\text{tenure}] = \frac{1}{\lambda_g} = \frac{1}{1.17} = 0.85 \text{ months} \approx 0.07 \text{ years}$$

**What the data actually shows:**

- Median tenure: **3.83 years** (46 months)
- Mean tenure: **6.52 years** (78 months)

**The CTMC with $\theta_1 = 0.97$ predicts a mean tenure of 0.85 months.
The observed mean tenure is 78 months — a factor of ~92× larger.**

The only way to match long tenure spells under the link is to push $\theta_1$
down, which increases $1/\lambda_g$. For example, $\theta_1 = 0.50$ gives:

$$\lambda_g = \frac{-\log(0.50)}{3} = 0.231 \text{ /month} \quad \Rightarrow \quad E[\text{tenure}] = 4.3 \text{ months}$$

Still much too short, but at least in the same order of magnitude — which is
exactly where the EM converges.

```{r}
#| label: toy-tenure-discrepancy
#| code-summary: "Verify: CTMC-implied vs observed tenure"

# CTMC link: lambda_g = -log(1 - theta1) / 3
ctmc_lambda <- function(theta) -log(1 - theta) / 3

# What theta1 = 0.97 implies
theta1_true <- 0.97
lambda_g_97 <- ctmc_lambda(theta1_true)
E_tenure_97 <- 1 / lambda_g_97  # in months

cat(sprintf("theta1 = %.2f  =>  lambda_g = %.4f /month  =>  E[tenure] = %.2f months (%.2f years)\n",
            theta1_true, lambda_g_97, E_tenure_97, E_tenure_97 / 12))

# Observed tenure stats (from data diagnostics)
obs_median_tenure_yrs <- 3.83
obs_mean_tenure_yrs   <- 6.52

cat(sprintf("Observed mean tenure:   %.1f months (%.2f years)\n",
            obs_mean_tenure_yrs * 12, obs_mean_tenure_yrs))
cat(sprintf("Discrepancy factor:    %.0f×\n",
            (obs_mean_tenure_yrs * 12) / E_tenure_97))

# What theta1 = 0.50 implies (where the EM converges)
theta1_em <- 0.50
lambda_g_50 <- ctmc_lambda(theta1_em)
E_tenure_50 <- 1 / lambda_g_50

cat(sprintf("\ntheta1 = %.2f  =>  lambda_g = %.4f /month  =>  E[tenure] = %.1f months (%.2f years)\n",
            theta1_em, lambda_g_50, E_tenure_50, E_tenure_50 / 12))

# What theta1 would we NEED to match observed mean tenure?
# E[tenure] = 1/lambda = 78 months => lambda = 1/78
# theta = 1 - exp(-lambda * 3) = 1 - exp(-3/78)
theta1_needed <- 1 - exp(-3 / (obs_mean_tenure_yrs * 12))
cat(sprintf("\nTo match observed mean tenure (%.0f months): theta1 = %.4f\n",
            obs_mean_tenure_yrs * 12, theta1_needed))
cat("This is far below the transition-data estimate of 0.97\n")
```
:::

::: {.callout-important title="Toy Calculation: Timegap side is less severe but still mismatched"}

**What $\theta_0 = 0.08$ implies for nonemployment duration:**

$$\lambda_d = \frac{-\log(1 - 0.08)}{3} = \frac{-\log(0.92)}{3} = \frac{0.083}{3} = 0.028 \text{ /month}$$

$$E[\text{timegap}] = \frac{1}{\lambda_d} = \frac{1}{0.028} = 36 \text{ months} = 3.0 \text{ years}$$

**Observed mean timegap: 2.33 years (28 months)** — actually in the right
ballpark, but this includes the coarse categorical measurement (7 bins) and
the data is interval-censored, so the comparison is approximate.

The mismatch is much less severe on the timegap side, but the fundamental issue
is the same: forcing a single parameter to satisfy two constraints.

```{r}
#| label: toy-timegap-discrepancy
#| code-summary: "Verify: CTMC-implied vs observed timegap"

ctmc_lambda <- function(theta) -log(1 - theta) / 3

# What theta0 = 0.08 (naive observed) implies
theta0_naive <- 0.084
lambda_d <- ctmc_lambda(theta0_naive)
E_timegap <- 1 / lambda_d

cat(sprintf("theta0 = %.3f  =>  lambda_d = %.4f /month  =>  E[timegap] = %.1f months (%.2f years)\n",
            theta0_naive, lambda_d, E_timegap, E_timegap / 12))

# Observed timegap stats
obs_mean_timegap_yrs <- 2.33
cat(sprintf("Observed mean timegap: %.1f months (%.2f years)\n",
            obs_mean_timegap_yrs * 12, obs_mean_timegap_yrs))
cat(sprintf("Ratio (implied/observed): %.1f×\n", E_timegap / (obs_mean_timegap_yrs * 12)))

# What theta0 = 0.03 (simple model estimate) implies
theta0_simple <- 0.03
lambda_d_03 <- ctmc_lambda(theta0_simple)
E_timegap_03 <- 1 / lambda_d_03
cat(sprintf("\ntheta0 = %.2f (simple model)  =>  E[timegap] = %.0f months (%.1f years)\n",
            theta0_simple, E_timegap_03, E_timegap_03 / 12))
cat("Also in the right ballpark — timegap side is less problematic\n")
```
:::

```{r}
#| label: toy-mapping-table
#| code-summary: "Full mapping: theta -> implied mean spell duration"

ctmc_lambda <- function(theta) -log(1 - theta) / 3

thetas <- c(0.97, 0.90, 0.80, 0.50, 0.30, 0.10, 0.08, 0.03)
mapping <- data.frame(
  theta        = thetas,
  lambda       = round(ctmc_lambda(thetas), 4),
  E_spell_months = round(1 / ctmc_lambda(thetas), 1),
  E_spell_years  = round(1 / ctmc_lambda(thetas) / 12, 2)
)
cat("CTMC link: theta -> lambda -> E[spell duration]\n")
cat("(lambda in per-month units, delta = 3 months)\n\n")
mapping
```

::: {.callout-note title="Toy Calculation: Why synthetic data works"}

In synthetic data, durations are **generated from** the exponential model:

$$G = \text{Exp}(\lambda_g) + \epsilon_g, \quad \epsilon_g \sim N(0, \sigma^2_g)$$

The CTMC link is *exactly correct by construction*. True $\theta_1 = 0.97$
generates short spells ($E[G] \approx 0.85$ months) as expected.

In real data, employment spells exhibit **positive duration dependence** — longer
spells are more stable. The tenure distribution (median 3.8 yrs, mean 6.5 yrs,
max 70 yrs) has a heavy right tail completely inconsistent with any single
exponential rate. The memoryless property of the exponential fundamentally does
not hold for real employment tenure.
:::


```{r}
#| label: summary-tension
#| code-summary: "Summary: the two signals and their implied theta"

ctmc_lambda <- function(theta) -log(1 - theta) / 3
ctmc_theta  <- function(lambda) 1 - exp(-lambda * 3)

cat("=== EMPLOYMENT SIDE (theta1) ===")
cat("\n\nSignal 1 (transition counts): theta1 ≈ 0.904 (naive) or 0.97 (corrected)")
cat("\n  => lambda_g =", round(ctmc_lambda(0.97), 4), "/month")
cat("\n  => E[tenure] =", round(1/ctmc_lambda(0.97), 1), "months")
cat("\n  => But observed mean tenure = 78 months\n")

cat("\nSignal 2 (tenure distribution): need E[tenure] ≈ 78 months")
cat("\n  => lambda_g = 1/78 =", round(1/78, 5), "/month")
cat("\n  => theta1 =", round(ctmc_theta(1/78), 4))
cat("\n  => But transition data says theta1 ≈ 0.97\n")

cat("\nEM compromise: theta1 ≈ 0.50")
cat("\n  => E[tenure] =", round(1/ctmc_lambda(0.50), 1), "months (too short)")
cat("\n  => P(stay employed) = 0.50 (too low)")
cat("\n  => Satisfies NEITHER signal\n")

cat("\n=== NONEMPLOYMENT SIDE (theta0) ===")
cat("\n\nSignal 1 (transition counts): theta0 ≈ 0.084 (naive) or 0.03 (corrected)")
cat("\nSignal 2 (timegap distribution): mean timegap ≈ 28 months")
cat("\n  => lambda_d = 1/28 =", round(1/28, 5))
cat("\n  => theta0 =", round(ctmc_theta(1/28), 4))
cat("\n  Less severe mismatch — signals are in the same ballpark\n")
```

---

## Root Cause Diagnosis Summary

The estimation failure has **one primary cause and several contributing factors**:

### Primary: Model misspecification (exponential duration assumption)

The CTMC assumption implies exponential (memoryless) spell durations.
Real employment tenure is heavy-tailed with strong positive duration
dependence. This creates an irreconcilable tension within the linked
specification:

- **Transition counts** identify $\theta_1 \approx 0.97$ (short exit hazard)
- **Tenure levels** demand $\lambda_g$ be small (long spells), which via the
  link requires $\theta_1$ to be small
- The EM compromise sits at $\theta_1 \approx 0.5$, satisfying neither

### Contributing: Wrong-state duration imputation

The ingest script imputes "plausible" durations for the wrong state
(e.g., timegap values for employed waves) via nearest-non-zero donor
imputation. These fabricated values enter the likelihood for all
misclassifying latent histories. If the imputed values are "too
plausible," they inflate $\pi$ and further distort $\theta$.

### Contributing: Local optima

With 8 latent histories and a mixture likelihood, the EM has
multiple local optima. The default initialisation ($\theta_1 = 0.9$,
$\theta_0 = 0.1$) may place the algorithm in the basin of the wrong
optimum — but even with initialisation closer to the truth, the
fundamental tension from the CTMC link would remain.

### Contributing: Extreme tenure values and right-tail leverage {#sec-tail}

The tenure distribution has a very heavy right tail (mean 6.5 yrs vs median
3.8 yrs, max 70 yrs). Under the exponential model, each observation contributes
a term $-\lambda_g \cdot g_i$ to the log-likelihood. For an individual with
$g = 70$ years (840 months), this term dominates. A few hundred extreme-tenure
observations can overwhelm the transition-count signal from the entire sample.

::: {.callout-warning title="Extreme tenure leverage on the likelihood"}
Under the EMG/exponential model, the log-likelihood contribution of one
observation with tenure $g$ includes $-\lambda_g \cdot g$. At $\theta_1 = 0.97$
($\lambda_g = 1.17$/month), a 70-year tenure contributes $-1.17 \times 840 = -983$
to the log-likelihood — a catastrophic penalty. The EM reduces $\lambda_g$ (and
hence $\theta_1$) to make these observations less costly, even though doing so
wrecks the transition-probability estimates.
:::

```{r}
#| label: toy-tail-leverage
#| code-summary: "Verify: extreme tenure observations' leverage on log-likelihood"

ctmc_lambda <- function(theta) -log(1 - theta) / 3

# Log-likelihood contribution of Exp(lambda) at observation g:
#   log(lambda) - lambda * g
ll_exp <- function(g_months, lambda) log(lambda) - lambda * g_months

# Representative observations
g_vals <- c(0.5, 1, 3.83, 6.52, 20, 40, 70) # in years
g_months <- g_vals * 12

# Compare LL contributions at theta1 = 0.97 vs 0.50
lambda_97 <- ctmc_lambda(0.97)
lambda_50 <- ctmc_lambda(0.50)

leverage <- data.frame(
  tenure_yrs = g_vals,
  tenure_months = g_months,
  ll_theta097 = round(ll_exp(g_months, lambda_97), 1),
  ll_theta050 = round(ll_exp(g_months, lambda_50), 1),
  ll_gain_from_lower_theta = round(ll_exp(g_months, lambda_50) - ll_exp(g_months, lambda_97), 1)
)

cat("Log-likelihood contribution per observation (exponential component only)\n")
cat("Positive 'll_gain' = observation prefers theta1=0.50 over theta1=0.97\n\n")
leverage

cat("\n\nKey insight: the 70-year tenure observation gains",
    round(ll_exp(70*12, lambda_50) - ll_exp(70*12, lambda_97), 0),
    "log-lik points\nby moving theta1 from 0.97 to 0.50.")
cat("\nA median-tenure (3.83 yr) observation loses only",
    round(abs(ll_exp(3.83*12, lambda_50) - ll_exp(3.83*12, lambda_97)), 1),
    "points.\n")
cat("So a SINGLE 70-year obs outweighs ~",
    round(abs(ll_exp(70*12, lambda_50) - ll_exp(70*12, lambda_97)) /
          abs(ll_exp(3.83*12, lambda_50) - ll_exp(3.83*12, lambda_97))),
    " median-tenure observations.\n")
```

::: {.callout-note}
**Diagnostic to try**: Trim or winsorise tenure at the 99th percentile and
re-run the EM. If $\theta_1$ moves substantially toward 0.97, the right tail
is a major driver — independent of the broader CTMC misspecification question.
This is a quick test that doesn't require any model changes.
:::

### Contributing: Coarse timegap categories

Nonemployment duration is observed in 7 coarse bins. While the
discrete interval-censored model handles this correctly in
principle, the coarseness reduces the information content and may
interact with the exponential misspecification.

---

## Proposed Next Steps

### Immediate diagnostics (I can do these now)

1. **Inspect sufficient statistics at each iteration**: Print $T_{11}/D_1$
   and $T_{01}/D_0$ (the Markov-only $\theta$ estimates) alongside the
   EMG gradient $E_g$ to confirm the tension quantitatively.

2. **Run with free $\lambda$**: Decouple $\lambda_g$ and $\lambda_d$ from
   $\theta$ and re-estimate. If the Markov-only estimates normalise to
   $\theta_1 \approx 0.97$, $\theta_0 \approx 0.03$, this confirms the
   link is the sole driver.

3. **Trim the tenure right tail**: Winsorise tenure at the 95th or 99th
   percentile and re-run. If $\theta_1$ moves substantially toward 0.97,
   the extreme-value leverage (see @sec-tail) is a major driver and may
   be addressable without changing the model structure.

4. **Try initialisation closer to truth**: Set $\theta_1 = 0.97$,
   $\theta_0 = 0.03$, $\pi = 0.03$ and check whether the EM stays near
   these values or drifts away.

### Model modifications (need supervisor input)

4. **Drop the CTMC link entirely**: Use $\lambda_g$ and $\lambda_d$ as
   free parameters (or reparametrise via a more flexible duration model).
   The free specification has identification concerns (see
   @sec-ctmc), but these may be resolvable with
   regularisation or informative priors.

5. **Replace exponential with a more flexible duration model**: Options
   include:
   - **Mixture of exponentials**: $G \sim p \cdot \text{Exp}(\lambda_1) + (1-p) \cdot \text{Exp}(\lambda_2)$ — captures short and long spells with 3 parameters per side
   - **Weibull**: $f(t) = k\lambda(t\lambda)^{k-1}e^{-(t\lambda)^k}$ — shape parameter $k > 1$ captures positive duration dependence
   - **Piecewise exponential**: different rates for different tenure segments
   - Each option breaks the clean CTMC link but may be necessary

6. **Use duration data only for identification, not via the link**:
   Keep the binary-sequence Markov model for $\theta$ estimation.
   Use tenure/timegap data separately to identify $\pi$ via the
   different distributional shapes of correctly-classified vs
   misclassified observations — without linking $\lambda$ to $\theta$.

---

## Help Needed from Supervisor {#sec-help}

::: {.callout-caution title="Questions for discussion"}

**Q1. Is the exponential duration assumption viable for South African
labour markets?**

Observed tenure has median 3.8 years and mean 6.5 years with a max of
70 years. This heavy right tail is inconsistent with memoryless
exponential spells. Is there an economic argument for maintaining the
CTMC assumption, or should we accept that duration dependence is
present and move to a more flexible model?

**Q2. Should we run a free-$\lambda$ diagnostic?**

The TeX document argues the free specification is "potentially poorly
identified" and "theoretically redundant." But running it as a
*diagnostic* (not final model) would isolate whether the link itself
drives the divergence vs other data issues. Is this worth doing?

**Q3. What is the best path to a more flexible duration model?**

If we drop the CTMC link, we need an alternative duration model that:

- Is flexible enough to capture heavy-tailed tenure distributions
- Still allows identification of $\pi$ (misclassification)
- Doesn't add too many parameters given only 3 waves

The leading candidates are mixture-of-exponentials, Weibull, or
dropping the duration emission from $\theta$ estimation entirely. Which
direction makes the most economic sense?

**Q4. Should we consider a staged approach?**

- **Stage 1**: Estimate $(\alpha, \theta_0, \theta_1, \pi)$ from
  binary sequences only (this already works and gives plausible
  results)
- **Stage 2**: Conditional on $\hat\theta$, estimate the duration
  parameters $(\sigma^2_g, \lambda_g)$ to validate or refine the
  misclassification estimate

This sacrifices joint efficiency but avoids the link tension entirely.
:::

---

## Appendix: Key References Within the Project

| File | Description |
|:---|:---|
| `EM-tenure/estimate_pipeline.R` | End-to-end estimation script (currently running) |
| `EM-tenure/R/` | All EM functions (E-step, M-step, emissions, driver) |
| `EM-tenure/CONCERNS.md` | Known methodological concerns (13 items) |
| `troubleshoot/diagnose_em_tenure.R` | Diagnostic script (§1–§5) |
| `documents/EM tenure.tex` | Full model derivation |
| `scripts/ingest_data_3waves_SA.R` | Data ingestion and imputation |
