---
title: "Covariate-model revisions on `codex/rulof_new_work`"
subtitle: "Changes relative to the `efficient` branch"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: true
geometry: margin=1in
header-includes:
  - \usepackage{booktabs}
  - \usepackage{amsmath}
  - \usepackage{float}
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
project_root <- normalizePath("..", winslash = "/", mustWork = TRUE)

include_tex_table <- function(relative_path, tabular_scale = NULL) {
  path <- file.path(project_root, relative_path)
  if (!file.exists(path)) {
    stop("Missing generated table: ", path,
         ". Run Rscript build_tables.R from the project root first.")
  }
  lines <- readLines(path, warn = FALSE)
  if (!is.null(tabular_scale)) {
    begin_tabular <- grep("^\\\\begin\\{tabular\\}", lines)
    end_tabular <- grep("^\\\\end\\{tabular\\}", lines)
    if (length(begin_tabular) != 1L || length(end_tabular) != 1L)
      stop("Expected exactly one tabular environment in ", path)
    lines <- append(lines, "}", after = end_tabular)
    lines <- append(
      lines,
      sprintf("\\scalebox{%s}{%%", tabular_scale),
      after = begin_tabular - 1L
    )
  }
  cat(lines, sep = "\n")
}
```

# Purpose and scope

This note records the changes made on `codex/rulof_new_work` relative to the
`efficient` branch. It concentrates on the observable-covariate extension and
the revised version of Table 4. The baseline, finite-mixture, inconsistency,
AR(2), and tenure specifications are not reinterpreted here because the work on
this branch did not substantively change their economic specification.

The central objective was to replace numerically unreliable covariate-model
solutions with valid maximum-likelihood estimates, handle contract type without
discarding non-employed respondents, add informal-sector status to Set 3, and
report heterogeneous transition risks in a way that remains comparable with
the paper's earlier tables.

# Changes to the code and estimation approach

## Numerical optimisation

The covariate model is estimated by a Generalised EM algorithm. The earlier
M-step used separate probit IRLS updates. That approach did not maximise the
correct joint expected complete-data likelihood when the initial-state term
coupled the entry and persistence equations, and it could return a solution
whose likelihood was below that of the nested no-misclassification model.

The revised estimator:

1. normalises survey weights during optimisation to improve numerical
   conditioning, while restoring the original likelihood scale in saved output;
2. updates the entry and persistence coefficient blocks jointly with limited
   BFGS steps;
3. backtracks until both the complete-data objective and observed-data
   likelihood are non-decreasing;
4. uses several starting values and initializes each symmetric-error model from
   its corresponding no-error estimate;
5. explicitly checks that the symmetric model's maximised likelihood is no
   lower than that of its nested no-error model; and
6. requires both a small likelihood improvement and a small parameter change
   before declaring convergence.

These safeguards address the invalid-MLE problem directly: they alter how the
model is estimated, rather than merely changing how its results are displayed.

## Contract type and informal-sector status

Contract type is structurally undefined when a respondent is not employed.
The adopted treatment is therefore to include it only in the
employment-persistence equation. Its coefficient in the job-entry equation is
fixed at zero, so non-employed observations are retained instead of being
dropped or assigned an economically artificial contract category.

Set 3 now also includes an informal-sector indicator, defined from the QLFS
sector classification and including informal agricultural employment. Like
contract type, it enters only the persistence/exit equation. Rural or urban
residence is not included, avoiding the sample loss caused by missing residence
information.

Both employment characteristics are measured at the origin of each transition:
wave 1 values are used for the wave 1--2 transition and wave 2 values for the
wave 2--3 transition. They can consequently change between waves. Employment
is still represented by a single latent state; formal and informal employment
are not modelled as separate states.

## Initial conditions and reported transition rates

Because the Set 3 transition probabilities can vary between the two transition
periods, the main comparison no longer imposes a time-homogeneous stationary
initial distribution. All six columns of the revised Table 4 estimate the
initial employment probability freely.

Average entry and exit rates are computed using the survey weights and the
posterior probability that each observation belongs to the appropriate latent
origin-state risk set. This makes the averages conditional hazards: entry is
averaged over latent non-employment exposure and exit over latent employment
exposure. The table also reports the posterior employed risk share and total
employment churn. The separate entry-flow and exit-flow rows have been removed
to keep the presentation close to the earlier implied-probability tables.

## Analytical standard errors

Pending a full bootstrap, inference for the covariate models uses an
individual-level survey-weighted sandwich estimator. The observed-information
``bread'' is obtained by numerically differentiating the analytic observed-data
score, while the ``meat'' is the outer product of weighted individual scores.
The delta method propagates this covariance matrix to average transition rates,
total churn, and Set 3 discrete effects. The linearisation also includes direct
sampling variation from averaging predicted quantities over the observed
covariate and posterior-risk distribution.

These are approximate analytical standard errors. They use survey weights but
do not incorporate stratification or PSU clustering because those identifiers
are unavailable in the estimation extract. A person-level bootstrap remains a
useful robustness exercise.

## Output and reproducibility

The extension pipeline saves fitted model objects and convergence diagnostics.
`EM-baseline-ext/estimate_analytical_se_table4.R` produces the analytical
covariance estimates, and `build_tables.R` creates the LaTeX results below.
Generated model objects and tables remain excluded from version control and can
be rebuilt locally from the underlying confidential QLFS files.

# New and revised tables

The following are the two tables introduced by this branch: the revised main
Table 4 and its new coefficient/distribution appendix. Unchanged tables already
present on `efficient` are not duplicated.

## Revised Table 4

```{r revised-table-4, results='asis'}
include_tex_table("EM-baseline-ext/output/tables/table_cov_risk_weighted.tex")
```

The no-error specifications imply quarterly entry and exit rates of about
8.14% and 8.58%, almost invariant to the three covariate sets. Allowing for
symmetric reporting error lowers these rates to 3.62--3.99% and 3.71--4.16%,
respectively. The estimated per-wave misclassification probability remains
narrowly between 2.39% and 2.58%. Observable characteristics therefore explain
substantial heterogeneity across workers, but they do not explain away the
measurement-error signal. Ignoring misclassification still approximately
doubles the estimated average transition hazards and total churn.

The exit rate is only modestly higher than the entry rate in the symmetric
models. In Set 3 the difference is about 0.17 percentage points, compared with
roughly 0.44 percentage points in the no-error specification. This is more
consistent with the approximately stable employment share than a comparison of
the two conditional hazards alone might initially suggest.

## Coefficients, hazard distributions, and Set 3 discrete effects

```{r coefficient-appendix, results='asis'}
include_tex_table(
  "EM-baseline-ext/output/tables/table_cov_coefficients_appendix.tex",
  tabular_scale = 0.62
)
```

The appendix demonstrates why a single mean hazard is incomplete. In symmetric
Set 3, the risk-weighted 10th--90th percentile range is approximately
1.07--7.97% for entry and 0.46--9.61% for exit. Thus workers with different
observed characteristics face markedly different transition risks even though
the population-average entry and exit rates are similar.

The economically clearest Set 3 results concern employment stability. Holding
the other covariates at their observed values, a permanent contract lowers the
average quarterly exit probability by about 3.38 percentage points in the
symmetric-error model. Informal-sector employment raises it by about 1.80
percentage points. The corresponding no-error effects are larger in magnitude
(-4.69 and +2.65 percentage points), showing that some observed instability is
otherwise attributed to genuine labour-market transitions rather than reporting
error.

# Interpretation for the paper

The revisions strengthen the paper's core result. The earlier numerical problem
could have made the misclassification specification appear to fit worse than a
model it mathematically nests. Once the likelihood is optimised with appropriate
monotonicity and nesting checks, a small but precisely estimated reporting-error
rate continues to produce a large reduction in inferred employment turnover.

At the same time, the new Set 3 results sharpen the economic interpretation.
Misclassification and genuine heterogeneity are not competing explanations in
which only one survives: contract security and informal employment strongly
predict true exit risk, while a separate reporting-error component remains.
The paper should therefore distinguish clearly between variation in transition
risk across workers and the upward bias in observed transitions caused by
classification error.

# Remaining checks

Before final submission, the analytical standard errors should ideally be
compared with a person-level bootstrap. If the survey extract can be augmented
with strata and PSU identifiers, design-based clustered inference would also be
preferable. The main qualitative conclusions should be assessed against those
two alternatives, while retaining the current analytical estimates as a fast
and reproducible diagnostic.
