---
title: "Estimation revisions on `codex/rulof_new_work`"
subtitle: "Changes relative to the `efficient` branch"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: true
geometry: margin=1in
header-includes:
  - \usepackage{booktabs}
  - \usepackage{amsmath}
  - \usepackage{float}
  - \usepackage{caption}
  - \usepackage{needspace}
  - \usepackage{graphicx}
  - \usepackage{longtable}
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
project_root <- normalizePath("..", winslash = "/", mustWork = TRUE)

include_tex_table <- function(relative_path, tabular_scale = NULL,
                              table_number = NULL, appendix_label = NULL) {
  path <- file.path(project_root, relative_path)
  if (!file.exists(path)) {
    stop("Missing generated table: ", path,
         ". Run Rscript build_tables.R from the project root first.")
  }
  lines <- readLines(path, warn = FALSE)
  lines <- sub("^\\\\begin\\{table\\}\\[[^]]+\\]$", "\\\\begin{table}[H]", lines)
  if (!is.null(table_number)) {
    lines <- c(sprintf("\\setcounter{table}{%d}", as.integer(table_number) - 1L),
               lines)
  }
  if (!is.null(appendix_label)) {
    caption_line <- grep("^\\\\caption\\{", lines)
    if (length(caption_line) != 1L)
      stop("Expected exactly one caption in ", path)
    caption_text <- sub("^\\\\caption\\{(.*)\\}$", "\\1", lines[caption_line])
    lines[caption_line] <- sprintf("\\caption*{%s: %s}", appendix_label,
                                   caption_text)
  }
  if (!is.null(tabular_scale)) {
    begin_tabular <- grep("^\\\\begin\\{tabular\\}", lines)
    end_tabular <- grep("^\\\\end\\{tabular\\}", lines)
    if (length(begin_tabular) != 1L || length(end_tabular) != 1L)
      stop("Expected exactly one tabular environment in ", path)
    lines <- append(lines, "}", after = end_tabular)
    lines <- append(lines, sprintf("\\scalebox{%s}{%%", tabular_scale),
                    after = begin_tabular - 1L)
  }
  cat(lines, sep = "\n")
}

# The paper-facing sections and tables have one authoritative source. The long
# summary renders the short summary body verbatim, then adds technical notes.
short_path <- file.path(dirname(knitr::current_input(dir = TRUE)),
                        "branch_changes_summary_short.Rmd")
short_lines <- readLines(short_path, warn = FALSE)
body_start <- grep("^# 1\\. Baseline transition model", short_lines)
if (length(body_start) != 1L)
  stop("Could not locate the numbered body in branch_changes_summary_short.Rmd")
short_body <- short_lines[body_start:length(short_lines)]
```

```{r synchronized-paper-facing-summary, results='asis'}
cat(knitr::knit_child(text = short_body, quiet = TRUE), sep = "\n")
```

\newpage

# Detailed changes to code and estimation

The preceding numbered sections, tables, interpretations, and to-do lists are
identical to the short summary. This section records the additional estimation
and implementation detail needed to reproduce and audit those results.

## Tables 1--2: exact baseline likelihood and matching robustness

The three-wave baseline estimator works with the exact eight possible observed
employment histories. Collapsing to these histories preserves the weighted
observed likelihood while making multiple-start optimization inexpensive. All
probabilities are represented by unconstrained logits. Stationary models impose
the restriction inside the likelihood; free-initial-condition models estimate
the initial employment probability separately.

The pipeline uses nested warm starts and dispersed deterministic starts, checks
the analytical score against numerical derivatives, requires positive-definite
observed information, and verifies that each unrestricted specification weakly
dominates its nested restriction. Saved fits carry the panel identity, sample
signature, sample size, total weight, likelihood, and optimizer diagnostics so
the table builder can reject stale results.

The main table now reports only free-initial-condition models. Appendix Table
A1 retains the stationary specifications, raw parameters, implied entry, exit,
false-positive and false-negative rates, and tests the stationarity restriction.
Because the fitted pseudo-likelihood uses population-scale survey weights, all
LR statistics normalize weights to sum to the estimation sample size. The test
of zero misclassification uses the 50:50 point-mass/$\chi^2_1$ boundary
reference; asymmetry and stationarity use the regular $\chi^2_1$ reference.

Table 2 reruns the nonstationary no-error and symmetric-error models on panels
A, B, and C. These panels apply progressively stricter longitudinal matching
rules. The sensitivity script changes the panel input explicitly rather than
silently modifying the default ingestion path, which remains
`df_qlfs_A.rds`.

## Table 3: observed indicators of linked-record reliability

The inconsistency likelihood permits the symmetric status-error probability to
depend on age inconsistency, education inconsistency, both inconsistencies, and
in the expanded models the magnitude of those discrepancies. The most recent
specifications also include indicators for observations retained in panel B
but not C and in panel A but not B. These indicators are constructed using the
same nested matching algorithm as Table 2.

The EM update for the error equation now maximizes its correct conditional
objective. Both stationary and free-initial-condition results are retained,
and the robustness model allows true entry and exit rates to differ across
reliability groups. Reported means use the same survey and posterior weighting
conventions as the other tables. The mechanism language remains deliberately
neutral: inconsistencies may reflect response problems, incorrect longitudinal
matches, or both.

## Table 4: observable transition heterogeneity

Sets 1--4 are nested. Set 1 contains age, age squared, and education; Set 2 adds
race and sex; Set 3 adds origin-wave log tenure to persistence and origin-wave
log time since work plus never-worked status to entry; Set 4 adds origin-wave
contract type and informal-sector status to persistence. Contract type and
sector may change between transitions. They do not enter entry because they are
undefined for a genuinely nonemployed origin state, so respondents are retained
without assigning artificial job characteristics.

The eight constant-error/no-error specifications retain the repaired
generalized-EM implementation: normalized weights, joint transition-block
updates, likelihood-preserving backtracking, nested starts, and explicit
convergence and nesting checks. Posterior risk-set and survey-weighted hazards
replace unweighted averages. Individual-level survey-weighted sandwich
covariances and delta-method transformations provide approximate analytical
errors where bootstrap estimates are unavailable.

Column (9) is estimated by the exact observed likelihood rather than by the old
separable M-step. It combines the Set 4 transition equations with a symmetric
error probability depending on the full Table 3 reliability specification:
age and education inconsistencies, severity indicators, panel B but not C, and
panel A but not B. Five starts were evaluated. The polished solution has 30 of
30 information-matrix directions identified, condition number about 18,126,
and maximum normalized score $1.1\times10^{-8}$. The new mean entry, exit, and
error rates are 4.67%, 4.62%, and 2.55%. Detailed error slopes and implied
reliability-specific rates appear in Appendix Table A3B.

## Table 5: four-wave AR(1) and AR(2)

The four-wave modules enumerate latent histories and evaluate the exact
collapsed observed likelihood. The AR(1) benchmark has free initial employment
and symmetric error. The unrestricted AR(2) allows the previous two latent
states to determine the next transition. Its duration-controlled specification
adds the same state-appropriate tenure and time-since-work variables used in
Table 4 and permits those variables to change with the origin wave.

The analytical covariance uses the observed information and collapsed
person-level score outer products. The duration-controlled AR(2) information
matrix is locally full rank, although its high condition number means the
initial-state and missing-duration directions warrant caution.

The new extension is estimated as three exact nested likelihoods on the same
308,240 observations. The first combines AR(2) with the Table 4 Set 4 transition
equations and constant symmetric error. The second adds age/education
inconsistency and severity predictors to the error equation. The third adds
panel B-not-C and panel A-not-B matching-quality predictors. Stage-specific
collapsing leaves 197,933, 220,795, and 228,195 exact cells. Each stage uses two
substantively distinct starts, checks likelihood nesting, applies an
observed-information Newton correction where useful, and is accepted only with
a maximum normalized score below $2\times10^{-5}$ and full information rank.
The final scores are $2.8\times10^{-6}$, $4.2\times10^{-6}$, and
$2.8\times10^{-6}$. All three models are locally full rank, but condition
numbers rise from about 746,558 to 2,800,426, motivating profile and
long-start sensitivity checks. Appendix Table A4C is generated from a separate
prospective decomposition that conditions only on the two reports forming an
apparent transition and integrates out the subsequent report.

## Table 6: finite-mixture heterogeneity

The two-type four-wave model uses exact likelihood refinement after EM warm
starts, label normalization, multiple starts, and analytical sandwich/delta
inference. The comparison now contains the no-error FMM, the uncontrolled
symmetric-error FMM, the controlled common-error model, and the controlled
type-specific-error model in one table. Observed transition slopes are common
across latent types in the parsimonious controlled models, while type-specific
intercepts retain mobility heterogeneity.

Age and education inconsistencies determine error risk without assigning them
to a particular behavioral mechanism. The preferred type-specific model allows
the two classes to have different error intercepts and common inconsistency
slopes. Its local information matrix is full rank. The failed three-wave
symmetric-error FMM is not reported as an identified result: three waves do not
separate two-type transition heterogeneity from symmetric misclassification.

## Table 7: endogenous duration dependence

The tenure-contamination module preserves the central consistency restriction:
latent employment paths must agree with clean reported tenure and time-since-
work clocks. Successive specifications relax constant transition hazards,
integrate unavailable wrong-state clocks, use tail-safe piecewise hazards, and
allow alternative contamination processes for tenure and timegap reports.
These are likelihood changes, not presentation-only adjustments.

The preferred timegap treatment is the joint per-wave marginal-contamination
model. The table reports approximate analytical errors from the full numerical
observed-information matrix and delta method, not bootstrap errors. Appendix
Table A6 retains hazard profiles, auxiliary comparisons, wrong-state-clock
integration checks, and the within-quarter reset decomposition.

## Appendix Table A7: no-error duration counterfactual

The counterfactual diagnostic estimates flexible no-status-error AR(1) and
AR(2) transition models without using destination-wave durations in the
likelihood. It then propagates implied spell clocks and compares them with the
held-out reported tenure and timegap. Large conditional contradictions show
that reported employment changes and duration histories cannot generally both
be correct. The exercise motivates joint modeling but does not distinguish
response inconsistency from longitudinal matching error.

## Reproducibility and generated files

`build_tables.R` generates Tables 1, 3, and 4 and their coefficient appendices;
the matching-sensitivity runner generates Table 2. The four-wave and duration
modules retain their own reproducible runners, tests, result summaries, and
diagnostic outputs. Confidential data and intermediate fitted objects remain
ignored by Git. Curated paper-facing snapshots and both summary sources are
tracked. A complete rerun still requires the confidential QLFS inputs and the R
environment recorded by `renv`.

The remaining inference priorities are the person-level bootstrap for the main
covariate and duration specifications, robustness to alternative reliability
links, and likelihood-profile and longer-start checks for the ill-conditioned
Table 5 Set 4 AR(2) extensions.
