---
title: "Estimation revisions on `codex/rulof_new_work`"
subtitle: "Changes relative to the `efficient` branch"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: true
geometry: margin=1in
header-includes:
  - \usepackage{booktabs}
  - \usepackage{amsmath}
  - \usepackage{float}
  - \usepackage{caption}
  - \usepackage{needspace}
  - \usepackage{graphicx}
  - \usepackage{longtable}
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
project_root <- normalizePath("..", winslash = "/", mustWork = TRUE)

include_tex_table <- function(relative_path, tabular_scale = NULL,
                              table_number = NULL, appendix_label = NULL) {
  path <- file.path(project_root, relative_path)
  if (!file.exists(path)) {
    stop("Missing generated table: ", path,
         ". Run Rscript build_tables.R from the project root first.")
  }
  lines <- readLines(path, warn = FALSE)
  lines <- sub("^\\\\begin\\{table\\}\\[[^]]+\\]$", "\\\\begin{table}[H]", lines)
  if (!is.null(table_number)) {
    lines <- c(sprintf("\\setcounter{table}{%d}", as.integer(table_number) - 1L),
               lines)
  }
  if (!is.null(appendix_label)) {
    caption_line <- grep("^\\\\caption\\{", lines)
    if (length(caption_line) != 1L)
      stop("Expected exactly one caption in ", path)
    caption_text <- sub("^\\\\caption\\{(.*)\\}$", "\\1", lines[caption_line])
    lines[caption_line] <- sprintf("\\caption*{%s: %s}", appendix_label,
                                   caption_text)
  }
  if (!is.null(tabular_scale)) {
    begin_tabular <- grep("^\\\\begin\\{tabular\\}", lines)
    end_tabular <- grep("^\\\\end\\{tabular\\}", lines)
    if (length(begin_tabular) != 1L || length(end_tabular) != 1L)
      stop("Expected exactly one tabular environment in ", path)
    lines <- append(lines, "}", after = end_tabular)
    lines <- append(lines, sprintf("\\scalebox{%s}{%%", tabular_scale),
                    after = begin_tabular - 1L)
  }
  cat(lines, sep = "\n")
}

# The paper-facing sections and tables have one authoritative source. The long
# summary renders the short summary body verbatim, then adds technical notes.
short_path <- file.path(dirname(knitr::current_input(dir = TRUE)),
                        "branch_changes_summary_short.Rmd")
short_lines <- readLines(short_path, warn = FALSE)
body_start <- grep("^# 1\\. Baseline transition model", short_lines)
if (length(body_start) != 1L)
  stop("Could not locate the numbered body in branch_changes_summary_short.Rmd")
short_body <- short_lines[body_start:length(short_lines)]
```

```{r synchronized-paper-facing-summary, results='asis'}
cat(knitr::knit_child(text = short_body, quiet = TRUE), sep = "\n")
```

\newpage

# Detailed changes to code and estimation

The preceding numbered sections, tables, interpretations, and to-do lists are
identical to the short summary. This section records the additional estimation
and implementation detail needed to reproduce and audit those results.

## Tables 1--2: exact baseline likelihood and matching robustness

The three-wave baseline estimator works with the exact eight possible observed
employment histories. Collapsing to these histories preserves the weighted
observed likelihood while making multiple-start optimization inexpensive. All
probabilities are represented by unconstrained logits. Stationary models impose
the restriction inside the likelihood; free-initial-condition models estimate
the initial employment probability separately.

The pipeline uses nested warm starts and dispersed deterministic starts, checks
the analytical score against numerical derivatives, requires positive-definite
observed information, and verifies that each unrestricted specification weakly
dominates its nested restriction. Saved fits carry the panel identity, sample
signature, sample size, total weight, likelihood, and optimizer diagnostics so
the table builder can reject stale results.

The main table now reports only free-initial-condition models. Appendix Table
A1 retains the stationary specifications, raw parameters, implied entry, exit,
false-positive and false-negative rates, and tests the stationarity restriction.
Because the fitted pseudo-likelihood uses population-scale survey weights, all
LR statistics normalize weights to sum to the estimation sample size. The test
of zero misclassification uses the 50:50 point-mass/$\chi^2_1$ boundary
reference; asymmetry and stationarity use the regular $\chi^2_1$ reference.

Table 2 reruns the nonstationary no-error and symmetric-error models on panels
A, B, and C. These panels apply progressively stricter longitudinal matching
rules. The sensitivity script changes the panel input explicitly rather than
silently modifying the default ingestion path, which remains
`df_qlfs_A.rds`.

Appendix Tables A2a--A2d diagnose the declining transition estimates without
altering Table 2 or its existing appendix. A2a reports weighted observable
composition. A2b fits Set 2 transition controls separately in panels A--C and
evaluates each model on both its own and panel C's covariate distribution. A2c
decomposes each A--C and B--C predictive-margin gap into the component due to
age, education, race, and sex composition and a residual component. A2d pools
panel A and adds the matching-group indicators sequentially to the transition
equations, the symmetric-error equation, and both. All fits use the exact
Table 2 ingestion samples, free initial employment, multiple starts, full-rank
information checks, and analytical sandwich/delta inference. Predictive margins
in these diagnostic appendices average conditional hazards over a common
covariate population; they are not the posterior risk-set means in Table 2.

## Table 3: observed indicators of linked-record reliability

The inconsistency likelihood permits the symmetric status-error probability to
depend on inconsistencies in reported age, education, race, and gender. Race
and gender disagreements are attributed to waves using the same adjacent-wave
logic as matching algorithm B in Table 2. Expanded models add the magnitude of
age and education discrepancies and indicators for observations retained in
panel B but not C and in panel A but not B.

Two further specifications combine the indicators-only error equation with Set
2 transition covariates (age, age squared, education, race, and gender). The
preferred specification also adds the four origin-wave inconsistency indicators
to both entry and persistence. Table 3 now retains only the
inconsistency-indicator, severity plus matching-rule, and transition-covariate
specifications; the severity-only and Set 2-only results remain in the
appendices.

The preferred model is estimated both with a free common initial-employment
probability and under conditional stationarity. Conditional stationarity sets
each person's wave-1 initial probability to the stationary probability implied
by their wave-1 covariate-specific entry and exit hazards. It does not require
the later hazards to remain constant when origin-wave inconsistencies change.
Because this restriction also introduces observed heterogeneity into the
initial state, its likelihood is not nested in the free-common-alpha model.

The direct likelihood maximization uses multiple starts and explicit score,
curvature, rank, and nesting checks. Table 3 reports only the unrestricted
initial-condition versions. Appendix Table A3 reports the stationary and free
initial-condition results together, including all consistently calculated
implied rates and the raw transition and error coefficients for the
transition-covariate specification.
The mechanism language remains deliberately neutral: inconsistencies may
reflect response problems, incorrect longitudinal matches, or both.

## Table 4: observable transition heterogeneity

Table 4 now reports two nested sets. Set 1 contains age, age squared, education,
race, and sex (the old Set 2). Set 2 adds origin-wave log tenure to persistence,
origin-wave log time since work plus never-worked status to entry, and
origin-wave contract type and informal-sector status to persistence (the old
Set 4). The old Sets 1 and 3 are omitted. Contract type and sector may change
between transitions. They do not enter entry because they are undefined for a
genuinely nonemployed origin state, so respondents are retained without
assigning artificial job characteristics.

The four reported constant-error/no-error specifications retain the repaired
generalized-EM implementation: normalized weights, joint transition-block
updates, likelihood-preserving backtracking, nested starts, and explicit
convergence and nesting checks. Posterior risk-set and survey-weighted hazards
replace unweighted averages. Individual-level survey-weighted sandwich
covariances and delta-method transformations provide approximate analytical
errors where bootstrap estimates are unavailable.

Column (5) is estimated by the exact observed likelihood rather than by the old
separable M-step. It combines the new Set 2 transition equations with a symmetric
error equation containing exactly the same covariates as Table 3 column (1):
age, education, race, and gender inconsistency indicators plus mutually
exclusive indicators for exactly two, three, or four inconsistencies. Zero or
one inconsistency is the omitted count category. Two substantively distinct
starts converged to the same likelihood region, after which Hessian-guided
polishing reduced the maximum normalized score to $8.7\times10^{-9}$. The
polished solution has 31 of 31 information-matrix directions identified,
condition number about 61,200, and maximum Newton correction 0.0021. The mean
entry, exit, and error rates are 4.67%, 4.62%, and 2.55%. Predicted error is
1.28% at the minimum and median and 43.78% at the observed maximum. Detailed
error slopes and the predicted-error distribution appear in Appendix Table A4B.

## Table 5: four-wave AR(1) and AR(2)

The four-wave modules enumerate latent histories and evaluate the exact
collapsed observed likelihood. The AR(1) benchmark has free initial employment
and symmetric error. The unrestricted AR(2) allows the previous two latent
states to determine the next transition. Its duration-controlled specification
adds the same state-appropriate tenure and time-since-work variables used in
Table 4 and permits those variables to change with the origin wave.

The analytical covariance uses the observed information and collapsed
person-level score outer products. The duration-controlled AR(2) information
matrix is locally full rank, although its high condition number means the
initial-state and missing-duration directions warrant caution.

The new extension is estimated as two exact nested likelihoods on the same
308,240 observations. The first combines AR(2) with the Table 4 Set 2 transition
equations and constant symmetric error. The second makes symmetric error depend
on the same predictors as Table 3 column (1): age, education, race, and gender
inconsistency plus mutually exclusive indicators for exactly two, three, or
four inconsistencies. Each stage uses two substantively distinct starts, checks
likelihood nesting, applies an observed-information Newton correction where
useful, and is accepted only with a maximum normalized score below
$2\times10^{-5}$ and full information rank. Stage-specific collapsing leaves
197,933 and 223,907 exact cells. The final scores are $2.8\times10^{-6}$ and
$3.4\times10^{-6}$; both models are locally full rank, with information
condition numbers of about 746,558 and 7,710,884. Appendix Table A5C is
reserved for the new duration-hazard profiles. The piecewise model replaces
each log-duration slope with five indicators, uses 0--3 months as the omitted
category, retains separate never-worked and missing-duration indicators, and
keeps the same Set 2 sample and constant-error equation. Two starts converge to
nearby likelihoods before a Newton correction; the accepted fit has score
$2.6\times10^{-6}$, full information rank (37/37), and condition number about
1,387,680. Because the piecewise and log-linear specifications are non-nested,
their likelihood difference is descriptive rather than an LR statistic.
Table 8 is generated from a common decomposition routine spanning
the requested Tables 1, 4, 5, and 6 models. These comparisons condition only on
the two reports forming an apparent transition and integrate out the subsequent
report. The Table 7 row is explicitly distinguished: because duration reports
are identifying measurement equations in that model, it uses the fitted
full-record posterior including tenure and timegap evidence.

## Table 6: finite-mixture heterogeneity

The two-type four-wave model uses exact likelihood refinement after EM warm
starts, label normalization, multiple starts, and analytical sandwich/delta
inference. The comparison now contains the no-error FMM, the uncontrolled
symmetric-error FMM, the controlled common-error model, and two controlled
type-specific-error models. Observed transition slopes are common across latent
types in the controlled models, while type-specific intercepts retain mobility
heterogeneity.

The first new model uses only type-specific misclassification intercepts. The
second adds the complete Table 3 column (1) predictor set---age, education,
race, and gender inconsistencies plus indicators for exactly two, three, or
four inconsistencies---with common slopes across types. This keeps the
stable/mobile error contrast separate from observed reliability predictors
without assigning either to a particular behavioral mechanism. Both local
information matrices are full rank and have positive curvature. The richer
model estimates posterior mean misclassification rates of 1.34% and 5.33% in
the stable and mobile classes, respectively, and reduces the mobile-class entry
and exit rates relative to the intercept-only model. The failed three-wave
symmetric-error FMM is not reported as an identified result: three waves do not
separate two-type transition heterogeneity from symmetric misclassification.

## Table 7: endogenous duration dependence

The tenure-contamination module preserves the central consistency restriction:
latent employment paths must agree with clean reported tenure and time-since-
work clocks. Successive specifications relax constant transition hazards,
integrate unavailable wrong-state clocks, use tail-safe piecewise hazards, and
allow alternative contamination processes for tenure and timegap reports.
These are likelihood changes, not presentation-only adjustments.

The preferred timegap treatment is the joint per-wave marginal-contamination
model. The table reports approximate analytical errors from the full numerical
observed-information matrix and delta method, not bootstrap errors. Appendix
Table A7 retains hazard profiles, auxiliary comparisons, wrong-state-clock
integration checks, and the within-quarter reset decomposition.

## Table 8: apparent-transition implications

Table 8 uses a common decomposition routine to compare how the principal
models classify apparent transitions as classification only, genuine reversal,
or genuine persistence. It is presented after Table 7 because it synthesizes
results across Tables 1 and 4--7 rather than documenting only the AR(2) model.

## Appendix Table A8: no-error duration counterfactual

The counterfactual diagnostic estimates flexible no-status-error AR(1) and
AR(2) transition models without using destination-wave durations in the
likelihood. It then propagates implied spell clocks and compares them with the
held-out reported tenure and timegap. Large conditional contradictions show
that reported employment changes and duration histories cannot generally both
be correct. The exercise motivates joint modeling but does not distinguish
response inconsistency from longitudinal matching error.

## Reproducibility and generated files

`build_tables.R` generates Tables 1, 3, and 4 and their coefficient appendices;
the matching-sensitivity runner generates Table 2, and
`EM-baseline/estimate_matching_composition.R` generates supplementary Tables
A2a--A2d without overwriting Table 2 or Appendix Table A2. The four-wave and duration
modules retain their own reproducible runners, tests, result summaries, and
diagnostic outputs. Confidential data and intermediate fitted objects remain
ignored by Git. Curated paper-facing snapshots and both summary sources are
tracked. A complete rerun still requires the confidential QLFS inputs and the R
environment recorded by `renv`.

The remaining inference priorities are the person-level bootstrap for the main
covariate and duration specifications, robustness to alternative reliability
links, and likelihood-profile and longer-start checks for the ill-conditioned
Table 5 Set 2 AR(2) extensions.
