---
title: "Estimation revisions on `codex/rulof_new_work`"
subtitle: "Changes relative to the `efficient` branch"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: true
geometry: margin=1in
header-includes:
  - \usepackage{booktabs}
  - \usepackage{amsmath}
  - \usepackage{float}
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
project_root <- normalizePath("..", winslash = "/", mustWork = TRUE)

include_tex_table <- function(relative_path, tabular_scale = NULL) {
  path <- file.path(project_root, relative_path)
  if (!file.exists(path)) {
    stop("Missing generated table: ", path,
         ". Run Rscript build_tables.R from the project root first.")
  }
  lines <- readLines(path, warn = FALSE)
  if (!is.null(tabular_scale)) {
    begin_tabular <- grep("^\\\\begin\\{tabular\\}", lines)
    end_tabular <- grep("^\\\\end\\{tabular\\}", lines)
    if (length(begin_tabular) != 1L || length(end_tabular) != 1L)
      stop("Expected exactly one tabular environment in ", path)
    lines <- append(lines, "}", after = end_tabular)
    lines <- append(
      lines,
      sprintf("\\scalebox{%s}{%%", tabular_scale),
      after = begin_tabular - 1L
    )
  }
  cat(lines, sep = "\n")
}
```

# Purpose and scope

This note records the changes made on `codex/rulof_new_work` relative to the
`efficient` branch. It covers the exact baseline estimator and analytical
inference used in Tables 2 and 3, together with the observable-covariate
extension and revised Table 4. It also records the repaired four-wave AR(1)
models, the two-type finite-mixture extensions, and the models combining
observable transition heterogeneity with inconsistency-dependent
misclassification. The AR(2) and tenure specifications are not reinterpreted
here.

The central objectives were to replace stale or numerically unreliable results
with validated maximum-likelihood estimates, provide fast analytical inference,
handle state-specific duration and job characteristics without discarding
respondents, and report heterogeneous transition risks in a
way that remains comparable with the paper's earlier tables. The later
four-wave work tests whether latent classes partly absorb observed heterogeneity
or heterogeneous measurement reliability.

# Changes to the code and estimation approach

## Exact baseline estimator and result provenance

The simple model is now maximised directly on the eight possible observed
three-wave employment histories. This collapse is exactly equivalent to the
row-level likelihood and makes the optimisation extremely fast. It also fixes
the stationary model: the earlier closed-form transition update imposed the
stationary initial probability only afterwards and therefore did not maximise
the exact constrained likelihood.

All probabilities are mapped from unconstrained logit parameters. Each model
uses multiple starts, with symmetric models initialized from the corresponding
no-error estimates and asymmetric models from the symmetric estimates. Results
are saved only when the likelihood score is small, observed information is
positive definite, and every unrestricted model has a likelihood at least as
large as its nested restriction. Each saved fit also records its source panel,
sample signature, sample size, weight total, and optimiser diagnostics. The
table builder rejects stale fits or inference files that do not match the
current `df_qlfs_A.rds` estimation sample.

## Numerical optimisation

The covariate model is estimated by a Generalised EM algorithm. The earlier
M-step used separate probit IRLS updates. That approach did not maximise the
correct joint expected complete-data likelihood when the initial-state term
coupled the entry and persistence equations, and it could return a solution
whose likelihood was below that of the nested no-misclassification model.

The revised estimator:

1. normalises survey weights during optimisation to improve numerical
   conditioning, while restoring the original likelihood scale in saved output;
2. updates the entry and persistence coefficient blocks jointly with limited
   BFGS steps;
3. backtracks until both the complete-data objective and observed-data
   likelihood are non-decreasing;
4. uses several starting values and initializes each symmetric-error model from
   its corresponding no-error estimate;
5. explicitly checks that the symmetric model's maximised likelihood is no
   lower than that of its nested no-error model; and
6. requires both a small likelihood improvement and a small parameter change
   before declaring convergence.

These safeguards address the invalid-MLE problem directly: they alter how the
model is estimated, rather than merely changing how its results are displayed.

## Covariate sets and state-specific variables

The revised sequence of specifications is:

1. Set 1: age, age squared, and education;
2. Set 2: Set 1 plus race and sex;
3. Set 3: Set 2 plus log tenure in the employment-persistence equation, and
   log time since last employment plus a never-worked indicator in the entry
   equation; and
4. Set 4: Set 3 plus contract type and informal-sector status in the
   persistence equation.

Tenure and time since work are measured at the origin wave and transformed as
$\log(1+d)$. Their coefficients are restricted to the equation in which the
reported duration is economically defined. The ingestion script preserves
these state-appropriate reported values before the separate tenure module's
nearest-spell filling. A missing-tenure indicator prevents missing reports from
being treated as genuine zero tenure; zero-variance missingness indicators are
automatically omitted to avoid unidentified information-matrix directions.

## Contract type and informal-sector status

Contract type is structurally undefined when a respondent is not employed.
The adopted treatment is therefore to include it only in the
employment-persistence equation. Its coefficient in the job-entry equation is
fixed at zero, so non-employed observations are retained instead of being
dropped or assigned an economically artificial contract category.

Set 4 also includes an informal-sector indicator, defined from the QLFS
sector classification and including informal agricultural employment. Like
contract type, it enters only the persistence/exit equation. Rural or urban
residence is not included, avoiding the sample loss caused by missing residence
information.

Both employment characteristics are measured at the origin of each transition:
wave 1 values are used for the wave 1--2 transition and wave 2 values for the
wave 2--3 transition. They can consequently change between waves. Employment
is still represented by a single latent state; formal and informal employment
are not modelled as separate states.

## Initial conditions and reported transition rates

Because the Set 3 and Set 4 transition covariates can vary between the two transition
periods, the main comparison no longer imposes a time-homogeneous stationary
initial distribution. All eight columns of the revised Table 4 estimate the
initial employment probability freely.

Average entry and exit rates are computed using the survey weights and the
posterior probability that each observation belongs to the appropriate latent
origin-state risk set. This makes the averages conditional hazards: entry is
averaged over latent non-employment exposure and exit over latent employment
exposure. The table also reports the posterior employed risk share and total
employment churn. The separate entry-flow and exit-flow rows have been removed
to keep the presentation close to the earlier implied-probability tables.

## Analytical standard errors

Pending a full bootstrap, inference for the covariate models uses an
individual-level survey-weighted sandwich estimator. The observed-information
``bread'' is obtained by numerically differentiating the analytic observed-data
score, while the ``meat'' is the outer product of weighted individual scores.
The delta method propagates this covariance matrix to average transition rates,
total churn, and Set 4 discrete effects. The linearisation also includes direct
sampling variation from averaging predicted quantities over the observed
covariate and posterior-risk distribution.

These are approximate analytical standard errors. They use survey weights but
do not incorporate stratification or PSU clustering because those identifiers
are unavailable in the estimation extract. A person-level bootstrap remains a
useful robustness exercise.

## Output and reproducibility

The extension pipeline saves fitted model objects and convergence diagnostics.
`EM-baseline-ext/estimate_analytical_se_table4.R` produces the analytical
covariance estimates, and `build_tables.R` creates the LaTeX results below.
Generated model objects and tables remain excluded from version control and can
be rebuilt locally from the underlying confidential QLFS files.

# New and revised tables

The following are the baseline tables rebuilt with the exact estimator and
analytical inference, followed by the two covariate tables introduced on this
branch. Unchanged tables are not duplicated.

## Table 2: baseline implied probabilities

```{r baseline-implied, results='asis'}
include_tex_table("EM-baseline/output/tables/table_baseline_implied.tex")
```

Without a reporting-error correction, quarterly entry and exit rates are about
8.2--8.6%. Allowing symmetric misclassification lowers both rates to roughly
3%, while the estimated per-wave misclassification probability is itself only
2.99%. The result is robust to freeing the initial employment probability and
to allowing asymmetric reporting errors.

## Table 3: baseline parameters

```{r baseline-parameters, results='asis'}
include_tex_table("EM-baseline/output/tables/table_baseline_params.tex")
```

The asymmetric estimates now satisfy the required likelihood nesting
relationship. False-positive and false-negative probabilities are approximately
3.05% and 2.93%; their small difference has virtually no effect on the implied
transition rates. The table distinguishes the initial employment probability
$\alpha$ from the steady-state employment share $\alpha^*$ reported in Table 2.

## Revised Table 4

```{r revised-table-4, results='asis'}
include_tex_table("EM-baseline-ext/output/tables/table_cov_risk_weighted.tex")
```

The no-error specifications imply quarterly entry and exit rates of about
8.14% and 8.58%, almost invariant to the four covariate sets. Allowing for
symmetric reporting error lowers these rates to 3.62--4.87% and 3.71--5.09%,
respectively. The estimated per-wave misclassification probability declines
from 2.58% in Set 1 to 2.18% in Set 4. Observable characteristics therefore explain
substantial heterogeneity across workers, but they do not explain away the
measurement-error signal. Ignoring misclassification still approximately
doubles the estimated average transition hazards and total churn.

The exit rate is only modestly higher than the entry rate in the symmetric
models. In Set 4 the difference is about 0.22 percentage points, compared with
roughly 0.44 percentage points in the no-error specification. This is more
consistent with the approximately stable employment share than a comparison of
the two conditional hazards alone might initially suggest.

## Coefficients, hazard distributions, and Set 4 discrete effects

```{r coefficient-appendix, results='asis'}
include_tex_table(
  "EM-baseline-ext/output/tables/table_cov_coefficients_appendix.tex",
  tabular_scale = 0.55
)
```

The appendix demonstrates why a single mean hazard is incomplete. In symmetric
Set 4, the risk-weighted 10th--90th percentile range is approximately
1.00--13.82% for entry and 0.19--14.29% for exit. Thus workers with different
observed characteristics face markedly different transition risks even though
the population-average entry and exit rates are similar.

The duration coefficients have the expected signs. In symmetric Set 4, a
one-standard-deviation increase in log tenure raises the persistence probit
index by 0.508 (SE 0.012), while a one-standard-deviation increase in log time
since work lowers the entry index by 0.307 (SE 0.006). Never having worked also
substantially lowers entry (coefficient -0.835, SE 0.021). The missing-tenure
coefficient is imprecise and should not be interpreted substantively.

Holding other covariates at their observed values, a permanent contract lowers
the average quarterly exit probability by about 2.43 percentage points in the
symmetric-error model, while informal-sector employment raises it by about
2.36 percentage points. The corresponding no-error effects are -3.28 and +3.01
percentage points. Adding duration variables therefore exposes much wider
heterogeneity in predicted hazards while preserving the main conclusion that
misclassification materially inflates uncorrected transition rates.

# Four-wave finite-mixture results

The new `EM-AR1-4W` module uses every complete four-wave history in the
method-A panel. The controlled specifications contain 308,240 people, exactly
collapsed to 105,257 likelihood cells. Initial employment probabilities are
free rather than restricted to stationarity. Estimation combines an EM warm
start with exact observed-likelihood analytical scores, dispersed starting
values, label normalisation, and survey-weighted sandwich/delta-method
inference.

## Uncontrolled and controlled two-type models

The uncontrolled model estimates constant transition and symmetric
misclassification probabilities within two time-invariant latent classes. The
controlled model instead gives the classes distinct transition intercepts but
constrains all transition-covariate slopes to be common. Baseline age,
age-squared, education, race, and sex enter both transition equations.
Origin-wave permanent-contract and informal-sector status can change over time
and enter persistence only. Age and education inconsistencies predict
misclassification without attempting to distinguish inconsistent respondent
reports from incorrect panel matches.

| Model and latent type | Population share | Initial employment | Risk-weighted entry | Risk-weighted exit |
|:--|--:|--:|--:|--:|
| Uncontrolled: stable | 92.13 | 47.74 | 2.36 | 2.21 |
| Uncontrolled: mobile | 7.87 | 47.95 | 42.16 | 47.04 |
| Controlled, common misclassification: stable | 77.54 | 45.19 | 1.12 | 0.59 |
| Controlled, common misclassification: mobile | 22.46 | 56.27 | 24.54 | 19.60 |
| Controlled, type-specific misclassification: stable | 72.92 | 44.49 | 0.84 | 0.26 |
| Controlled, type-specific misclassification: mobile | 27.08 | 56.58 | 21.28 | 16.72 |

All entries are percentages. In the preferred type-specific-misclassification
model, analytical standard errors are 1.23 percentage points for the class
shares, 0.15 and 0.10 points for stable entry and exit, and 0.67 and 0.65 points
for mobile entry and exit. The class distinction remains economically large,
but observable characteristics and heterogeneous reliability make the mobile
class larger and substantially less extreme than in the uncontrolled model.
This shows that the original latent classes combined residual mobility,
observed heterogeneity, and heterogeneous misclassification.

The common-slope estimates are economically coherent. Permanent contracts
raise latent employment persistence (probit coefficient 0.584, SE 0.022),
whereas informal-sector employment lowers it (-0.237, SE 0.021). Age effects
are nonlinear: entry peaks at approximately age 38 and persistence at
approximately age 53. Education raises persistence, while women have lower
conditional entry and persistence probabilities. These are conditional
associations rather than causal effects.

## Type-specific misclassification

The parsimonious relaxation gives the stable and mobile classes different
misclassification intercepts while keeping the age- and education-inconsistency
slopes common.

| Response pattern | Stable type | Mobile type | Mobile minus stable |
|:--|--:|--:|--:|
| Consistent age and education | 0.46 (0.03) | 1.72 (0.10) | 1.26 |
| Age inconsistency | 10.31 (0.38) | 24.95 (1.02) | 14.64 |
| Education inconsistency | 2.29 (0.13) | 7.77 (0.48) | 5.48 |
| Both inconsistencies | 28.65 (0.52) | 41.86 (0.63) | 13.21 |

Probabilities and analytical standard errors are percentages. The mobile
class has about 3.7 times the baseline misclassification probability. The
difference between the two transformed intercepts is 1.344 (SE 0.105), giving
a robust Wald statistic of 12.79. The weighted log likelihood improves from
-435,310.33 to -435,207.29 with one additional parameter. Because the objective
uses survey weights, this likelihood gain should not be assigned the usual
unweighted likelihood-ratio reference distribution; the sandwich Wald result
provides separate evidence against a common intercept.

The estimates do not mean that latent mobility causes misclassification. Class
membership and misclassification are inferred jointly. The higher mobile-type
probability may represent persistent differences in response reliability,
match quality, or other omitted characteristics.

## Observable characteristics and posterior class membership

The mixture share is constant *a priori*, but posterior class probabilities can
still be associated with observed characteristics through the transition and
measurement likelihoods. Simple survey-weighted correlations between the
posterior mobile probability and age, education, sex, and individual race
indicators are small in absolute value (approximately 0.01--0.05). Mean mobile
probability rises from 24.3% at ages 18--24 to 28.1% at ages 45--55, is 27.7%
for men and 26.5% for women, and ranges only from 26.7% to 29.1% across race
groups.

The stronger simple correlations are with permanent-contract exposure (0.131),
informal-sector exposure (0.084), age inconsistencies (0.061), and education
inconsistencies (0.073). Conditional on observed employment, however, average
mobile probabilities differ only modestly by permanent contract (31.3% versus
32.2% without one) and by informal status (32.9% versus 31.3% outside the
informal sector).

Observable blocks explain little of the posterior variation. A survey-weighted
linear projection on baseline demographics has an $R^2$ of 0.8%; adding the
inconsistency measures raises it to 1.5%. Among people observed employed at
least once in the first three waves, demographics and job characteristics
explain 3.0%, rising to 3.8% after adding inconsistencies. The classes therefore
are not merely demographic or sectoral labels; most posterior separation still
comes from employment histories. These regressions are descriptive and should
not be interpreted as a model of, or causal effects on, class membership.

## Four-wave diagnostics and reproducibility

The preferred model has full information-matrix rank (27/27), a positive
minimum eigenvalue of 0.000275, and condition number 913. All four refined
starts agree within 0.001 log-likelihood units, and the maximum weighted score
is $3.0\times10^{-6}$. The four-wave test suite contains 37 passing checks,
including posterior normalisation and numerical verification of the analytical
scores for both common and type-specific misclassification.

The main scripts are `EM-AR1-4W/estimate_fmm_covariates_inconsistency.R` and
`EM-AR1-4W/estimate_fmm_type_specific_misclassification.R`. Generated fitted
objects remain ignored by Git and can be reconstructed from the confidential
QLFS inputs.

# Interpretation for the paper

The revisions strengthen the paper's core result. The earlier numerical problem
could have made the misclassification specification appear to fit worse than a
model it mathematically nests. Once the likelihood is optimised with appropriate
monotonicity and nesting checks, a small but precisely estimated reporting-error
rate continues to produce a large reduction in inferred employment turnover.

At the same time, the new Sets 3--4 results sharpen the economic interpretation.
Misclassification and genuine heterogeneity are not competing explanations in
which only one survives: duration dependence, contract security, and informal
employment strongly predict true transition risk, while a separate
reporting-error component remains.
The paper should therefore distinguish clearly between variation in transition
risk across workers and the upward bias in observed transitions caused by
classification error.

# Remaining checks

Before final submission, the analytical standard errors should ideally be
compared with a person-level bootstrap. If the survey extract can be augmented
with strata and PSU identifiers, design-based clustered inference would also be
preferable. The main qualitative conclusions should be assessed against those
two alternatives, while retaining the current analytical estimates as a fast
and reproducible diagnostic.
