---
title: "Four-wave duration model: paper table"
output:
  pdf_document:
    latex_engine: xelatex
geometry: margin=1in
header-includes:
  - \usepackage{booktabs}
  - \usepackage{float}
  - \usepackage{caption}
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
project_root <- normalizePath("..",winslash="/",mustWork=TRUE)
```

The table compares the repaired four-wave AR(1) model with its two-parameter
AR(2) extension on Panel A. Duration-reporting assumptions are unchanged.
Table 8 also includes these models' apparent-transition implications.

```{r table, results='asis'}
path <- file.path(project_root,
  "paper/generated/four_wave_duration/table_four_wave_duration.tex")
if (!file.exists(path)) stop("Run scripts/build_four_wave_duration_paper_table.R first")
lines <- readLines(path,warn=FALSE)
lines <- lines[!grepl("^%",lines)]
lines <- gsub("[htbp]","[H]",lines,fixed=TRUE)
cat("\\setcounter{table}{8}\n",lines,sep="\n")
```

The AR(2) model attributes more turnover to genuine short-lived transitions,
while retaining a substantial role for status misclassification. Recent
entrants and leavers have much higher next-quarter transition probabilities
than people whose status was unchanged in the preceding interval. These
comparisons combine duration composition with the estimated history effects;
they are not causal effects of switching.

The questionnaire distinction is important: timegap measures time since last
work, while tenure is derived from the reported employer/business start date.
The current findings remain conditional on the treatment of new nonemployment
spells and interruptions followed by return to the same employer. Those
sensitivity checks and analytical uncertainty are still pending.

Rebuild this table, without estimating any model, from the repository root:

```sh
Rscript scripts/build_four_wave_duration_paper_table.R
```
