---
title: "Short summary of estimation revisions"
subtitle: "Changes on codex/rulof_new_work relative to efficient"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: true
geometry: margin=1in
header-includes:
  - \usepackage{booktabs}
  - \usepackage{amsmath}
  - \usepackage{float}
  - \usepackage{caption}
  - \usepackage{needspace}
  - \usepackage{graphicx}
  - \usepackage{longtable}
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
project_root <- normalizePath("..", winslash = "/", mustWork = TRUE)

include_tex_table <- function(relative_path, tabular_scale = NULL,
                              table_number = NULL, appendix_label = NULL) {
  path <- file.path(project_root, relative_path)
  if (!file.exists(path)) {
    stop("Missing generated table: ", path,
         ". Run Rscript build_tables.R from the project root first.")
  }
  lines <- readLines(path, warn = FALSE)
  lines <- sub("^\\\\begin\\{table\\}\\[[^]]+\\]$", "\\\\begin{table}[H]", lines)
  if (!is.null(table_number)) {
    lines <- c(sprintf("\\setcounter{table}{%d}", as.integer(table_number) - 1L),
               lines)
  }
  if (!is.null(appendix_label)) {
    caption_line <- grep("^\\\\caption\\{", lines)
    if (length(caption_line) != 1L)
      stop("Expected exactly one caption in ", path)
    caption_text <- sub("^\\\\caption\\{(.*)\\}$", "\\1", lines[caption_line])
    lines[caption_line] <- sprintf("\\caption*{%s: %s}", appendix_label,
                                   caption_text)
  }
  if (!is.null(tabular_scale)) {
    begin_tabular <- grep("^\\\\begin\\{tabular\\}", lines)
    end_tabular <- grep("^\\\\end\\{tabular\\}", lines)
    if (length(begin_tabular) != 1L || length(end_tabular) != 1L)
      stop("Expected exactly one tabular environment in ", path)
    lines <- append(lines, "}", after = end_tabular)
    lines <- append(lines, sprintf("\\scalebox{%s}{%%", tabular_scale),
                    after = begin_tabular - 1L)
  }
  cat(lines, sep = "\n")
}
```

# 1. Baseline transition model: Table 1 and Appendix Table A1

- Re-estimated the three-wave model using the exact eight-history likelihood.
- Repaired stationarity; used 12 reproducible starts (including nested-model
  starts); and required optimizer convergence, a small score, positive observed
  information, and valid likelihood nesting before saving results.
- Replaced stale inference with analytical survey-weighted sandwich and delta
  standard errors.
- Separated the free initial employment probability from the model-implied
  long-run employment share.
- Restricted the main table to free-initial-condition specifications and moved
  the stationary counterparts, implied rates, and stationarity tests to A1.
- Added normalized-weight LR tests of non-zero misclassification, asymmetric
  error, and non-stationarity, using the boundary reference where required.

```{r baseline-implied, results='asis'}
include_tex_table("EM-baseline/output/tables/table_baseline_implied.tex",
                  table_number = 1)
```

```{r baseline-parameters, results='asis'}
include_tex_table("EM-baseline/output/tables/table_baseline_params.tex",
                  tabular_scale = 0.82, appendix_label = "Appendix Table A1")
```

## Interpretation

- The no-error model implies quarterly entry and exit of about 8.2--8.6%.
- Allowing symmetric error lowers both rates to about 3%, with a per-wave
  misclassification probability of 2.99%.
- The no-error restriction is decisively rejected (LR 22,438.36, boundary
  $p<0.001$). False-positive and false-negative probabilities remain very
  similar at 3.05% and 2.92%; equality is statistically rejected (LR 7.49,
  $p=0.006$) but the substantive difference is only 0.13 percentage points.
- Appendix Table A1 rejects stationarity under no error, symmetric error, and
  asymmetric error (LR 12.51, 20.77, and 21.93; all $p<0.001$). The restrictions
  nevertheless have only modest effects on the estimated transition rates.
- The repaired likelihood strengthens the paper's central conclusion:
  relatively infrequent classification error can create a large upward bias in
  measured labour-market turnover.

## To do

- Check whether transition probabilities vary across wave pairs.
- Confirm the normalized-weight LR conclusions with a person-level bootstrap
  or likelihood profiles for the misclassification and transition rates.

# 2. Panel-matching robustness: Table 2

- Applied the same exact, non-stationary estimator, 12-start search, and
  convergence and curvature checks to panels A, B, and C.
- Held all sample filters fixed apart from progressively stricter matching.
- Added consistent analytical standard errors to every panel comparison.

```{r matching-sensitivity, results='asis'}
include_tex_table("EM-baseline/output/tables/table_matching_implied.tex",
                  tabular_scale = 0.82, table_number = 2)
```

```{r matching-composition, results='asis'}
include_tex_table("EM-baseline/output/tables/table_matching_composition.tex",
                  appendix_label = "Appendix Table A2a")
```

```{r matching-standardized, results='asis'}
include_tex_table("EM-baseline/output/tables/table_matching_standardized.tex",
                  tabular_scale = 0.72, appendix_label = "Appendix Table A2b")
```

```{r matching-decomposition, results='asis'}
include_tex_table("EM-baseline/output/tables/table_matching_decomposition.tex",
                  tabular_scale = 0.88, appendix_label = "Appendix Table A2c")
```

```{r matching-pooled-groups, results='asis'}
include_tex_table("EM-baseline/output/tables/table_matching_pooled_groups.tex",
                  tabular_scale = 0.80, appendix_label = "Appendix Table A2d")
```

## Interpretation

- Stricter matching lowers both observed and corrected transition rates.
- The estimated misclassification probability declines from 2.99% in panel A
  to 2.15% in panel C.
- Corrected entry and exit remain much lower than no-error rates under every
  matching rule.
- Matching quality is therefore part of the inconsistency mechanism, but the
  main correction is not driven solely by the baseline matching rule.
- Observable composition changes very little: from panel A to C, mean age rises
  by 0.38 years, mean education falls by 0.08 years, and the female share rises
  by 0.88 percentage points.
- Standardizing all models to panel C's age, education, race, and sex
  distribution does not remove the transition gradient. In the symmetric-error
  model it changes panel A entry from 4.63% to 4.66% and exit from 5.92% to
  5.83%. Observable composition explains about $-4.7$% of the A--C entry gap
  and 6.3% of the exit gap; the corresponding B--C shares are $-5.7$% and 9.3%.
- In the pooled joint model, holding Set 2 covariates fixed, entry, exit, and
  misclassification are 3.98%, 4.27%, and 1.70% for records retained in C;
  5.64%, 7.93%, and 3.39% for B-but-not-C records; and 8.41%, 15.12%, and 5.82%
  for A-but-not-B records. The stricter-match gradient therefore reflects both
  higher underlying mobility and higher classification error among excluded
  records, not ordinary demographic composition alone.
- The pooled estimates are descriptive: matching-group membership is selected
  by the linkage rules, so the coefficients should not be read as causal effects
  of applying a stricter algorithm.

## To do

- Repeat the standardization with richer duration and employment covariates and
  test sensitivity to alternative predictive-margin target populations.
- Profile the pooled transition/error allocation and use longer-start checks;
  the data clearly reject equal groups, but either channel alone explains much
  of the raw gradient.
- Check whether the Table 4 and age/education-consistency results are also
  stable under the stricter matching rules.

# 3. Reported-characteristic inconsistencies: Table 3 and Appendix Table A3

- Re-estimated the inconsistency-indicator, severity, and matching-rule
  models after adding race and gender disagreement indicators to every error
  equation.
- Added mutually exclusive indicators for exactly two, three, or four reported-
  characteristic inconsistencies to each misclassification equation, with zero
  or one inconsistency as the omitted category.
- Retained unrestricted-initial-employment estimates in Table 3 and report the
  stationary counterparts alongside them in A3.
- Repaired the EM update, used multiple starts and likelihood checks, and added
  analytical standard errors.
- Interpreted age and education inconsistencies neutrally: they may arise from
  response error, matching error, or both.
- Added robustness allowing true transition rates to differ between records
  with and without an inconsistency and allowing misclassification to depend on
  inconsistency severity.
- Added stationary and free-initial-employment specifications in which the
  misclassification rate also depends on person-wave indicators for membership
  in matching panel B but not C and panel A but not B, defined using the Table 2
  sample-construction rules.
- Reduced the main table to the inconsistency-indicator, severity plus
  matching-rule, and transition-covariate specifications. A3 now contains only
  the stationary/free pairs for these same three specifications; the
  severity-only and Set 2-only results remain in separate appendix outputs.
- Added a conditional-stationarity counterpart to the preferred model. It makes
  wave-1 initial employment depend on the wave-1 covariate-specific entry and
  exit hazards, while retaining the wave-varying transition equations.
- Simplified A3 by replacing pattern-specific probabilities and probability
  increments with the lowest, median, mean, and highest predicted
  misclassification probabilities, with analytical standard errors. It also
  reports the preferred-model transition and error coefficients under
  conditional stationarity and free initial employment.

```{r inconsistency-main, results='asis'}
include_tex_table("EM-baseline-ext/output/tables/table_inconsistency.tex",
                  tabular_scale = 0.78, table_number = 3)
```

```{r inconsistency-details, results='asis'}
include_tex_table("EM-baseline-ext/output/tables/table_inconsistency_details.tex",
                  tabular_scale = 0.50,
                  appendix_label = "Appendix Table A3, Panel A")
```

```{r inconsistency-details-coefficients, results='asis'}
include_tex_table(
  "EM-baseline-ext/output/tables/table_inconsistency_details_coefficients.tex",
  tabular_scale = 0.78,
  appendix_label = "Appendix Table A3, Panel B")
```

```{r inconsistency-details-diagnostics, results='asis'}
include_tex_table(
  "EM-baseline-ext/output/tables/table_inconsistency_details_diagnostics.tex",
  tabular_scale = 0.90,
  appendix_label = "Appendix Table A3, Panel B (continued)")
```

```{r inconsistency-transition-covariates, results='asis'}
include_tex_table("EM-baseline-ext/output/tables/table_inconsistency_transition_covariates.tex",
                  tabular_scale = 0.76, appendix_label = "Appendix Table A3A")
```

The additional reliability-group robustness table is generated as
table_inconsistency_reliability.tex in EM-baseline-ext/output/tables.

## Interpretation

- In the unrestricted indicators-only model, mean misclassification is 3.39%;
  the entry and exit rates are 2.57% and 2.15%, and initial employment is
  47.60%.
- Conditional on the other indicators, a race disagreement barely changes
  predicted error in the first two models. A gender disagreement is much more
  informative: predicted error rises from 2.21% to 8.85% in the indicators-only
  model.
- Stationarity materially changes the entry and exit decomposition but barely
  changes the inconsistency-related misclassification probabilities. In the
  free model, initial employment is 6.56 percentage points below its implied
  steady-state rate.
- Allowing additional large-increment effects substantially sharpens the age
  gradient: predicted misclassification rises from 6.31% for a mild age
  inconsistency to 16.66% for a large one. The corresponding education increase
  is only 0.62 percentage points.
- In the augmented free-initial-employment model, predicted misclassification
  for a record with no reported-characteristic inconsistency is 2.01% when
  retained by both stricter panels and 2.93% in either B-but-not-C or A-but-not-B
  observations. Thus the matching-rule indicators retain a reliability
  gradient after controlling for race and gender disagreements.
- Adding the matching-rule indicators attenuates but does not eliminate the
  age/education gradients: predicted error is 4.79% for a mild and 13.89% for a
  large age inconsistency, and 4.34% and 4.77% for mild and large education
  inconsistencies. Mean error changes little, from 3.40% to 3.41%; entry and
  exit move from 2.57% and 2.14% to 2.54% and 2.10%.
- In the matching-rule-indicator model, a race disagreement predicts 1.70% error and
  a gender disagreement 6.67%, compared with 2.01% when none of the four
  reported characteristics is inconsistent. The race contrast is small and
  negative; the gender contrast remains large and positive.
- The omitted Set 2-only main-table column remains in A3A: its posterior
  risk-set mean entry and exit rates are 3.37% and 3.17%, while mean
  misclassification is 2.93%.
- In the preferred free-initial model, adding the inconsistency indicators to
  both transition equations raises the
  mean entry and exit rates to 3.55% and 3.51% and lowers mean error only
  modestly, to 2.85%. The likelihood gain is large, so the indicators capture
  real transition heterogeneity as well as measurement reliability.
- Under conditional stationarity, the corresponding entry, exit, and mean
  misclassification rates are 3.47% (SE 0.05), 3.78% (0.05), and 2.71% (0.03);
  mean initial employment is 47.93% (0.08). This model is full rank (29/29).
  Its much larger likelihood is not an LR test of stationarity: unlike the free
  model's common initial probability, conditional stationarity lets initial
  employment vary with observed covariates.
- The error gradient remains economically large after this control: predicted
  error is 1.76% with no inconsistency, 12.52% with an age inconsistency,
  5.01% with an education inconsistency, and 7.53% with a gender inconsistency.
- The relationship is not explained solely by different true transition rates
  among records with age or education inconsistencies.
- Larger age inconsistencies are much more informative than mild ones;
  inconsistency severity matters less for education.
- Age, education, and gender inconsistencies predict higher status misclassification,
  without distinguishing response errors from incorrect panel matches.

## To do

- Check alternative nonlinear definitions of inconsistency extent and more
  flexible inconsistency-group transition rates.
- Bootstrap the age/education-inconsistency effects and repeat the analysis
  under stricter panel matching.
- Test alternative timing conventions for putting inconsistency indicators in
  the transition equations and allow the initial employment probability to
  depend on baseline covariates.

# 4. Observable transition heterogeneity: Table 4

- Repaired the covariate-model optimiser so every specification maximises the
  appropriate likelihood, uses five starts including a nested-model warm start,
  and dominates its nested no-error model.
- Added analytical standard errors and posterior risk-set and survey-weighted
  transition rates.
- Retained the old Set 2 as new Set 1 and the old Set 4 as new Set 2; omitted
  the old Sets 1 and 3 to focus the table on the two preferred nested sets.
- New Set 2 adds tenure, time since work, time-varying contract type, and
  informal-sector status without restricting the sample to respondents observed
  in employment at every wave.
- Added a new Set 2 specification in which the symmetric misclassification rate
  uses exactly the Table 3 column (1) covariates: age, education, race, and
  gender inconsistency indicators plus indicators for exactly two, three, or
  four inconsistencies.
- Moved raw coefficients, the initial-state parameter, total churn, and hazard
  distributions to an appendix table.

```{r revised-table-4, results='asis'}
include_tex_table("EM-baseline-ext/output/tables/table_cov_risk_weighted.tex",
                  tabular_scale = 0.60, table_number = 4)
```

```{r covariate-appendix, results='asis'}
include_tex_table(
  "EM-baseline-ext/output/tables/table_cov_coefficients_appendix.tex",
  tabular_scale = 0.55, appendix_label = "Appendix Table A4")
```

```{r covariate-reliability-appendix, results='asis'}
include_tex_table(
  "EM-baseline-ext/output/tables/table_cov_reliability_appendix.tex",
  tabular_scale = 0.82, appendix_label = "Appendix Table A4B")
```

## Interpretation

- No-error entry and exit remain close to 8.1% and 8.6% across covariate sets.
- Symmetric-error estimates range from 3.77--4.87% for entry and 3.89--5.09%
  for exit; estimated misclassification falls from 2.50% in Set 1 to 2.18% in
  Set 2.
- In symmetric Set 2, predicted risks are highly dispersed: the weighted
  10th--90th percentile range is 1.00--13.82% for entry and 0.19--14.29% for
  exit.
- Allowing error rates to depend on linked-record reliability gives entry and
  exit rates of 4.67% and 4.62%, and raises the mean error rate from 2.18% in
  constant-error Set 2 to 2.55%.
- Predicted error has a weighted median of 1.28% and mean of 2.55%, but reaches
  43.78% at the observed maximum. Within inconsistency-count groups, the
  survey-weighted means rise from 1.28% with no inconsistency to 6.79%, 30.68%,
  40.78%, and 43.78% with one through four inconsistencies. Reliability
  heterogeneity is therefore substantial, although corrected entry and exit
  remain close to the constant-error Set 2 values.
- Observable heterogeneity therefore explains some differences in transition
  risk but does not explain away the classification-error signal.
- Duration is economically important: entry falls with time since work and exit
  falls with tenure.
- Permanent contracts reduce exit and informal employment raises it. These are
  conditional associations, not causal estimates.
- The paper should distinguish heterogeneous true transition risk from the
  upward bias in observed transitions caused by classification error.

## To do

- Allow the initial employment probability to depend on baseline covariates and
  check whether this changes the transition and error estimates.
- Test more flexible duration functions and confirm the main inference with a
  person-level bootstrap.
- Check robustness to alternative definitions and interactions for the
  reliability indicators in the misclassification equation.

# 5. Higher-order state dependence: Table 5

- Added the comparable four-wave AR(1) benchmark with symmetric error and free
  initial employment.
- Added the same tenure and time-since-work controls used in Table 4.
- Allowed duration covariates to change across origin waves.
- Added a nested AR(2) extension with the full Table 4 Set 2 transition
  controls, followed by a single model whose symmetric-error predictors match
  Table 3 column (1).
- Replaced the log-linear duration terms with piecewise-constant hazards using
  0--3, 4--6, 7--12, 13--24, 25--60, and over-60-month categories.
- Used informed and perturbed starts, exact stage-specific likelihood collapse,
  likelihood-nesting checks, Newton polishing, and score/rank diagnostics.
- Replaced approximate optimisation and inference with exact collapsed
  likelihood estimation and analytical standard errors.

\Needspace{15\baselineskip}

```{r ar2-main-table, results='asis'}
ar2_prob <- read.csv(file.path(project_root,
  "EM-AR2/output/results/ar2_duration_probabilities_latest.csv"))
ar2_fit <- readRDS(file.path(project_root,
  "EM-AR2/output/results/ar2_duration_covariates_4w_latest.rds"))
ar2_runs <- read.csv(file.path(project_root,
  "EM-AR2/output/results/run_summary_latest.csv"))
ar2_base_inf <- read.csv(file.path(project_root,
  "EM-AR2/output/results/analytical_se_latest.csv"))
ar2_base_inf <- ar2_base_inf[ar2_base_inf$model == "sym", ]
ar2_base_run <- ar2_runs[ar2_runs$model == "sym", ]
ar2_base_files <- list.files(file.path(project_root, "EM-AR2/output/results"),
  pattern = "^em_ar2_sym_[0-9_]+[.]rds$", full.names = TRUE)
ar2_base_fit_main <- readRDS(tail(sort(ar2_base_files), 1L))
ar1_inf <- read.csv(file.path(project_root,
  "EM-AR1-4W/output/results/analytical_se_latest.csv"))
ar1_inf <- ar1_inf[ar1_inf$model == "sym_free", ]
ar1_runs <- read.csv(file.path(project_root,
  "EM-AR1-4W/output/results/run_summary_latest.csv"))
ar1_run <- ar1_runs[ar1_runs$model == "sym_free", ]
ar2_extension <- read.csv(file.path(project_root,
  "EM-AR2/output/results/set4_reliability/table5_extension_summary.csv"))
piecewise_fit <- readRDS(file.path(project_root,
  "EM-AR2/output/results/set2_piecewise_duration/fit_ar2_set2_piecewise_latest.rds"))
piecewise_prob <- read.csv(file.path(project_root,
  "EM-AR2/output/results/set2_piecewise_duration/probabilities_ar2_set2_piecewise.csv"))
piecewise_summary <- data.frame(stage = "piecewise", quantity = piecewise_prob$quantity,
  estimate = piecewise_prob$estimate, se = piecewise_prob$se,
  loglik = piecewise_fit$loglik, N = piecewise_fit$n_obs,
  K = piecewise_fit$analytical_inference$diagnostics$parameter_count,
  rank = piecewise_fit$analytical_inference$diagnostics$information_rank,
  condition = piecewise_fit$analytical_inference$diagnostics$information_condition)
ar2_extension <- rbind(ar2_extension, piecewise_summary)
extension_stages <- c("constant", "piecewise", "table3_column1")
ar2_rows <- c(
  "Entry rate: nonemployed in both preceding waves" = "risk_weighted_entry_00",
  "Entry rate: employed two waves ago" = "risk_weighted_entry_10",
  "Exit rate: nonemployed two waves ago" = "risk_weighted_exit_01",
  "Exit rate: employed in both preceding waves" = "risk_weighted_exit_11",
  "Misclassification rate" = "pi")
ar2_base_rows <- c(
  "Entry rate: nonemployed in both preceding waves" = "p_00",
  "Entry rate: employed two waves ago" = "p_10",
  "Exit rate: nonemployed two waves ago" = "p_01",
  "Exit rate: employed in both preceding waves" = "p_11",
  "Misclassification rate" = "pi")
ar1_rows <- c(
  "Entry rate: nonemployed in both preceding waves" = "entry_rate",
  "Entry rate: employed two waves ago" = "entry_rate",
  "Exit rate: nonemployed two waves ago" = "exit_rate",
  "Exit rate: employed in both preceding waves" = "exit_rate",
  "Misclassification rate" = "pi")
cat("\\setcounter{table}{4}\n\\begin{table}[H]\n\\centering\n",
    "\\caption{Four-wave AR(1) and AR(2) estimates}\n",
    "\\scriptsize\n\\setlength{\\tabcolsep}{3pt}\n",
    "\\resizebox{\\textwidth}{!}{%\n",
    "\\begin{tabular}{@{}lrrrrrr@{}}\n\\toprule\n",
    " & AR(1) & AR(2) & \\shortstack{AR(2), duration\\\\ controls} & AR(2), Set 2 & \\shortstack{AR(2), Set 2,\\\\ duration bins} & \\shortstack{+ Table 3 column (1)\\\\ error predictors} \\\\\n\\midrule\n",
    sep = "")
for (label in names(ar2_rows)) {
  ar1 <- ar1_inf[match(ar1_rows[[label]], ar1_inf$quantity), ]
  current <- ar2_prob[match(ar2_rows[[label]], ar2_prob$quantity), ]
  earlier <- ar2_base_inf[
    match(ar2_base_rows[[label]], ar2_base_inf$quantity), ]
  earlier_estimate <- earlier$estimate
  if (grepl("^Exit rate", label)) earlier_estimate <- 1 - earlier_estimate
  extension_quantity <- switch(label,
    "Entry rate: nonemployed in both preceding waves" = "entry_00",
    "Entry rate: employed two waves ago" = "entry_10",
    "Exit rate: nonemployed two waves ago" = "exit_01",
    "Exit rate: employed in both preceding waves" = "exit_11",
    "Misclassification rate" = "mean_misclassification")
  extension <- ar2_extension[
    ar2_extension$quantity == extension_quantity, ]
  extension <- extension[match(extension_stages, extension$stage), ]
  cat(label, " & ", sprintf("%.2f", 100 * ar1$estimate), " & ",
      sprintf("%.2f", 100 * earlier_estimate), " & ",
      sprintf("%.2f", 100 * current$estimate), " & ",
      paste(sprintf("%.2f", 100 * extension$estimate), collapse = " & "),
      " \\\\\n",
      " & (", sprintf("%.2f", 100 * ar1$se), ") & (",
      sprintf("%.2f", 100 * earlier$se), ") & (",
      sprintf("%.2f", 100 * current$se), ") & ",
      paste(sprintf("(%.2f)", 100 * extension$se), collapse = " & "),
      " \\\\\n", sep = "")
}
initial_pair_summary <- function(fit) {
  fn <- function(z) {
    shift <- max(c(0, z)); a <- c(exp(-shift), exp(z - shift))
    p <- a / sum(a); p[3] + p[4]
  }
  z <- fit$eta[1:3]; estimate <- fn(z); gradient <- numeric(3)
  for (j in 1:3) {
    h <- 1e-5 * max(1, abs(z[j])); zp <- zm <- z
    zp[j] <- zp[j] + h; zm[j] <- zm[j] - h
    gradient[j] <- (fn(zp) - fn(zm)) / (2 * h)
  }
  V <- fit$analytical_inference$vcov_eta[1:3, 1:3, drop = FALSE]
  c(estimate = estimate,
    se = sqrt(as.numeric(t(gradient) %*% V %*% gradient)))
}
base_vq <- ar2_base_fit_main$analytical_inference$vcov_quantities
base_initial <- c(
  estimate = sum(ar2_base_fit_main$params$alpha[c("01", "11")]),
  se = sqrt(sum(base_vq[c("alpha_01", "alpha_11"),
                        c("alpha_01", "alpha_11")])))
duration_initial <- initial_pair_summary(ar2_fit)
extension_initial <- ar2_extension[
  ar2_extension$quantity == "initial_wave2_employment.01", ]
extension_initial <- extension_initial[
  match(extension_stages, extension_initial$stage), ]
ar1_initial <- ar1_inf[ar1_inf$quantity == "initial_employment", ]
cat("Initial employment rate & ", sprintf("%.2f", 100 * ar1_initial$estimate),
    " & ", sprintf("%.2f", 100 * base_initial["estimate"]),
    " & ", sprintf("%.2f", 100 * duration_initial["estimate"]),
    " & ", paste(sprintf("%.2f", 100 * extension_initial$estimate),
                   collapse = " & "), " \\\\\n",
    " & (", sprintf("%.2f", 100 * ar1_initial$se), ") & (",
    sprintf("%.2f", 100 * base_initial["se"]), ") & (",
    sprintf("%.2f", 100 * duration_initial["se"]), ") & ",
    paste(sprintf("(%.2f)", 100 * extension_initial$se), collapse = " & "),
    " \\\\\n", sep = "")
extension_stats <- ar2_extension[ar2_extension$quantity == "entry_00", ]
extension_stats <- extension_stats[match(extension_stages, extension_stats$stage), ]
cat("\\addlinespace\nLog likelihood & ",
    formatC(ar1_run$loglik, format = "f", digits = 1, big.mark = ","),
    " & ",
    formatC(ar2_base_run$loglik, format = "f", digits = 1, big.mark = ","),
    " & ",
    formatC(ar2_fit$loglik, format = "f", digits = 1, big.mark = ","),
    " & ", paste(formatC(extension_stats$loglik, format = "f", digits = 1,
                          big.mark = ","), collapse = " & "),
    " \\\\\nN & ", formatC(ar1_run$n_obs, format = "d", big.mark = ","),
    " & ", formatC(ar2_base_run$n_obs, format = "d", big.mark = ","),
    " & ", formatC(max(ar2_runs$n_obs), format = "d", big.mark = ","),
    " & ", paste(formatC(extension_stats$N, format = "d", big.mark = ","),
                  collapse = " & "),
    " \\\\\nParameters & 4 & 8 & ", length(ar2_fit$eta), " & ",
    paste(extension_stats$K, collapse = " & "),
    " \\\\\nSecond employment lag & No & Yes & Yes & Yes & Yes & Yes",
    " \\\\\nDuration controls & No & No & Yes & Yes & Yes & Yes",
    " \\\\\nDuration form & --- & --- & Log-linear & Log-linear & Piecewise constant & Log-linear",
    " \\\\\nSet 2 transition controls & No & No & No & Yes & Yes & Yes",
    " \\\\\nMisclassification predictors & Constant & Constant & Constant & Constant & Constant & \\shortstack{Table 3\\\\ column (1)}",
    " \\\\\n\\bottomrule\n\\end{tabular}}\\\\[4pt]\n",
    "{\\footnotesize\\textit{Notes.} Entries are model-implied quarterly probabilities in percent. Covariate specifications report posterior risk-set and survey-weighted averages; the misclassification row is the corresponding posterior mean probability. Initial employment is the latent employment probability in wave 2. Parentheses contain analytical sandwich/delta-method standard errors. The AR(1) rate is repeated across the two corresponding AR(2) rows because the first-order model imposes equality across the second-lag state. Every specification has one latent type, symmetric misclassification, and free initial conditions. Set 2 includes the Table 4 demographic, tenure, time-since-work, contract, and informal-sector transition controls. Piecewise duration uses 0--3 months as the omitted category, followed by 4--6, 7--12, 13--24, 25--60, and over 60 months; never-worked and missing durations remain separate. The final error equation uses age, education, race, and gender inconsistency indicators plus mutually exclusive indicators for exactly two, three, or four inconsistencies, matching Table 3 column (1). The three Set 2 columns use the same 308,240 observations. The log-linear constant-error and final error models are nested; the piecewise and log-linear duration forms are non-nested.}\n",
    "\\end{table}\n", sep = "")
```

```{r ar2-coefficient-appendix, results='asis'}
ar2_coef <- read.csv(file.path(project_root,
  "EM-AR2/output/results/ar2_duration_coefficients_latest.csv"))
latest_result <- function(path, pattern) {
  files <- list.files(file.path(project_root, path), pattern = pattern,
    full.names = TRUE)
  if (!length(files)) stop("Missing saved fit: ", pattern)
  tail(sort(files), 1L)
}
numeric_jacobian <- function(fn, x, step = 1e-5) {
  base <- fn(x)
  out <- matrix(NA_real_, length(base), length(x),
    dimnames = list(names(base), names(x)))
  for (j in seq_along(x)) {
    h <- step * max(1, abs(x[j])); xp <- xm <- x
    xp[j] <- xp[j] + h; xm[j] <- xm[j] - h
    out[, j] <- (fn(xp) - fn(xm)) / (2 * h)
  }
  out
}
coefficient_inference <- function(eta, vcov_eta, transform) {
  estimate <- transform(eta)
  jacobian <- numeric_jacobian(transform, eta)
  se <- sqrt(pmax(diag(jacobian %*% vcov_eta %*% t(jacobian)), 0))
  data.frame(term = names(estimate), estimate = unname(estimate), se = unname(se))
}

ar1_fit <- readRDS(latest_result("EM-AR1-4W/output/results",
  "^ar1_4w_sym_free_[0-9_]+[.]rds$"))
ar1_coefficient_transform <- function(z) setNames(unname(c(
  z["alpha"], qnorm(plogis(z["theta0"])),
  qnorm(plogis(z["theta1"])), z["pi"])),
  c("initial_employment_logit", "entry_intercept",
    "persistence_intercept", "misclassification_logit"))
ar1_coef <- coefficient_inference(ar1_fit$eta,
  ar1_fit$analytical_inference$vcov_eta, ar1_coefficient_transform)

ar2_base_fit <- readRDS(latest_result("EM-AR2/output/results",
  "^em_ar2_sym_[0-9_]+[.]rds$"))
ar2_base_eta <- setNames(unname(c(
  log(ar2_base_fit$params$alpha["10"] / ar2_base_fit$params$alpha["00"]),
  log(ar2_base_fit$params$alpha["01"] / ar2_base_fit$params$alpha["00"]),
  log(ar2_base_fit$params$alpha["11"] / ar2_base_fit$params$alpha["00"]),
  qlogis(ar2_base_fit$params$theta0),
  qlogis(ar2_base_fit$params$theta01 / (1 - ar2_base_fit$params$theta0)),
  qlogis(1 - ar2_base_fit$params$theta1 - ar2_base_fit$params$theta10),
  qlogis(ar2_base_fit$params$theta10 /
    (ar2_base_fit$params$theta1 + ar2_base_fit$params$theta10)),
  qlogis(ar2_base_fit$params$pi / 0.5))),
  c("alpha10", "alpha01", "alpha11", "p00", "gap10", "p01", "gap11", "pi"))
ar2_base_coefficient_transform <- function(z) {
  p00 <- plogis(z["p00"])
  p10 <- p00 + (1 - p00) * plogis(z["gap10"])
  p01 <- plogis(z["p01"])
  p11 <- p01 + (1 - p01) * plogis(z["gap11"])
  setNames(unname(c(z["alpha10"], z["alpha01"], z["alpha11"],
    qnorm(p00), qnorm(p10) - qnorm(p00), qnorm(p01),
    qnorm(p11) - qnorm(p01), z["pi"])),
    c("alpha_logit_10", "alpha_logit_01", "alpha_logit_11",
      "entry_intercept", "entry_lag2", "persistence_intercept",
      "persistence_lag2", "misclassification_logit"))
}
ar2_base_coef <- coefficient_inference(ar2_base_eta,
  ar2_base_fit$analytical_inference$vcov_eta,
  ar2_base_coefficient_transform)

coefficient_labels <- c(
  initial_employment_logit = "Initial employment logit",
  alpha_logit_10 = "Initial-pair logit: 10",
  alpha_logit_01 = "Initial-pair logit: 01",
  alpha_logit_11 = "Initial-pair logit: 11",
  entry_intercept = "Entry: intercept",
  entry_lag2 = "Entry: employed two waves ago",
  entry_log_time_since_work = "Entry: log time since work",
  entry_never_worked = "Entry: never worked",
  entry_timegap_missing = "Entry: timegap missing",
  persistence_intercept = "Persistence: intercept",
  persistence_lag2 = "Persistence: employed two waves ago",
  persistence_log_tenure = "Persistence: log tenure",
  persistence_tenure_missing = "Persistence: tenure missing",
  misclassification_logit = "Misclassification logit")
coefficient_terms <- names(coefficient_labels)
coefficient_cell <- function(tbl, term, field) {
  value <- tbl[[field]][match(term, tbl$term)]
  if (!length(value) || is.na(value)) return(NA_real_)
  value
}
duration_coef <- ar2_coef
cat("\\begin{table}[H]\n\\centering\n",
    "\\caption*{Appendix Table A5: Four-wave AR(1) and AR(2) coefficients}\n",
    "\\scriptsize\n\\setlength{\\tabcolsep}{5pt}\n",
    "\\begin{tabular}{@{}lrrr@{}}\n\\toprule\n",
    "Coefficient & AR(1) & AR(2) & AR(2), duration controls \\\\\n\\midrule\n",
    sep = "")
for (term in coefficient_terms) {
  estimates <- c(coefficient_cell(ar1_coef, term, "estimate"),
    coefficient_cell(ar2_base_coef, term, "estimate"),
    coefficient_cell(duration_coef, term, "estimate"))
  ses <- c(coefficient_cell(ar1_coef, term, "se"),
    coefficient_cell(ar2_base_coef, term, "se"),
    coefficient_cell(duration_coef, term, "se"))
  estimate_cells <- ifelse(is.na(estimates), "---", sprintf("%.3f", estimates))
  se_cells <- ifelse(is.na(ses), "", paste0("(", sprintf("%.3f", ses), ")"))
  cat(coefficient_labels[[term]], " & ", paste(estimate_cells, collapse = " & "),
      " \\\\\n & ", paste(se_cells, collapse = " & "), " \\\\\n", sep = "")
}
cat("\\bottomrule\n\\end{tabular}\n\\end{table}\n")
```

```{r ar2-extension-coefficient-appendix, results='asis'}
extension_coef_paths <- c(
  constant = "EM-AR2/output/results/set4_reliability/coefficients_ar2_set4_constant.csv",
  piecewise = "EM-AR2/output/results/set2_piecewise_duration/coefficients_ar2_set2_piecewise.csv",
  table3_column1 = "EM-AR2/output/results/set4_reliability/coefficients_ar2_set4_table3_column1.csv")
extension_coef <- lapply(extension_coef_paths, function(path)
  read.csv(file.path(project_root, path)))
names(extension_coef) <- extension_stages
extension_terms <- unique(unlist(lapply(extension_coef, `[[`, "term")))
extension_term_labels <- c(
  alpha_logit_10 = "Initial-pair logit: 10",
  alpha_logit_01 = "Initial-pair logit: 01",
  alpha_logit_11 = "Initial-pair logit: 11",
  entry_intercept = "Entry: intercept",
  entry_lag2 = "Entry: employed two waves ago",
  entry_age = "Entry: age",
  entry_age_sq = "Entry: age squared",
  entry_educ = "Entry: education",
  entry_race_2 = "Entry: race 2",
  entry_race_3 = "Entry: race 3",
  entry_race_4 = "Entry: race 4",
  entry_female = "Entry: female",
  entry_log_time_since_work = "Entry: log time since work",
  entry_timegap_4_6m = "Entry: timegap 4--6 months",
  entry_timegap_7_12m = "Entry: timegap 7--12 months",
  entry_timegap_13_24m = "Entry: timegap 13--24 months",
  entry_timegap_25_60m = "Entry: timegap 25--60 months",
  entry_timegap_over_60m = "Entry: timegap over 60 months",
  entry_never_worked = "Entry: never worked",
  entry_timegap_missing = "Entry: timegap missing",
  persistence_intercept = "Persistence: intercept",
  persistence_lag2 = "Persistence: employed two waves ago",
  persistence_age = "Persistence: age",
  persistence_age_sq = "Persistence: age squared",
  persistence_educ = "Persistence: education",
  persistence_race_2 = "Persistence: race 2",
  persistence_race_3 = "Persistence: race 3",
  persistence_race_4 = "Persistence: race 4",
  persistence_female = "Persistence: female",
  persistence_log_tenure = "Persistence: log tenure",
  persistence_tenure_4_6m = "Persistence: tenure 4--6 months",
  persistence_tenure_7_12m = "Persistence: tenure 7--12 months",
  persistence_tenure_13_24m = "Persistence: tenure 13--24 months",
  persistence_tenure_25_60m = "Persistence: tenure 25--60 months",
  persistence_tenure_over_60m = "Persistence: tenure over 60 months",
  persistence_tenure_missing = "Persistence: tenure missing",
  persistence_permanent_contract = "Persistence: permanent contract",
  persistence_informal_sector = "Persistence: informal sector",
  error_intercept = "Error: intercept",
  age_inconsistency = "Error: age inconsistency",
  education_inconsistency = "Error: education inconsistency",
  race_inconsistency = "Error: race inconsistency",
  gender_inconsistency = "Error: gender inconsistency",
  two_inconsistencies = "Error: exactly two inconsistencies",
  three_inconsistencies = "Error: exactly three inconsistencies",
  four_inconsistencies = "Error: exactly four inconsistencies")
cat("\\begingroup\n\\scriptsize\n\\setlength{\\tabcolsep}{5pt}\n",
    "\\begin{longtable}{@{}lrrr@{}}\n",
    "\\caption*{Appendix Table A5B: Set 2 AR(2) extension coefficients}\\\\\n",
    "\\toprule\nCoefficient & \\shortstack{Set 2,\\\\ log-linear} & \\shortstack{Set 2,\\\\ duration bins} & \\shortstack{+ Table 3 column (1)\\\\ error predictors} \\\\\n\\midrule\n",
    "\\endfirsthead\n\\toprule\nCoefficient & \\shortstack{Set 2,\\\\ log-linear} & \\shortstack{Set 2,\\\\ duration bins} & \\shortstack{+ Table 3 column (1)\\\\ error predictors} \\\\\n\\midrule\n",
    "\\endhead\n", sep = "")
for (term in extension_terms) {
  estimates <- vapply(extension_coef, function(tbl)
    coefficient_cell(tbl, term, "estimate"), numeric(1))
  ses <- vapply(extension_coef, function(tbl)
    coefficient_cell(tbl, term, "se"), numeric(1))
  estimate_cells <- ifelse(is.na(estimates), "---", sprintf("%.3f", estimates))
  se_cells <- ifelse(is.na(ses), "", paste0("(", sprintf("%.3f", ses), ")"))
  cat(extension_term_labels[[term]], " & ",
      paste(estimate_cells, collapse = " & "), " \\\\\n & ",
      paste(se_cells, collapse = " & "), " \\\\\n", sep = "")
}
cat("\\bottomrule\n\\end{longtable}\n\\endgroup\n")
```

```{r ar2-piecewise-hazard-appendix, results='asis'}
piecewise_profile <- read.csv(file.path(project_root,
  "EM-AR2/output/results/set2_piecewise_duration/piecewise_duration_hazard_profile.csv"))
duration_order <- c("0--3 months", "4--6 months", "7--12 months",
  "13--24 months", "25--60 months", "Over 60 months")
profile_cell <- function(equation, lag, duration, field) {
  row <- piecewise_profile$equation == equation &
    piecewise_profile$second_lag_employed == lag &
    piecewise_profile$duration == duration
  piecewise_profile[[field]][row]
}
cat("\\begin{table}[H]\n\\centering\n",
    "\\caption*{Appendix Table A5C: Reference piecewise duration hazards}\n",
    "\\small\n\\setlength{\\tabcolsep}{6pt}\n",
    "\\begin{tabular}{@{}lrrrr@{}}\n\\toprule\n",
    " & \\multicolumn{2}{c}{Entry} & \\multicolumn{2}{c}{Exit} \\\\\n",
    "\\cmidrule(lr){2-3}\\cmidrule(lr){4-5}\n",
    "Duration & \\shortstack{Two waves\\\\ nonemployed} & \\shortstack{Recent\\\\ exit} & \\shortstack{Recent\\\\ entry} & \\shortstack{Two waves\\\\ employed} \\\\\n\\midrule\n",
    sep = "")
for (duration in duration_order) {
  estimates <- c(profile_cell("Entry", 0, duration, "estimate"),
    profile_cell("Entry", 1, duration, "estimate"),
    profile_cell("Exit", 0, duration, "estimate"),
    profile_cell("Exit", 1, duration, "estimate"))
  ses <- c(profile_cell("Entry", 0, duration, "se"),
    profile_cell("Entry", 1, duration, "se"),
    profile_cell("Exit", 0, duration, "se"),
    profile_cell("Exit", 1, duration, "se"))
  cat(duration, " & ", paste(sprintf("%.2f", 100 * estimates), collapse = " & "),
      " \\\\\n & ", paste(sprintf("(%.2f)", 100 * ses), collapse = " & "),
      " \\\\\n", sep = "")
}
cat("\\bottomrule\n\\end{tabular}\\\\[4pt]\n",
    "{\\footnotesize\\textit{Notes.} Entries are quarterly probabilities in percent at reference covariate values, with analytical delta-method standard errors in parentheses. Age, age squared, and education are at their standardized means; the reference respondent is male and in the baseline race category. Exit profiles set permanent-contract and informal-sector indicators to zero. Never-worked and missing-duration indicators are also zero. The 0--3 month category is omitted in estimation.}\n",
    "\\end{table}\n", sep = "")
```

```{r ar2-apparent-transition-appendix, results='asis'}
transition_implications <- read.csv(file.path(project_root,
  "EM-AR2/output/results/set4_reliability/apparent_transition_decomposition.csv"),
  check.names = FALSE)
direction_order <- c("All apparent transitions", "Apparent entries",
                     "Apparent exits")
model_order <- c("Table 1: raw/no error", "Table 1: symmetric error",
                 "Table 5: final AR(2)")
model_labels <- c(
  `Table 1: raw/no error` = "Table 1: raw/no error",
  `Table 1: symmetric error` = "Table 1: symmetric-error adjustment",
  `Table 5: final AR(2)` = "Table 5: final AR(2)")
cat("\\begin{table}[H]\n\\centering\n",
    "\\caption*{Appendix Table A5D: Implications for apparent two-wave transitions}\n",
    "\\small\n\\setlength{\\tabcolsep}{7pt}\n",
    "\\begin{tabular}{@{}lrrr@{}}\n\\toprule\n",
    "Model & Classification only & Genuine reversal & Genuine persistence \\\\\n",
    "\\midrule\n", sep = "")
for (direction in direction_order) {
  cat("\\multicolumn{4}{l}{\\textit{", direction, "}} \\\\\n", sep = "")
  rows <- transition_implications[
    transition_implications$direction == direction, ]
  rows <- rows[match(model_order, rows$model), ]
  for (i in seq_len(nrow(rows))) {
    cat(model_labels[[rows$model[i]]], " & ",
        sprintf("%.1f", 100 * rows$classification_only[i]), " & ",
        sprintf("%.1f", 100 * rows$true_reversal[i]), " & ",
        sprintf("%.1f", 100 * rows$true_persistent[i]),
        " \\\\\n\\addlinespace[1pt]\n", sep = "")
  }
  if (direction != tail(direction_order, 1L)) cat("\\addlinespace\n")
}
cat("\\bottomrule\n\\end{tabular}\\\\[4pt]\n",
    "{\\footnotesize\\textit{Notes.} Entries are survey-weighted percentages conditional only on the two reports forming an apparent transition; the subsequent report is integrated out. Classification only means that reported status changes but latent status does not. Genuine reversal means that latent status changes and returns to its original state in the next period; genuine persistence means that it remains in the new state. The three outcomes are mutually exclusive and exhaustive. The Table 1 columns use the nonstationary three-wave estimates; the final AR(2) column uses the four-wave Set 2 model with the Table 3 column (1) predictors of symmetric classification error. Because the Table 1 and Table 5 estimation samples differ, comparisons combine model and sample differences.}\n",
    "\\end{table}\n", sep = "")
```

Transition coefficients are on a common probit scale. The unrestricted AR(1)
and no-duration AR(2) transition probabilities are re-expressed in their
observationally equivalent probit parameterisations; the duration-controlled
AR(2) is estimated directly by probit. Initial-state coefficients are logits,
and the misclassification coefficient is $\operatorname{logit}(2\pi)$ in all
three original models and is reported directly for the staged extension.
Parentheses contain analytical sandwich/delta-method standard errors. Duration
variables and their missingness indicators vary across origin waves. The
information-matrix condition numbers are 14.6 for AR(1), 130.7 for AR(2), and
388,646 for duration-controlled AR(2), 746,558 for the constant-error Set 2
extension, 1,387,680 for the piecewise-duration model, and 7,710,884 for the
Table 3 column (1) error model. All three Set 2 models are locally full rank and
pass the normalized-score criterion.

## Interpretation

- The AR(1) model estimates quarterly entry and exit of 3.20% and 3.13%, with
  a 2.70% symmetric misclassification probability.
- Duration controls substantially reduce the large AR(2) transition
  differences but do not eliminate them.
- With Set 2 controls and constant error, entry and exit after two periods in
  the same state are 4.72% and 4.56%; following a recent reversal they are
  31.13% and 32.37%. Observed transition heterogeneity therefore does not
  explain away the second-lag pattern.
- Piecewise duration hazards improve the log likelihood by 156.4 despite being
  non-nested with the log-linear model. Entry declines from 15.00% at 0--3
  months to 4.85% after 60 months for otherwise reference nonemployed records;
  exit declines from 13.20% to 3.63% for otherwise reference records employed
  in both preceding waves. Thus the model finds strong nonlinear duration
  dependence.
- Nevertheless, the posterior risk-set averages remain 4.73% versus 31.34%
  for entry and 32.25% versus 4.59% for exit. Flexible piecewise duration
  controls therefore do not materially reduce the large AR(2) effects.
- Making misclassification depend on the Table 3 column (1) predictors raises
  its posterior mean from 0.93% to 1.79%. The corresponding transition rates
  are 4.32%, 28.98%, 30.18%, and 4.03%, so the pronounced second-lag pattern
  remains after applying the same observed reliability controls used in Table 3.
- Appendix Table A5D shows that the final model attributes 38.5% of apparent
  transitions to classification error, 18.0% to genuine transitions that
  reverse next period, and 43.5% to genuine persistent transitions. Among the
  61.5% inferred to be genuine, 29.3% reverse in the following period. This is
  more balanced than the Table 1 symmetric-error model, which attributes 67.0%
  to error and only 1.0% to genuine short-lived transitions.
- Surviving AR(2) dependence may reflect unobserved duration, state dependence,
  latent heterogeneity, or within-quarter transitions; the current estimates do
  not distinguish these mechanisms.
- The extension is locally identified and has very small final scores, but its
  rising information condition number calls for sensitivity checks before it
  is treated as the preferred specification.

## To do

- Compare piecewise hazards with spline duration controls and allow the duration
  profiles to interact with the second-lag state.
- Profile the weakly curved directions in the two Set 2 extensions and check
  stability with additional long starts and alternative covariate scaling.
- Compare the staged extensions on a common sample with the simpler AR(1),
  AR(2), and duration-controlled benchmarks.
- With additional waves, test latent heterogeneity, wave-varying transitions,
  and correlated classification errors as competing explanations.

# 6. Unobserved heterogeneity: Table 6 and Appendix Table A6

- Re-estimated the two-type finite-mixture model by direct likelihood
  maximisation with multiple starts and audited local identification.
- Removed the underidentified three-wave FMM table: only the no-error model with free initial
  conditions is locally identified; symmetric error cannot be separated from
  two-type transition heterogeneity.
- Developed a four-wave alternative with observed covariates, duration
  controls, and class-specific misclassification intercepts.
- Compared all four specifications in one main table, including the nested
  no-misclassification model, and moved inconsistency-specific error rates to
  a secondary table.

\Needspace{44\baselineskip}

**Table 6. Comparison of all four-wave two-type FMM specifications**

\begin{center}
\scriptsize
\setlength{\tabcolsep}{2.7pt}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lrrrrrrrr@{}}
\toprule
& \multicolumn{4}{c}{No controls} & \multicolumn{4}{c}{Controls} \\
\cmidrule(lr){2-5}\cmidrule(lr){6-9}
& \multicolumn{2}{c}{No error} & \multicolumn{2}{c}{Common error}
& \multicolumn{2}{c}{Common error} & \multicolumn{2}{c}{Type-specific error} \\
\cmidrule(lr){2-3}\cmidrule(lr){4-5}\cmidrule(lr){6-7}\cmidrule(lr){8-9}
& Stable & Mobile & Stable & Mobile & Stable & Mobile & Stable & Mobile \\
\midrule
Latent-class share (\%) & 75.65 & 24.35 & 92.13 & 7.87 & 89.63 & 10.37 & 58.17 & 41.83 \\
& (1.00) & (1.00) & (0.37) & (0.37) & (1.26) & (1.26) & (3.05) & (3.05) \\
Entry rate (\%) & 0.46 & 30.46 & 2.36 & 42.16 & 3.20 & 27.30 & 1.33 & 11.98 \\
& (0.18) & (0.75) & (0.07) & (1.15) & (0.12) & (2.49) & (0.15) & (0.64) \\
Exit rate (\%) & 0.004 & 33.63 & 2.21 & 47.04 & 3.01 & 26.87 & 0.75 & 11.18 \\
& (0.240) & (0.77) & (0.08) & (1.16) & (0.16) & (1.57) & (0.20) & (0.65) \\
Misclassification rate (\%) & 0.00 & 0.00 & 1.35 & 1.35 & 1.74 & 1.98 & 0.68 & 4.60 \\
& (fixed) & (fixed) & (0.03) & (0.03) & (0.03) & (0.03) & (0.09) & (0.18) \\
Initial employment rate (\%) & 47.73 & 48.17 & 47.74 & 47.95 & 47.38 & 50.59 & 44.61 & 52.33 \\
& (0.18) & (0.45) & (0.13) & (0.74) & (0.22) & (1.75) & (0.58) & (0.56) \\
\addlinespace
Log-likelihood & \multicolumn{2}{c}{-451,874.52}
& \multicolumn{2}{c}{-450,805.93}
& \multicolumn{2}{c}{-428,777.56}
& \multicolumn{2}{c}{-428,568.60} \\
$N$ & \multicolumn{2}{c}{308,240}
& \multicolumn{2}{c}{308,240}
& \multicolumn{2}{c}{308,240}
& \multicolumn{2}{c}{308,240} \\
\bottomrule
\end{tabular}
}
\end{center}

All estimates except the log likelihood and observation count are percentages;
analytical survey-weighted sandwich standard errors, transformed by the delta
method, are in parentheses. Entry and exit rates are posterior risk-set weighted
in the controlled models. Posterior mean misclassification averages each
respondent-wave's fitted posterior mismatch probability using survey weights and
posterior class membership. Controls comprise observed covariates and duration
controls. Both controlled models use common slopes and class-specific transition
intercepts. The last model additionally makes the misclassification intercept
class-specific; the age- and education-inconsistency slopes remain common. In
the no-error model, all misclassification probabilities are fixed at zero. The
no-error model has full probability-Jacobian rank, but weak curvature around
its near-zero stable-class exit estimate warrants caution. Pattern-specific
implied misclassification rates are reported separately in Appendix Table A6.

\Needspace{13\baselineskip}

**Appendix Table A6. Misclassification rates by inconsistency pattern**

| Inconsistency pattern | Common error: Stable | Common error: Mobile | Type-specific error: Stable | Type-specific error: Mobile | Mobile minus stable |
|:--|--:|--:|--:|--:|--:|
| None | 0.72 (0.02) | 0.72 (0.02) | 0.14 (0.04) | 2.37 (0.15) | 2.22 |
| Age | 14.72 (0.31) | 14.72 (0.31) | 6.28 (0.98) | 35.52 (1.05) | 29.25 |
| Education | 3.73 (0.12) | 3.73 (0.12) | 0.89 (0.21) | 11.83 (0.53) | 10.94 |
| Age and education | 34.85 (0.40) | 34.85 (0.40) | 23.62 (1.97) | 46.93 (0.34) | 23.31 |

Estimates are misclassification rates in percent; analytical standard errors
are in parentheses. Both specifications include observed covariates and
duration controls. The common-error model has a common misclassification
intercept; the type-specific model has separate stable- and mobile-class
intercepts. Age- and education-inconsistency slopes are common across classes.

**Appendix Table A6, Panel B. Preferred FMM coefficients and diagnostics**

| Quantity | Estimate | Analytical SE |
|:--|--:|--:|
| Persistence: log tenure | 0.650 | 0.015 |
| Entry: log time since work | -0.325 | 0.007 |
| Entry: never worked | -0.860 | 0.018 |
| Persistence: permanent contract | 0.399 | 0.016 |
| Persistence: informal sector | -0.266 | 0.016 |
| Mobile-minus-stable misclassification intercept | 2.839 | 0.225 |
| Wald statistic for common error intercept | 12.64 | -- |
| Entry-risk peak age | 38 | -- |
| Persistence peak age | 53 | -- |
| Information-matrix rank | 32/32 | -- |
| Minimum information eigenvalue | 0.0000120 | -- |
| Information condition number | 29,216 | -- |
| Maximum weighted score | 0.0000032 | -- |

Coefficients are probit-index effects except the logit-scale difference between
the class-specific misclassification intercepts. The controlled sample contains
308,240 people and collapses exactly to 229,809 likelihood cells.

\Needspace{16\baselineskip}

## Interpretation

- The three-wave symmetric-error FMM is underidentified and should not be
  reported structurally; four waves are required to separate the two mechanisms.
- Imposing no misclassification pushes the stable-class exit estimate to
  essentially zero and attributes 30--34% quarterly transition rates to the
  mobile class. Allowing error materially improves fit and avoids this extreme
  decomposition.
- Four waves support a large stable/mobile distinction after controlling for
  observed characteristics and duration.
- Allowing the misclassification probability to differ by type makes the mobile
  class larger and less extreme, showing that the uncontrolled FMM mixed
  genuine mobility, observed heterogeneity, and heterogeneous classification
  error.
- In the preferred controlled model, the survey-weighted posterior mean
  misclassification rate is 0.68% in the stable class and 4.60% in the mobile
  class, a much sharper reliability contrast than the baseline probabilities
  alone convey.
- The type-specific probabilities describe joint latent classifications; they
  do not show that mobility causes reporting or matching error.
- The four-wave results are best presented as a replacement or appendix
  extension rather than as validation of the underidentified three-wave results.

## To do

- Complete profile-likelihood, multistart, and simulation checks for the
  four-wave model's identification and class separation.
- Test covariate-dependent class membership and, with a longer panel, whether
  latent types can be combined with AR(2) state dependence.

# 7. Endogenous duration dependence: Table 7

- Modelled transition hazards as functions of endogenous tenure and time since
  work, separately from the age/education-consistency analysis in Table 3.
- Repaired the tenure-contamination likelihood and optimisation diagnostics.
- Retained exact consistency between clean duration reports and latent
  employment histories.
- Integrated unavailable wrong-state duration clocks instead of imputing them.
- Replaced constant and power-law transition hazards with tail-safe piecewise
  hazards.
- Added contamination in reported time since work and tested local and joint
  alternatives.

\Needspace{42\baselineskip}

```{r tenure-main-table, results='asis'}
piece <- read.csv(file.path(project_root,
  "EM-tenure/output/results/piecewise_hazard/model_comparison_latest.csv"),
  check.names = FALSE)
robust <- read.csv(file.path(project_root,
  "EM-tenure/output/results/timegap_contamination_robustness/model_comparison_latest.csv"),
  check.names = FALSE)
tenure_inf <- read.csv(file.path(project_root,
  "EM-tenure/output/results/analytical_inference/table7_analytical_se_latest.csv"))
tenure_tab <- rbind(
  data.frame(model = piece$model[1:3], loglik = piece$loglik[1:3],
    pi = piece$pi[1:3], eps_g = piece$eps[1:3], eps_d = 0,
    weighted_exit = piece$weighted_exit[1:3],
    weighted_entry = piece$weighted_entry[1:3]),
  robust[2:4, c("model", "loglik", "pi", "eps_g", "eps_d",
                 "weighted_exit", "weighted_entry")])
tenure_tab$model <- c("Constant, exact", "Power law, exact",
  "Piecewise, exact", "Independent error", "Local error",
  "Joint per-wave error")
inference_value <- function(quantity, field) vapply(tenure_tab$model, function(m) {
  z <- tenure_inf[tenure_inf$model == m & tenure_inf$quantity == quantity, field]
  if (length(z) != 1L) stop("Missing Table 7 inference: ", m, " / ", quantity)
  z
}, numeric(1L))
quantity_map <- c(
  "Entry rate" = "entry_rate",
  "Exit rate" = "exit_rate",
  "Misclassification rate" = "misclassification_rate",
  "Initial employment rate" = "initial_employment_rate",
  "Tenure contamination rate" = "tenure_contamination_rate",
  "Timegap contamination rate" = "timegap_contamination_rate")
values <- 100 * t(vapply(quantity_map,
  function(q) inference_value(q, "estimate"), numeric(nrow(tenure_tab))))
se_values <- 100 * t(vapply(quantity_map,
  function(q) inference_value(q, "se"), numeric(nrow(tenure_tab))))
rownames(values) <- rownames(se_values) <- names(quantity_map)
colnames(values) <- colnames(se_values) <- tenure_tab$model
cat("\\setcounter{table}{6}\n\\begin{table}[H]\n\\centering\n",
    "\\caption{Tenure-contamination misclassification and transition rates}\n",
    "\\scriptsize\n\\setlength{\\tabcolsep}{4pt}\n",
    "\\resizebox{\\textwidth}{!}{%\n",
    "\\begin{tabular}{@{}lrrrrrr@{}}\n\\toprule\n & ",
    paste(tenure_tab$model, collapse = " & "), " \\\\\n\\midrule\n", sep = "")
for (label in rownames(values)) {
  cat(label, " & ", paste(sprintf("%.2f", values[label, ]),
                           collapse = " & "), " \\\\\n", sep = "")
  se_text <- ifelse(is.finite(se_values[label, ]),
                    sprintf("(%.2f)", se_values[label, ]), "fixed")
  cat(" & ", paste(se_text, collapse = " & "), " \\\\\n", sep = "")
}
cat("\\addlinespace\nLog likelihood & ",
    paste(formatC(tenure_tab$loglik, format = "f", digits = 1,
                  big.mark = ","), collapse = " & "), " \\\\\n",
    "N & ", paste(rep("402,923", nrow(tenure_tab)), collapse = " & "),
    " \\\\\n\\addlinespace\n",
    "Transition-hazard form & Constant & Power law & Piecewise & Piecewise & Piecewise & Piecewise \\\\\n",
    "Tenure-error model & Independent marginal & Independent marginal & Independent marginal & Independent marginal & Independent marginal & Independent marginal \\\\\n",
    "Timegap-error model & None (exact) & None (exact) & None (exact) & Independent marginal & Local category & Joint per-wave marginal",
    " \\\\\n\\bottomrule\n\\end{tabular}}\\\\[4pt]\n",
    "{\\footnotesize\\textit{Notes.} Rates are in percent and follow the common entry, exit, misclassification, and initial-employment order used in the other main tables; model-specific duration-error rates follow. Entry and exit are posterior risk-set and survey-weighted quarterly probabilities. Parentheses contain numerical observed-information/delta-method standard errors from the full covariance matrix. In every model, a contaminated tenure report is an independent draw from the marginal tenure distribution; otherwise it must follow the exact clock. `Exact' means that no separate timegap-contamination process is allowed; status misclassification and tenure contamination remain allowed.}\n",
    "\\end{table}\n", sep = "")
```

For the joint model, 16.62% is the per-report probability. It implies a 30.48%
probability that at least one report in a two-report pair is contaminated.

\Needspace{12\baselineskip}

**Appendix Table A7, Panel A. Exact and alternative duration profiles**

| Duration interval | Exact exit hazard | Exact exit (%) | Independent-error exit (%) | Joint-model exit (%) |
|:--|--:|--:|--:|--:|
| 0--3 months | 0.3420 | 8.35 | 7.79 | 7.79 |
| 3--12 months | 0.3558 | 8.51 | 7.90 | 7.90 |
| 1--3 years | 0.1757 | 4.30 | 4.23 | 4.23 |
| 3--5 years | 0.1522 | 3.73 | 3.67 | 3.67 |
| 5+ years | 0.1353 | 3.33 | 3.27 | 3.27 |

| Duration interval | Exact entry hazard | Exact entry (%) | Independent-error entry (%) | Joint-model entry (%) |
|:--|--:|--:|--:|--:|
| 0--3 months | 0.5060 | 9.15 | 7.66 | 7.88 |
| 3--12 months | 0.2620 | 6.34 | 5.00 | 5.07 |
| 1--3 years | 0.1098 | 2.71 | 2.27 | 2.27 |
| 3--5 years | 0.1177 | 2.90 | 2.57 | 2.57 |
| 5+ years | 0.1425 | 3.50 | 2.82 | 2.83 |

\Needspace{24\baselineskip}

**Appendix Table A7, Panel B. Auxiliary model comparisons**

| Quantity | Constant, exact | Power law, exact | Piecewise, exact |
|:--|--:|--:|--:|
| Parameters | 5 | 7 | 13 |
| AIC | 5,020,870.7 | 4,951,367.6 | 4,924,420.8 |
| Mean employment duration (months) | 68.9 | 68.1 | 67.0 |
| Mean nonemployment duration (months) | 72.0 | 1,257.2 | 76.7 |
| Comparable two-report timegap contamination (%) | 0 | 0 | 0 |
| Analytical SE | fixed | fixed | fixed |

| Quantity | Independent error | Local error | Joint per-wave error |
|:--|--:|--:|--:|
| Parameters | 14 | 14 | 14 |
| AIC | 4,660,178.6 | 4,686,130.8 | 4,640,617.9 |
| Mean employment duration (months) | 69.6 | 69.6 | 69.6 |
| Mean nonemployment duration (months) | 95.6 | 102.2 | 94.8 |
| Comparable two-report timegap contamination (%) | 28.04 | 39.20 | 30.48 |
| Analytical SE | 0.11 | 0.15 | 0.12 |

**Appendix Table A7, Panel C. Effect of integrating unavailable wrong-state clocks**

| Quantity | Imputed clock | Integrated clock |
|:--|--:|--:|
| Entry rate (%) | 2.08 | 1.96 |
| Exit rate (%) | 4.91 | 4.51 |
| Misclassification rate (%) | 10.35 | 10.53 |
| Tenure contamination rate (%) | 24.81 | 25.09 |
| Exit duration slope | -0.4413 | -0.3742 |
| Entry duration slope | -0.9148 | -0.8782 |
| Mean nonemployment duration (months) | 3,800 | 1,257 |

**Appendix Table A7, Panel D. Timegap-contamination mechanism decomposition**

| Interview interval | Clock-feasible (%) | Reset-compatible contradiction (%) | Other contradiction (%) | Posterior contamination (%) |
|:--|--:|--:|--:|--:|
| Wave 1 to wave 2 | 84.42 | 2.82 | 12.76 | 28.20 |
| Wave 2 to wave 3 | 84.95 | 2.98 | 12.07 | 27.87 |

In the power-law model, exit falls from 7.42% at zero tenure to 3.22% at
ten years, while entry falls from 8.51% to 1.18%. Integrating rather than
imputing wrong-state clocks affects 15.5% of the posterior employment risk set
and 4.5% of the nonemployment risk set. In the piecewise exact model, median
employment and nonemployment durations are 35.3 and 51.4 months. For the
independent-error model, clock-impossible shares are 15.6% and 15.0%, while
average contaminated-branch probabilities among feasible reports are 14.96%
and 15.09%. Roughly 10% of fitted contamination mass is reset-compatible,
43--45% comes from contradictions that a hidden reset cannot explain, and
45--46% is allocated among clock-feasible reports. Joint-model exit risks
differ from the independent-error risks by less than 0.03 percentage points in
every interval.

The independent-error information matrix has rank 14/14, condition number 45,
and a sharply peaked likelihood around 28.04%; the boundary likelihood-ratio
diagnostic is 264,244. Refined maximum scaled gradients are $2.0\times10^{-4}$
for the exact piecewise fit, $2.1\times10^{-6}$ for the local-error fit, and
$2.5\times10^{-6}$ for the joint model. These are numerical diagnostics, not
substitutes for the standard errors reported in Table 7.

## Interpretation

- Piecewise hazards fit much better than constant or power-law hazards and
  improve the weighted-likelihood AIC by about 26,947 relative to the power
  law, while avoiding its implausibly long nonemployment tail.
- Exit falls after the first year of tenure. Entry falls sharply with time
  since work and then levels off rather than approaching zero.
- Allowing timegap contamination reduces estimated status error from about 10%
  to 2--3%, while tenure contamination remains close to 25%.
- The joint per-wave model fits best and leaves transition rates and duration
  profiles almost unchanged relative to the simpler contamination model.
- Only about 3 percentage points of timegap contradictions can be explained by
  an unobserved within-quarter employment spell; hidden resets are not the main
  source of the contamination estimate.
- The results identify inconsistency among status, tenure, timegap, and linked
  histories. They do not determine whether the underlying cause is respondent
  error, matching error, or another data problem.

## To do

- Add bootstrap or profile uncertainty and posterior-predictive checks for the
  joint status and duration distributions.
- Test alternative contamination densities, rounding and heaping rules, and
  tail cut-points while retaining transition-duration consistency.

# 8. No-error duration counterfactual (Appendix Table A8)

- Estimated flexible AR(1) and AR(2) transition models treating reported status
  changes as error-free.
- Simulated the duration clocks implied by those transitions.
- Compared simulated and observed tenure and timegap distributions, with
  transition-conditional graphs as the primary diagnostic.

\Needspace{18\baselineskip}

```{r no-error-clock-table, results='asis'}
clock <- read.csv(file.path(project_root,
  "diagnostics/output/no_error_duration_counterfactual/clock_inconsistency.csv"))
clock_labels <- c(
  "Entry followed by tenure >3 months" =
    "Apparent entry followed by tenure over three months",
  "Exit followed by timegap >3 months" =
    "Apparent exit followed by timegap over three months",
  "Employment continuation with inconsistent clock" =
    "Employment continuation with inconsistent tenure",
  "Nonemployment continuation with inconsistent clock" =
    "Nonemployment continuation with unreachable timegap")
cat("\\begin{table}[H]\n\\centering\n",
    "\\caption*{Appendix Table A8: No-error duration-clock contradictions}\n",
    "\\small\n\\begin{tabular}{@{}lrr@{}}\n\\toprule\n",
    " & AR(1) & AR(2) \\\\\n\\midrule\n", sep = "")
for (key in names(clock_labels)) {
  z <- clock[match(c("AR(1)", "AR(2)"),
    clock$model[clock$diagnostic == key]), ]
  # Re-select explicitly because match above indexes within a subset.
  z <- clock[clock$diagnostic == key, ]
  z <- z[match(c("AR(1)", "AR(2)"), z$model), ]
  cat(clock_labels[[key]], " & ",
      paste(sprintf("%.1f", 100 * z$inconsistent_share), collapse = " & "),
      " \\\\\n & ",
      paste(sprintf("(%.2f)", 100 * z$standard_error), collapse = " & "),
      " \\\\\n", sep = "")
}
cat("\\bottomrule\n\\end{tabular}\\\\[4pt]\n",
    "{\\footnotesize\\textit{Notes.} Entries are survey-weighted shares in percent. Parentheses contain person-clustered linearized standard errors. Each no-error model allows flexible duration dependence in observed transitions; destination-wave durations are held out and used for these clock-consistency diagnostics.}\n",
    "\\end{table}\n", sep = "")
```

**Graph accompanying Panel A. Clock contradictions under AR(1) and AR(2)**

```{r no-error-clock-graph, out.width='100%', fig.align='center'}
knitr::include_graphics(file.path(project_root,
  "diagnostics/output/no_error_duration_counterfactual/clock_inconsistency_shares.png"))
```

**Appendix Table A8, Panel B. Observed transition benchmarks**

| Rate | AR(1) (%) | AR(2) (%) |
|:--|--:|--:|
| Entry rate | 8.23 | 7.78 |
| Exit rate | 8.47 | 7.83 |
| Admissible tenure after apparent entry | 37.0 | 37.2 |

The flexible transition equations reproduce these observed entry and exit
rates. Among apparent exits in the AR(1) analysis, 26.0% report that they have
never worked, reinforcing the contradiction between status changes and the
reported duration history.

**Graphs accompanying Panel B. Empirical and simulated duration distributions**

```{r no-error-tenure-density-graph, out.width='100%', fig.align='center'}
knitr::include_graphics(file.path(project_root,
  "diagnostics/output/no_error_duration_counterfactual/tenure_empirical_vs_simulated_pdf.pdf"))
```

```{r no-error-timegap-graph, out.width='100%', fig.align='center'}
knitr::include_graphics(file.path(project_root,
  "diagnostics/output/no_error_duration_counterfactual/timegap_empirical_vs_simulated.png"))
```

## Interpretation

- Flexible no-error transition models reproduce average observed changes but
  cannot reconcile them with destination-wave duration reports.
- Most apparent entries and exits imply duration clocks that are impossible
  under a genuine transition during the interview interval.
- Marginal duration distributions obscure the problem because most respondents
  do not change reported state; transition-conditional comparisons are more
  persuasive.
- The diagnostic motivates jointly modelling status and duration consistency
  without claiming which reported variable or linked record is incorrect.

## To do

- Add simulation uncertainty bands to the empirical-versus-implied duration
  graphs and report transition-conditional comparisons prominently.
- Repeat the counterfactual under alternative matching panels and initial-clock
  assumptions.

# Reproducibility and remaining inference

- A result-by-result audit against `branch_changes_summary.Rmd` maps the
  baseline estimates to Table 1/A1, matching-rule results to Table 2,
  age/education-consistency results to Table 3/A2, covariate estimates and
  hazard distributions to Table 4/A4, AR(2) results to Table 5/A5, four-wave
  FMM results to Table 6/A6, tenure-contamination results to Table 7/A7, and
  the no-error duration benchmarks to A8. No numerical estimate reported in
  the long summary is intentionally omitted here; only detailed implementation
  narrative is abbreviated.
- Tables 1, 3, and 4 and Appendix Tables A1--A4 are rebuilt by build_tables.R.
- Table 2 is rebuilt by the matching-sensitivity script; the four-wave, tenure,
  and diagnostic scripts save the inputs for Tables 5--7 and Appendix Tables
  A5--A8 outside version control.
- Analytical standard errors are available for Tables 1--7 and their associated
  model estimates. Table 7 uses the full numerical observed-information matrix
  and delta method; Appendix Table A8 uses person-clustered linearization for
  its weighted descriptive shares.
- Table 7's numerical observed-information errors should still be checked with
  a person-level bootstrap or profile likelihood before final submission.
- A person-level bootstrap remains the most useful common inference robustness
  check across the paper.
