---
title: "Audit and development log: current-job tenure model"
subtitle: "Reported start dates, job changes, and duration contamination"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex: false
geometry: margin=1in
header-includes:
  - \usepackage{booktabs}
  - \usepackage{float}
  - \usepackage{longtable}
fontsize: 11pt
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
project_root <- normalizePath("..", winslash = "/", mustWork = TRUE)
audit_root <- file.path(project_root, "diagnostics", "output",
                        "job_start_dates")
comparison_file <- file.path(audit_root, "matching_comparison.csv")
if (!file.exists(comparison_file)) {
  stop("Run diagnostics/audit_job_start_dates.R for panels A, B, and C first")
}
audit <- read.csv(comparison_file, stringsAsFactors = FALSE)
job_result_root <- file.path(project_root, "EM-tenure", "output", "results",
                             "job_change")
job_comparison <- read.csv(file.path(job_result_root,
                                     "model_comparison_latest.csv"))
job_lr <- read.csv(file.path(job_result_root, "boundary_lr_latest.csv"))
matching_comparison <- read.csv(file.path(job_result_root,
  "matching_robustness", "comparison_latest.csv"))
validation_root <- file.path(job_result_root, "validation")
validation_starts <- read.csv(file.path(validation_root,
  "dispersed_starts_latest.csv"))
validation_interval <- read.csv(file.path(validation_root,
  "profile_interval_latest.csv"))
validation_recovery <- read.csv(file.path(validation_root,
  "recovery_summary_latest.csv"))
validation_score <- read.csv(file.path(validation_root,
  "large_sample_score_latest.csv"))
monthly_root <- file.path(project_root, "EM-tenure", "output", "results",
  "job_change_monthly")
monthly_comparison <- read.csv(file.path(monthly_root,
  "model_comparison_latest.csv"))
monthly_starts <- read.csv(file.path(monthly_root,
  "starting_values_latest.csv"))
monthly_lr <- read.csv(file.path(monthly_root, "boundary_lr_latest.csv"))
monthly_recovery <- read.csv(file.path(monthly_root, "validation",
  "recovery_latest.csv"))
monthly_recovery_grid <- read.csv(file.path(monthly_root, "recovery_grid",
  "summary_latest.csv"))
monthly_recovery_reps <- read.csv(file.path(monthly_root, "recovery_grid",
  "replications_latest.csv"))
monthly_inference <- read.csv(file.path(monthly_root, "inference",
  "analytical_se_latest.csv"))
monthly_hessian <- read.csv(file.path(monthly_root, "inference",
  "hessian_diagnostics_latest.csv"))
monthly_profile_interval <- read.csv(file.path(monthly_root, "inference",
  "profile_interval_latest.csv"))
monthly_matching <- read.csv(file.path(monthly_root, "matching_robustness",
  "comparison_latest.csv"))
monthly_matching_starts <- read.csv(file.path(monthly_root,
  "matching_robustness", "starting_values_latest.csv"))
persistent_root <- file.path(monthly_root, "persistent_reporting")
persistent_comparison <- read.csv(file.path(persistent_root,
  "model_comparison_latest.csv"))
persistent_starts <- read.csv(file.path(persistent_root,
  "starting_values_latest.csv"))
persistent_lr <- read.csv(file.path(persistent_root,"boundary_lr_latest.csv"))
persistent_recovery <- read.csv(file.path(persistent_root,"validation",
  "recovery_latest.csv"))
persistent_recovery_fits <- read.csv(file.path(persistent_root,"validation",
  "fits_latest.csv"))
persistent_profile <- read.csv(file.path(persistent_root,"validation",
  "conditional_profile_latest.csv"))
monthly_fits <- readRDS(file.path(monthly_root, "fits_latest.rds"))
persistent_fits <- readRDS(file.path(persistent_root, "fits_latest.rds"))
heaping_root <- file.path(monthly_root,"calendar_heaping")
heaping_comparison <- read.csv(file.path(heaping_root,
  "model_comparison_latest.csv"))
heaping_starts <- read.csv(file.path(heaping_root,
  "starting_values_latest.csv"))
heaping_lr <- read.csv(file.path(heaping_root,"boundary_lr_latest.csv"))
heaping_observed <- read.csv(file.path(heaping_root,
  "observed_january_share_latest.csv"))
heaping_recovery <- read.csv(file.path(heaping_root,"validation",
  "recovery_latest.csv"))
heaping_recovery_fits <- read.csv(file.path(heaping_root,"validation",
  "fits_latest.csv"))
heaping_inference <- read.csv(file.path(heaping_root,"inference",
  "analytical_se_latest.csv"))
heaping_fits <- readRDS(file.path(heaping_root,"fits_latest.rds"))
revision_root <- file.path(monthly_root,"calendar_revision")
revision_comparison <- read.csv(file.path(revision_root,
  "model_comparison_latest.csv"))
revision_starts <- read.csv(file.path(revision_root,
  "starting_values_latest.csv"))
revision_lr <- read.csv(file.path(revision_root,"boundary_lr_latest.csv"))
revision_observed <- read.csv(file.path(revision_root,
  "observed_revision_share_latest.csv"))
revision_recovery <- read.csv(file.path(revision_root,"validation",
  "recovery_latest.csv"))
revision_recovery_fits <- read.csv(file.path(revision_root,"validation",
  "fits_latest.csv"))
revision_fits <- readRDS(file.path(revision_root,"fits_latest.rds"))
revision_inference <- read.csv(file.path(revision_root,"inference",
  "analytical_se_latest.csv"))
revision_hessian <- read.csv(file.path(revision_root,"inference",
  "hessian_diagnostics_latest.csv"))
revision_profile_interval <- read.csv(file.path(revision_root,"inference",
  "profile","profile_interval_latest.csv"))
revision_predictive <- read.csv(file.path(revision_root,"diagnostics",
  "posterior_predictive_latest.csv"))
revision_decomposition <- read.csv(file.path(revision_root,"diagnostics",
  "likelihood_gain_by_pattern_latest.csv"))
clean_revision_root <- file.path(monthly_root,
  "calendar_clean_anchor_revision")
clean_revision_comparison <- read.csv(file.path(clean_revision_root,
  "model_comparison_latest.csv"))
clean_revision_starts <- read.csv(file.path(clean_revision_root,
  "starting_values_latest.csv"))
clean_revision_lr <- read.csv(file.path(clean_revision_root,
  "boundary_lr_latest.csv"))
clean_revision_recovery <- read.csv(file.path(clean_revision_root,
  "validation","recovery_latest.csv"))
clean_revision_recovery_fits <- read.csv(file.path(clean_revision_root,
  "validation","fits_latest.csv"))
clean_revision_inference <- read.csv(file.path(clean_revision_root,
  "inference","analytical_se_latest.csv"))
clean_revision_inference_full <- readRDS(file.path(clean_revision_root,
  "inference","analytical_inference_latest.rds"))
clean_revision_hessian <- read.csv(file.path(clean_revision_root,
  "inference","hessian_diagnostics_latest.csv"))
clean_revision_predictive <- read.csv(file.path(clean_revision_root,
  "diagnostics","posterior_predictive_latest.csv"))
clean_revision_decomposition <- read.csv(file.path(clean_revision_root,
  "diagnostics","likelihood_gain_by_pattern_latest.csv"))
start_month_root <- file.path(monthly_root,"calendar_start_month_baseline")
start_month_comparison <- read.csv(file.path(start_month_root,
  "model_comparison_latest.csv"))
start_month_starts <- read.csv(file.path(start_month_root,
  "starting_values_latest.csv"))
start_month_distribution <- read.csv(file.path(start_month_root,
  "start_month_distribution_latest.csv"))
start_month_lr <- read.csv(file.path(start_month_root,"lr_latest.csv"))
start_month_predictive <- read.csv(file.path(start_month_root,"diagnostics",
  "posterior_predictive_latest.csv"))
start_month_decomposition <- read.csv(file.path(start_month_root,"diagnostics",
  "likelihood_gain_by_start_month_latest.csv"))
start_month_recovery_fits <- read.csv(file.path(start_month_root,"validation",
  "fits_latest.csv"))
start_month_recovery <- read.csv(file.path(start_month_root,"validation",
  "month_recovery_latest.csv"))
exact_anchor_root <- file.path(monthly_root,
  "calendar_exact_anchor_retention")
exact_anchor_comparison <- read.csv(file.path(exact_anchor_root,
  "model_comparison_latest.csv"))
exact_anchor_lr <- read.csv(file.path(exact_anchor_root,"lr_latest.csv"))
exact_anchor_predictive <- read.csv(file.path(exact_anchor_root,"diagnostics",
  "posterior_predictive_latest.csv"))
exact_anchor_decomposition <- read.csv(file.path(exact_anchor_root,
  "diagnostics","likelihood_gain_by_continuation_pattern_latest.csv"))
exact_anchor_recovery <- read.csv(file.path(exact_anchor_root,"validation",
  "recovery_summary_latest.csv"))
exact_anchor_refinement <- read.csv(file.path(exact_anchor_root,
  "refinement_summary_latest.csv"))
exact_anchor_joint_fits <- read.csv(file.path(exact_anchor_root,
  "validation_joint","fit_summary_latest.csv"))
exact_anchor_joint_recovery <- read.csv(file.path(exact_anchor_root,
  "validation_joint","recovery_summary_latest.csv"))
exact_anchor_opg <- read.csv(file.path(exact_anchor_root,"inference_opg",
  "analytical_se_latest.csv"))
exact_anchor_opg_diagnostics <- read.csv(file.path(exact_anchor_root,
  "inference_opg","information_diagnostics_latest.csv"))
exact_anchor_reporting_cor <- as.matrix(read.csv(file.path(exact_anchor_root,
  "inference_opg","reporting_probability_correlations_latest.csv"),
  row.names=1,check.names=FALSE))
local_anchor_root <- file.path(monthly_root,"calendar_local_anchor_revision")
local_anchor_comparison <- read.csv(file.path(local_anchor_root,
  "model_comparison_latest.csv"))
local_anchor_refinement <- read.csv(file.path(local_anchor_root,
  "refinement_summary_latest.csv"))
local_anchor_lr <- read.csv(file.path(local_anchor_root,"lr_latest.csv"))
local_anchor_predictive <- read.csv(file.path(local_anchor_root,"diagnostics",
  "posterior_predictive_latest.csv"))
local_anchor_decomposition <- read.csv(file.path(local_anchor_root,
  "diagnostics","likelihood_gain_by_continuation_pattern_latest.csv"))
local_anchor_curvature <- read.csv(file.path(local_anchor_root,"diagnostics",
  "conditional_curvature_latest.csv"))
shared_reliability_root <- file.path(monthly_root,
  "shared_duration_reliability")
shared_reliability_comparison <- read.csv(file.path(shared_reliability_root,
  "model_comparison_latest.csv"))
shared_reliability_refinement <- read.csv(file.path(shared_reliability_root,
  "refinement_summary_latest.csv"))
shared_reliability_components <- read.csv(file.path(shared_reliability_root,
  "component_probabilities_latest.csv"))
shared_reliability_posterior <- read.csv(file.path(shared_reliability_root,
  "posterior_unreliable_summary_latest.csv"))
shared_reliability_lr <- read.csv(file.path(shared_reliability_root,
  "lr_latest.csv"))
shared_reliability_predictive <- read.csv(file.path(shared_reliability_root,
  "diagnostics","posterior_predictive_latest.csv"))
shared_reliability_decomposition <- read.csv(file.path(shared_reliability_root,
  "diagnostics","likelihood_gain_by_duration_pattern_latest.csv"))
separate_reliability_root <- file.path(monthly_root,
  "separate_duration_reliability")
separate_reliability_comparison <- read.csv(file.path(
  separate_reliability_root,"model_comparison_latest.csv"))
separate_reliability_refinement <- read.csv(file.path(
  separate_reliability_root,"refinement_summary_latest.csv"))
separate_reliability_components <- read.csv(file.path(
  separate_reliability_root,"component_probabilities_latest.csv"))
separate_reliability_posterior <- read.csv(file.path(
  separate_reliability_root,"posterior_unreliable_summary_latest.csv"))
separate_reliability_lr <- read.csv(file.path(separate_reliability_root,
  "lr_latest.csv"))
separate_reliability_predictive <- read.csv(file.path(
  separate_reliability_root,"diagnostics","posterior_predictive_latest.csv"))
separate_reliability_decomposition <- read.csv(file.path(
  separate_reliability_root,"diagnostics",
  "likelihood_gain_by_duration_pattern_latest.csv"))
separate_reliability_joint <- read.csv(file.path(
  separate_reliability_root,"joint_continuation_latest.csv"))
separate_reliability_common_polish <- read.csv(file.path(
  separate_reliability_root,"common_polish_latest.csv"))
separate_reliability_recovery <- read.csv(file.path(
  separate_reliability_root,"validation","recovery_draws_latest.csv"))
separate_reliability_recovery_starts <- read.csv(file.path(
  separate_reliability_root,"validation","multistart_summary_latest.csv"))
separate_reliability_inference <- read.csv(file.path(
  separate_reliability_root,"inference_opg","analytical_se_latest.csv"))
separate_reliability_information <- read.csv(file.path(
  separate_reliability_root,"inference_opg",
  "information_diagnostics_latest.csv"))
separate_reliability_profile <- read.csv(file.path(
  separate_reliability_root,"profile_difference",
  "profile_likelihood_latest.csv"))
separate_reliability_profile_interval <- read.csv(file.path(
  separate_reliability_root,"profile_difference",
  "profile_interval_latest.csv"))

format_metric <- function(label, percent = TRUE, digits = 1L) {
  z <- audit[audit$quantity == label, c("panel", "estimate")]
  z$panel <- sub("df_qlfs_", "Panel ", z$panel, fixed = TRUE)
  if (percent) z$estimate <- 100 * z$estimate
  setNames(as.list(round(z$estimate[match(c("Panel A", "Panel B", "Panel C"),
                                         z$panel)], digits)),
           c("Panel A", "Panel B", "Panel C"))
}
```

# Purpose and status

This document is a reproducible audit and development log for improvements to
the tenure-contamination models reported in Table 7. It is deliberately
separate from the branch-wide summaries. Its purposes are to:

1. document the provenance and construction of the duration variables;
2. record descriptive evidence used to select measurement assumptions;
3. distinguish observed evidence from latent-model interpretation;
4. list every new estimation script and the checks applied to it; and
5. preserve point estimates, diagnostics, and model comparisons as the model
   is extended incrementally.

The start-date audit, job-reset extension, and discrete-month likelihood repair
are implemented and estimated separately from Table 7. The continuous-density
version fails full-model recovery and is retained only as a diagnostic. The
monthly probability-mass version passes repeated recovery checks, converges
from dispersed starts, has regular local inference, and remains stable under
stricter matching panels. It is the current candidate model, but the remaining
measurement-process extensions below should be assessed before it replaces the
paper's Table 7 specification.

# 1. Variable provenance and reconstruction

The upstream mapped QLFS file contains `yearstart` (year commenced working),
`monthstart` (month commenced working), survey `year`, quarterly `wave`, and
the derived tenure variable `timeago`. The matched panels retain `period` and
`tenure`, but not the start-year and start-month components.

For survey period (p), define the nominal interview year and month by

\[
Y(p)=2008+\left\lfloor\frac{p-1}{4}\right\rfloor,
\qquad
M(p)=3\{(p-1)\bmod 4+1\}.
\]

Thus the nominal interview months are March, June, September, and December.
For every upstream record with valid start-year and start-month fields,

\[
\text{tenure months}
=12\{Y(p)-Y^{start}\}+M(p)-M^{start}.
\]

Consequently, the reported start month can be recovered without an upstream
merge as

\[
R_{it}=12Y(p_{it})+M(p_{it})-g_{it},
\]

where (g_{it}) is tenure in months. This is a reported current-job start
month, not necessarily the date of the worker's most recent entry into
employment. The exact interview day is unavailable.

## Audit cautions

- A changed start date may reflect a new job, inconsistent reporting, or an
  incorrect record link.
- A date strictly after the previous nominal interview is strong evidence of
  a job change, but dates in the previous interview month are ambiguous because
  interview days are unavailable.
- The matching algorithms affect which records are linked. Comparisons across
  panels A, B, and C are therefore essential.
- January concentration can reflect genuine seasonality, rounding, or a
  reporting/coding convention. The available labels do not identify which.

# 2. Descriptive start-date audit

The audit uses the paper's age range (18--55), all available consecutive waves,
survey weights, and the three alternative matching panels. It classifies each
consecutive employed pair as an unchanged start date, a plausible job reset, an
ambiguous boundary-month reset, or a revision to a date that predates the
previous interview.

```{r headline-table, results='asis'}
metrics <- list(
  c("Same reported start month", "Same reported start month"),
  c("Plausible job-to-job reset", "Plausible job-to-job reset"),
  c("Boundary-month reset", "Boundary-month reset"),
  c("Any revised pre-existing start date", "Revised pre-existing date"),
  c("Revisions that are exact whole-year shifts", "Whole-year shift among revisions"),
  c("Revised date repeated in the next employed wave", "Revised date repeated next wave"),
  c("Plausible reset repeated in the next employed wave", "Plausible reset repeated next wave"),
  c("January share of reported start months", "Reported start month is January")
)
tab <- do.call(rbind, lapply(metrics, function(m) {
  values <- if (m[1] == "Boundary-month reset") {
    # This measure is stored in the pair-category output, not the summary.
    vapply(c("A", "B", "C"), function(p) {
      f <- file.path(audit_root, paste0("df_qlfs_", p),
                     "consecutive_employed_pair_categories.csv")
      z <- read.csv(f, stringsAsFactors = FALSE)
      100 * z$weighted_share[z$pair == "All" &
        z$category == "Boundary-month reset"]
    }, numeric(1))
  } else {
    unlist(format_metric(m[1]))
  }
  data.frame(Measure = m[2], `Panel A` = values[1], `Panel B` = values[2],
             `Panel C` = values[3], check.names = FALSE)
}))
rownames(tab) <- NULL
knitr::kable(tab, format = "latex", booktabs = TRUE, digits = 1,
             caption = "Start-date audit: survey-weighted percentages",
             linesep = "", longtable = FALSE, row.names = FALSE,
             table.envir = "table", position = "H")
```

```{r audit-figure, fig.cap='Reported current-job start dates and their revisions, matching panel A.', out.width='100%', fig.pos='H'}
knitr::include_graphics(file.path(audit_root, "df_qlfs_A",
                                  "job_start_date_audit.png"))
```

## Interpretation

- Most consecutive employed reports preserve exactly the same implied start
  month, and an unchanged date is repeated at the following employed wave in
  roughly nine out of ten cases.
- Stricter matching increases the unchanged share and reduces revisions, so
  matching quality contributes to the observed inconsistencies.
- Large revisions remain common under panel C. They should not be attributed
  exclusively to either respondent reporting or record linkage.
- Exact whole-year shifts and the January concentration favor a calendar-date
  reporting model over a smooth local-error distribution for calculated
  tenure.
- A non-negligible group has a reported job start after the previous interview.
  Current-job tenure can therefore reset without an intervening observed
  nonemployment state.

# 3. Audit of the existing Table 7 likelihood

The preferred current specification combines piecewise duration-dependent
entry and exit hazards, symmetric employment-status misclassification, gross
marginal tenure contamination, and joint per-wave marginal timegap
contamination. Its central duration restriction is that uncontaminated reports
must follow the deterministic clock implied by a latent employment or
nonemployment spell.

For tenure, this treats a maximal latent employment spell as one current-job
spell. That assumption is too strong when a worker can move directly between
jobs. A legitimate job-start reset is then forced into one of three incorrect
interpretations: a contaminated tenure report, a misclassified employment
state, or a latent exit and re-entry.

The local-plus-gross tenure extension does not repair this problem. Its local
component assigns independent bounded deviations around a single employment
clock; it does not allow the underlying current-job clock itself to restart.

# 4. First improved model: a job-change mixture

## Minimal defensible extension

For a latent employed-to-employed transition, introduce a job-continuity event
(J_{it}):

\[
J_{it}=0 \quad\text{same current job},
\qquad
J_{it}=1 \quad\text{new job while remaining employed}.
\]

Let

\[
P(J_{it}=1\mid H_{i,t-1}=H_{it}=1)=q.
\]

Conditional on (J_{it}=0), the uncontaminated tenure clock advances by the
quarterly interval. Conditional on (J_{it}=1), the current-job tenure clock
restarts within the interval. Initially, conditional event time can be taken
as uniform within the quarter; a continuous-time job-change hazard can replace
this approximation later.

This adds only one parameter, (q), and leaves unchanged:

- the latent employment-status transition process;
- employment-status misclassification;
- the piecewise exit and entry hazards;
- tenure-report contamination; and
- timegap contamination.

It is preferable to deterministically splitting spells whenever a date resets.
The mixture allows the likelihood to weigh job change against reporting/linking
inconsistency and remains coherent when reported employment status is itself
misclassified.

## Identification and required comparisons

The job-change probability is identified primarily by short destination tenure
within a latent employed-to-employed transition and by subsequent persistence
of the new start date. The first estimates must be compared on the same sample:

1. current preferred piecewise model;
2. preferred model plus (q);
3. model 2 with (q=0), verifying exact likelihood nesting; and
4. panels A, B, and C, verifying sensitivity to matching quality.

The estimate should not be interpreted as a pure job-to-job transition rate
until identification simulations show that it can be separated from tenure
contamination and status misclassification.

## First estimates

The joint maximum-likelihood fit converged. It estimates a quarterly
employed-to-employed current-job reset probability of
`r sprintf("%.2f", 100 * job_comparison$job_change_prob[2])`%, and the
survey-weighted posterior mean reset probability is
`r sprintf("%.2f", 100 * job_comparison$posterior_job_change_rate[2])`%.
The near equality of these two quantities is an internal consistency check.

```{r job-change-comparison}
shown <- job_comparison[, c("model", "loglik", "AIC", "job_change_prob",
  "status_misclassification", "tenure_contamination",
  "timegap_contamination", "weighted_entry", "weighted_exit")]
names(shown) <- c("Model", "Log likelihood", "AIC", "Job reset",
  "Status error", "Tenure error", "Timegap error", "Entry", "Exit")
shown[, 4:9] <- lapply(shown[, 4:9], function(x) sprintf("%.3f", x))
shown[, 2:3] <- lapply(shown[, 2:3], function(x) format(round(x, 1),
  big.mark = ",", scientific = FALSE))
knitr::kable(shown, booktabs = TRUE,
  caption = "Preferred duration model and the one-parameter job-change extension")
```

The log likelihood rises by
`r format(round(diff(job_comparison$loglik), 1), big.mark = ",")` and AIC falls
by `r format(round(-diff(job_comparison$AIC), 1), big.mark = ",")`. A boundary
likelihood-ratio statistic of `r round(job_lr$LR, 1)` provides overwhelming
evidence against (q=0). Substantively, allowing legitimate current-job resets
reduces the estimated gross tenure-contamination probability from
`r sprintf("%.1f", 100 * job_comparison$tenure_contamination[1])`% to
`r sprintf("%.1f", 100 * job_comparison$tenure_contamination[2])`%. It therefore
explains an important subset of tenure discontinuities, but most tenure
inconsistency is still assigned to the broader contamination mechanism. The
weighted entry and exit rates move only modestly, so the extension improves the
measurement model without overturning the employment-flow results.

## Matching-panel robustness

Panel A remains the preferred and main dataset. Panels B and C change only the
record-matching restrictions and diagnose whether the reset mechanism is
disproportionately supported by weaker links.

```{r matching-model-comparison}
match_show <- matching_comparison[, c("panel", "observations",
  "job_change_prob", "posterior_job_change_rate",
  "status_misclassification", "tenure_contamination",
  "timegap_contamination", "weighted_entry", "weighted_exit")]
names(match_show) <- c("Panel", "N", "Job reset", "Posterior reset",
  "Status error", "Tenure error", "Timegap error", "Entry", "Exit")
match_show$Panel <- paste("Panel", match_show$Panel)
match_show$N <- format(match_show$N, big.mark=",", scientific=FALSE)
match_show[, 3:9] <- lapply(match_show[, 3:9], function(x)
  sprintf("%.3f", x))
knitr::kable(match_show, booktabs=TRUE,
  caption="Job-change model under alternative matching restrictions")
```

The reset probability is nearly unchanged between Panels A and B
(`r sprintf('%.2f', 100 * matching_comparison$job_change_prob[1])`% versus
`r sprintf('%.2f', 100 * matching_comparison$job_change_prob[2])`%) and declines
to `r sprintf('%.2f', 100 * matching_comparison$job_change_prob[3])`% in Panel
C. By contrast, status misclassification falls from
`r sprintf('%.2f', 100 * matching_comparison$status_misclassification[1])`% to
`r sprintf('%.2f', 100 * matching_comparison$status_misclassification[3])`%,
and tenure contamination falls from
`r sprintf('%.1f', 100 * matching_comparison$tenure_contamination[1])`% to
`r sprintf('%.1f', 100 * matching_comparison$tenure_contamination[3])`%.
Stricter matching therefore removes substantial measurement inconsistency but
does not eliminate the current-job reset signal. The modest decline under C
also cautions against interpreting every reset as a genuine job-to-job move.

# 5. Numerical validation and recovery

## Optimizer stability

The high-reset dispersed start converges to the established optimum, and a
tighter refinement changes the reset estimate only from
`r sprintf('%.4f', 100 * validation_starts$job_change_prob[1])`% to
`r sprintf('%.4f', 100 * validation_interval$estimate)`%. A deliberately low
start converges to an inferior local mode, with a log likelihood lower by
`r format(round(validation_starts$loglik[1]-validation_starts$loglik[2], 1), big.mark=',')`.
Multiple starts and likelihood-based selection are therefore required.

Reoptimizing all 14 nuisance parameters at seven fixed reset probabilities
gives a provisional 95% profile-likelihood interval of
[`r sprintf('%.3f', 100 * validation_interval$lower)`%,
`r sprintf('%.3f', 100 * validation_interval$upper)`%]. The local quadratic
fit has (R^2=`r sprintf('%.5f', validation_interval$quadratic_R2)`). This
interval describes the implemented likelihood; it is not yet suitable for the
paper because the recovery check below fails.

## Full-likelihood recovery

Three samples of 20,000 observations were simulated from the intended model
and re-estimated from both low and high reset starts. The reset and tenure-error
parameters are not recovered without systematic bias.

```{r recovery-table}
recover_show <- validation_recovery[, c("parameter", "truth",
  "mean_estimate", "bias")]
recover_show$parameter <- c("Job reset", "Status error", "Tenure error",
  "Timegap error")
names(recover_show) <- c("Parameter", "Truth", "Mean estimate", "Bias")
recover_show[, 2:4] <- lapply(recover_show[, 2:4], function(x)
  sprintf("%.3f", x))
knitr::kable(recover_show, booktabs=TRUE,
  caption="Full-likelihood recovery diagnostic")
```

In particular, a true reset probability of
`r sprintf('%.3f', 100 * validation_recovery$truth[1])`% is estimated at
`r sprintf('%.3f', 100 * validation_recovery$mean_estimate[1])`% on average,
while tenure contamination rises from
`r sprintf('%.1f', 100 * validation_recovery$truth[3])`% to
`r sprintf('%.1f', 100 * validation_recovery$mean_estimate[3])`%. This is not
just small-sample variation: with 100,000 simulated observations, the numerical
likelihood scores at the generating values remain strongly nonzero
(`r round(validation_score$q_score, 1)` for resets and
`r round(validation_score$eps_score, 1)` for tenure contamination).

The leading cause is a units problem in the continuous-density approximation.
Mixture branches contain different numbers of independent tenure-density
evaluations, and hence have different physical units, but the current
likelihood adds them directly. Clean clock restrictions also create singular
components that are mixed with ordinary continuous densities. Since tenure is
constructed from a reported start **month**, the defensible repair is to model
tenure in discrete months: replace marginal densities by monthly interval
probabilities and replace the continuous within-quarter reset density by a
probability over possible start months. The job-change result should not be
promoted to Table 7 until that repair restores simulation recovery.

## Discrete-month repair and re-estimation

The repaired likelihood treats a reported tenure of (m) months as the mass

\[
P(m/12 \leq D < (m+1)/12),
\]

rather than evaluating a continuous density at (m/12). Clean tenure clocks
advance by exactly three months between quarterly interviews. A latent spell
that begins between waves, or a current-job reset within a persistent
employment spell, starts at 0, 1, or 2 reported months. This puts every branch
of the mixture on the same probability scale. The ingestion also retains
2,753 genuine zero-month reports and drops two records that previously received
an artificial quarter-month floor because no valid tenure report was available.

```{r monthly-comparison}
monthly_show <- monthly_comparison[, c("model", "loglik", "AIC",
  "job_change_prob", "status_misclassification", "tenure_contamination",
  "timegap_contamination", "entry_rate", "exit_rate")]
names(monthly_show) <- c("Model", "Log likelihood", "AIC", "Job reset",
  "Status error", "Tenure error", "Timegap error", "Entry", "Exit")
monthly_show[, 4:9] <- lapply(monthly_show[, 4:9], function(x)
  sprintf("%.3f", x))
monthly_show[, 2:3] <- lapply(monthly_show[, 2:3], function(x)
  format(round(x, 1), big.mark=",", scientific=FALSE))
knitr::kable(monthly_show, booktabs=TRUE,
  caption="Panel A discrete-month tenure likelihood")
```

The preferred reset fit estimates a quarterly current-job reset probability of
`r sprintf('%.3f', 100*monthly_comparison$job_change_prob[2])`%, almost exactly
equal to its posterior mean of
`r sprintf('%.3f', 100*monthly_comparison$posterior_job_change_rate[2])`%.
Allowing resets raises the log likelihood by
`r format(round(diff(monthly_comparison$loglik), 1), big.mark=',')` and gives a
boundary LR statistic of `r round(monthly_lr$LR, 1)`. Relative to the no-reset
monthly model, it lowers tenure contamination only modestly, from
`r sprintf('%.2f', 100*monthly_comparison$tenure_contamination[1])`% to
`r sprintf('%.2f', 100*monthly_comparison$tenure_contamination[2])`%.

All three reset starts converge near the same solution: their reset estimates
range from `r sprintf('%.3f',100*min(monthly_starts$job_change_prob))`% to
`r sprintf('%.3f',100*max(monthly_starts$job_change_prob))`%, and their log
likelihoods differ by less than
`r sprintf('%.1f',diff(range(monthly_starts$loglik)))`.

Six additional full-likelihood recovery fits use three independent samples at
each of 10,000 and 40,000 observations. All converge, and high dispersed starts
at both sample sizes do not improve the likelihood.

```{r monthly-recovery-grid}
keep <- monthly_recovery_grid$parameter %in%
  c("job_change_prob", "pi", "eps", "eps_d")
rec_show <- monthly_recovery_grid[keep,
  c("n", "parameter", "truth", "mean", "bias", "rmse")]
rec_show$parameter <- c("Job reset", "Status error", "Tenure error",
  "Timegap error")[match(rec_show$parameter,
    c("job_change_prob", "pi", "eps", "eps_d"))]
names(rec_show) <- c("N", "Parameter", "Truth", "Mean", "Bias", "RMSE")
rec_show$N <- format(rec_show$N, big.mark=",", scientific=FALSE)
rec_show[, 3:6] <- lapply(100*rec_show[, 3:6], function(x)
  sprintf("%.3f", x))
rownames(rec_show) <- NULL
knitr::kable(rec_show, booktabs=TRUE,
  caption="Repeated full-likelihood recovery (percentage points)")
```

At 40,000 observations, mean bias is only
`r sprintf('%+.3f',100*monthly_recovery_grid$bias[monthly_recovery_grid$n==40000 & monthly_recovery_grid$parameter=='job_change_prob'])`
percentage points for resets,
`r sprintf('%+.3f',100*monthly_recovery_grid$bias[monthly_recovery_grid$n==40000 & monthly_recovery_grid$parameter=='pi'])`
for status error,
`r sprintf('%+.3f',100*monthly_recovery_grid$bias[monthly_recovery_grid$n==40000 & monthly_recovery_grid$parameter=='eps'])`
for tenure contamination, and
`r sprintf('%+.3f',100*monthly_recovery_grid$bias[monthly_recovery_grid$n==40000 & monthly_recovery_grid$parameter=='eps_d'])`
for timegap contamination. RMSE also falls with sample size. The likelihood
scores at the generating parameters change sign across replications rather
than showing the systematic direction found under the continuous model. These
checks support accepting the discrete likelihood repair. A larger simulation
would still be required to assess confidence-interval coverage directly.

## Analytical and profile inference

The full 15-parameter numerical observed-information matrix is positive
definite and full rank. Its minimum eigenvalue is
`r sprintf('%.4g',monthly_hessian$minimum_eigenvalue)`, and its condition number
is `r sprintf('%.1f',monthly_hessian$condition_number)`. Delta-method standard
errors therefore propagate the joint uncertainty in all transition hazards,
contamination probabilities, initial employment, and job resets.

```{r monthly-inference}
order_q <- c("entry_rate", "exit_rate", "status_misclassification",
  "tenure_contamination", "timegap_contamination", "initial_employment",
  "job_change_prob")
inf_show <- monthly_inference[match(order_q, monthly_inference$quantity),
  c("quantity", "estimate", "se", "lower", "upper")]
inf_show$quantity <- c("Entry", "Exit", "Status error", "Tenure error",
  "Timegap error", "Initial employment", "Job reset")
names(inf_show) <- c("Quantity", "Estimate", "SE", "Lower 95%", "Upper 95%")
inf_show[,2:5] <- lapply(100*inf_show[,2:5], function(x) sprintf("%.3f",x))
knitr::kable(inf_show, booktabs=TRUE,
  caption="Discrete-month model: analytical standard errors (percent)")
```

The reset estimate is
`r sprintf('%.3f',100*monthly_profile_interval$estimate)`% with analytical SE
`r sprintf('%.3f',100*monthly_profile_interval$analytical_se)` percentage
points. Reoptimizing all 14 nuisance parameters at each fixed reset probability
gives a 95% profile interval of
[`r sprintf('%.3f',100*monthly_profile_interval$lower)`%,
`r sprintf('%.3f',100*monthly_profile_interval$upper)`%]. It is nearly
identical to the Wald interval, and the profile quadratic has
(R^2=`r sprintf('%.5f',monthly_profile_interval$quadratic_R2)`). Thus profile
asymmetry is negligible around the interior optimum.

The discrete model materially changes the substantive allocation. Compared
with the continuous reset fit, the reset probability falls from
`r sprintf('%.2f',100*job_comparison$job_change_prob[2])`% to
`r sprintf('%.2f',100*monthly_comparison$job_change_prob[2])`%, status
misclassification rises from
`r sprintf('%.2f',100*job_comparison$status_misclassification[2])`% to
`r sprintf('%.2f',100*monthly_comparison$status_misclassification[2])`%, and
tenure contamination falls from
`r sprintf('%.1f',100*job_comparison$tenure_contamination[2])`% to
`r sprintf('%.1f',100*monthly_comparison$tenure_contamination[2])`%. Entry also
falls, while exit changes less. Log likelihoods across the continuous and
monthly models are not comparable because one uses densities and the other
uses discrete probability masses.

\newpage

## Matching-panel sensitivity of the corrected model

The identical corrected monthly specification was re-estimated on Panels B and
C. Panel A remains the main sample; the stricter panels are sensitivity checks,
not alternative preferred datasets.

```{r monthly-matching-comparison}
mm_show <- monthly_matching[, c("panel", "observations", "job_change_prob",
  "posterior_job_change_rate", "status_misclassification",
  "tenure_contamination", "timegap_contamination", "entry_rate", "exit_rate")]
names(mm_show) <- c("Panel", "N", "Job reset", "Posterior reset",
  "Status error", "Tenure error", "Timegap error", "Entry", "Exit")
mm_show$N <- format(mm_show$N, big.mark=",", scientific=FALSE)
mm_show[,3:9] <- lapply(100*mm_show[,3:9], function(x) sprintf("%.3f", x))
knitr::kable(mm_show, booktabs=TRUE,
  caption="Corrected discrete-month model under alternative matching restrictions (percent)")
```

The reset estimate declines monotonically from
`r sprintf('%.3f',100*monthly_matching$job_change_prob[1])`% in Panel A to
`r sprintf('%.3f',100*monthly_matching$job_change_prob[2])`% in B and
`r sprintf('%.3f',100*monthly_matching$job_change_prob[3])`% in C. Status
misclassification falls from
`r sprintf('%.2f',100*monthly_matching$status_misclassification[1])`% to
`r sprintf('%.2f',100*monthly_matching$status_misclassification[3])`%, tenure
contamination from
`r sprintf('%.1f',100*monthly_matching$tenure_contamination[1])`% to
`r sprintf('%.1f',100*monthly_matching$tenure_contamination[3])`%, and timegap
contamination from
`r sprintf('%.1f',100*monthly_matching$timegap_contamination[1])`% to
`r sprintf('%.1f',100*monthly_matching$timegap_contamination[3])`%. Entry is
nearly unchanged in B and modestly lower in C; exit also falls slightly. Thus
stricter matching removes some apparent transition and duration inconsistency,
but it does not eliminate the estimated current-job reset process.

All six B/C optimizations converged normally. Within each panel, low, main, and
high reset starts reach log likelihoods agreeing within
`r sprintf('%.2f', max(tapply(monthly_matching_starts$loglik,
  monthly_matching_starts$panel, function(x) diff(range(x)))[c('B','C')]))`
log-likelihood units; reset estimates span only
`r sprintf('%.4f',100*max(tapply(monthly_matching_starts$job_change_prob,
  monthly_matching_starts$panel, function(x) diff(range(x)))[c('B','C')]))`
percentage points within either Panel B or C.

\newpage

## Reporting-process extensions

The five nested Panel A versions can be compared coefficient by coefficient.
The first rules out job resets and persistent gross reports. The second adds a
job-reset probability but treats gross start-date reports as independent over
time. The third estimates whether a gross report retains the preceding
reported start-date anchor. The fourth instead allows gross reports to follow
the empirically prominent January-heaping pattern. The fifth adds a correlated
whole-year revision of the previously reported start date.

```{r nested-model-coefficients}
no_reset <- monthly_fits$no_reset$params
independent <- monthly_fits$reset$params
persistent <- persistent_fits$persistent$params
heaping <- heaping_fits$heaping$params
revision <- revision_fits$revision$params

duration_labels <- c("0-3 months", "3-12 months", "1-3 years",
                     "3-5 years", "5+ years")
coefficient_labels <- c(
  "Initial employment probability",
  "Status misclassification probability",
  "Tenure contamination probability",
  "Timegap contamination probability",
  paste0("Exit hazard: ", duration_labels),
  paste0("Entry hazard: ", duration_labels),
  "Job-reset probability",
  "Gross-report persistence probability",
  "January-heaping probability",
  "Whole-year revision probability"
)

extract_coefficients <- function(x, persistence = 0, heaping = 0,
                                 revision = 0) {
  c(x$alpha, x$pi, x$eps, x$eps_d, x$lambda_g, x$lambda_d,
    x$job_change_prob, persistence, heaping, revision)
}
fmt_coefficient <- function(x) sprintf("%.4f", x)

coef_table <- data.frame(
  Coefficient = coefficient_labels,
  `No reset` = fmt_coefficient(extract_coefficients(no_reset)),
  `Independent gross` = fmt_coefficient(extract_coefficients(independent)),
  `Persistent gross` = fmt_coefficient(extract_coefficients(
    persistent, persistent$tenure_report_persistence)),
  `January heap` = fmt_coefficient(extract_coefficients(
    heaping, heaping=heaping$tenure_heaping_prob)),
  `January + year revision` = fmt_coefficient(extract_coefficients(
    revision, heaping=revision$tenure_heaping_prob,
    revision=revision$tenure_year_revision_prob)),
  check.names = FALSE
)
reset_row <- match("Job-reset probability",coef_table$Coefficient)
persistence_row <- match("Gross-report persistence probability",
  coef_table$Coefficient)
heaping_row <- match("January-heaping probability",coef_table$Coefficient)
revision_row <- match("Whole-year revision probability",
  coef_table$Coefficient)
coef_table[reset_row, "No reset"] <- "0 (fixed)"
coef_table[persistence_row,
  c("No reset", "Independent gross", "January heap",
    "January + year revision")] <-
  "0 (fixed)"
coef_table[persistence_row, "Persistent gross"] <- "0 (boundary)"
coef_table[heaping_row,
  c("No reset", "Independent gross", "Persistent gross")] <- "0 (fixed)"
coef_table[revision_row,
  c("No reset", "Independent gross", "Persistent gross",
    "January heap")] <- "0 (fixed)"

fit_rows <- data.frame(
  Coefficient = c("Log likelihood", "AIC", "Number of parameters"),
  `No reset` = c(
    format(round(monthly_fits$no_reset$loglik, 1), big.mark = ",",
           scientific = FALSE),
    format(round(monthly_fits$comparison$AIC[1], 1), big.mark = ",",
           scientific = FALSE),
    monthly_fits$comparison$parameters[1]
  ),
  `Independent gross` = c(
    format(round(monthly_fits$reset$loglik, 1), big.mark = ",",
           scientific = FALSE),
    format(round(monthly_fits$comparison$AIC[2], 1), big.mark = ",",
           scientific = FALSE),
    monthly_fits$comparison$parameters[2]
  ),
  `Persistent gross` = c(
    format(round(persistent_fits$persistent$loglik, 1), big.mark = ",",
           scientific = FALSE),
    format(round(persistent_fits$comparison$AIC[2], 1), big.mark = ",",
           scientific = FALSE),
    persistent_fits$comparison$parameters[2]
  ),
  `January heap` = c(
    format(round(heaping_fits$heaping$loglik, 1), big.mark = ",",
           scientific = FALSE),
    format(round(heaping_fits$comparison$AIC[2], 1), big.mark = ",",
           scientific = FALSE),
    heaping_fits$comparison$parameters[2]
  ),
  `January + year revision` = c(
    format(round(revision_fits$revision$loglik, 1), big.mark = ",",
           scientific = FALSE),
    format(round(revision_fits$comparison$AIC[2], 1), big.mark = ",",
           scientific = FALSE),
    revision_fits$comparison$parameters[2]
  ),
  check.names = FALSE
)

knitr::kable(rbind(coef_table, fit_rows), booktabs = TRUE,
  align = c("l", "r", "r", "r", "r", "r"),
  caption = paste("Natural-scale coefficients across nested corrected",
                  "discrete-month models (Panel A)"))
```

All probability and hazard coefficients are shown on the natural zero-to-one
scale. Exit and entry hazards are annualized piecewise-constant hazards for the
indicated duration band. The job-reset coefficient is the probability within
an observed quarterly interval; gross-report persistence is the conditional
probability that a contaminated report retains the previous reported anchor.
January heaping is the conditional probability that a gross report is drawn
from the marginal duration law restricted to an implied January start date.
Whole-year revision is the conditional probability that an eligible new gross
report preserves the previous reported start month but changes its year.
“Fixed” denotes a nesting restriction, while “boundary” denotes an estimated
coefficient whose constrained MLE is zero.

### Persistent reported-start-date extension

The first reporting-process extension allows a consecutive gross tenure report
within the same latent job to retain the preceding reported start-date anchor
with probability (rho). Otherwise it draws a new marginal anchor. Clean tenure
reports remain exactly consistent with the true employment spell, job changes
start a new segment, and the timegap process is unchanged. The independent
gross-report model is nested exactly at (rho=0).

```{r persistent-reporting-comparison}
rate_names <- c("entry_rate","exit_rate","status_misclassification",
  "tenure_contamination","timegap_contamination","job_change_prob",
  "tenure_report_persistence")
pr_show <- data.frame(
  Quantity=c("Entry","Exit","Status error","Tenure error","Timegap error",
    "Job reset","Persistence","Log likelihood","AIC"),
  Independent=c(sprintf("%.3f",100*unlist(persistent_comparison[1,rate_names])),
    format(round(persistent_comparison$loglik[1],1),big.mark=",",
      scientific=FALSE),format(round(persistent_comparison$AIC[1],1),
      big.mark=",",scientific=FALSE)),
  Persistent=c(sprintf("%.3f",100*unlist(persistent_comparison[2,rate_names])),
    format(round(persistent_comparison$loglik[2],1),big.mark=",",
      scientific=FALSE),format(round(persistent_comparison$AIC[2],1),
      big.mark=",",scientific=FALSE)),check.names=FALSE)
names(pr_show)[2:3] <- c("Independent gross reports","Persistent gross anchor")
knitr::kable(pr_show,booktabs=TRUE,
  caption="Persistent reported-start-date extension (rates in percent)")
```

The constrained MLE is exactly at (rho=0). Consequently, every other reported
rate is unchanged, the likelihood-ratio statistic is
`r sprintf('%.1f',persistent_lr$LR)`, and adding the unused parameter raises
AIC by two. The converged middle start reaches the numerical lower bound
(`r sprintf('%.6f',100*persistent_starts$tenure_report_persistence[2])`%),
while the high start returns to the same basin from (rho=0.90) but reaches the
iteration cap only after matching its likelihood closely. A low start finds an
inferior local solution approximately
`r format(round(max(persistent_starts$loglik)-persistent_starts$loglik[1],1),big.mark=',')`
log-likelihood units below the boundary solution. The exact nested fit is
therefore retained rather than a slightly inferior numerical approximation to
zero.

This boundary result is not caused by an intrinsically unidentified
persistence parameter. In a 12,000-record simulation with true (rho=0.60), the
conditional likelihood peaks at
`r sprintf('%.3f',persistent_profile$rho[which.max(persistent_profile$loglik)])`
on the reported grid, and the two full-likelihood starts converge to
`r paste(sprintf('%.3f',persistent_recovery_fits$tenure_report_persistence),collapse=' and ')`.
The likelihood can therefore distinguish substantial persistence, although
the best estimate is downward biased by
`r sprintf('%.3f',-persistent_recovery$bias[persistent_recovery$parameter=='tenure_report_persistence'])`
in this single moderate-size simulation. More replications would be needed for
coverage or bias correction.

Substantively, the data do not support this particular one-parameter mechanism
in addition to the clean-clock and independent gross-report branches. This is
not evidence that respondents never repeat start dates: clock-consistent
sequences are already assigned to the clean branch, and a persistently wrong
anchor can be observationally similar for jobs that began before the panel.
The next reporting extension should therefore focus on structured start-month
heaping or a persistent offset tied to jobs whose start is bounded by an
observed entry or reset, rather than adding further unrestricted persistence.

\newpage

### January-heaped gross-report extension

The next extension uses the strongest calendar pattern in the audit. For an
interview in nominal month (M), a tenure report of (m) months implies a January
start when $m \bmod 12=M-1$. Conditional on a gross tenure report, the model
now uses

\[
f^{gross}(m\mid M)=(1-h)f(m)+h
\frac{\mathbf{1}\{m\bmod 12=M-1\}f(m)}
{P_f(m\bmod 12=M-1)}.
\]

Thus the heaped component preserves the estimated marginal duration shape but
conditions it on an implied January start. It is normalized exactly, adds one
parameter, and nests the independent-gross model at (h=0). Nominal interview
months are retained when the likelihood cells are collapsed.

```{r calendar-heaping-comparison}
heap_rates <- c("entry_rate","exit_rate","status_misclassification",
  "tenure_contamination","timegap_contamination","job_change_prob",
  "gross_january_heaping_prob","prior_heaped_tenure_report_prob")
heap_show <- data.frame(
  Quantity=c("Entry","Exit","Status error","Tenure error","Timegap error",
    "Job reset","January heaping, conditional on gross",
    "Gross-and-heaped tenure report","Log likelihood","AIC"),
  Independent=c(sprintf("%.3f",100*unlist(
    heaping_comparison[1,heap_rates])),
    format(round(heaping_comparison$loglik[1],1),big.mark=",",
      scientific=FALSE),format(round(heaping_comparison$AIC[1],1),
      big.mark=",",scientific=FALSE)),
  January_heaping=c(sprintf("%.3f",100*unlist(
    heaping_comparison[2,heap_rates])),
    format(round(heaping_comparison$loglik[2],1),big.mark=",",
      scientific=FALSE),format(round(heaping_comparison$AIC[2],1),
      big.mark=",",scientific=FALSE)),check.names=FALSE)
names(heap_show)[2:3] <- c("Independent gross","January-heaped gross")
knitr::kable(heap_show,booktabs=TRUE,
  caption="January-heaped gross-report extension (rates in percent)")
```

All three starts converge normally. They estimate (h) between
`r sprintf('%.5f',min(heaping_starts$gross_january_heaping_prob))` and
`r sprintf('%.5f',max(heaping_starts$gross_january_heaping_prob))`, while their
log likelihoods differ by only
`r sprintf('%.3f',diff(range(heaping_starts$loglik)))`. The preferred estimate
is `r sprintf('%.2f',100*heaping_comparison$gross_january_heaping_prob[2])`%:
conditional on a gross tenure report, roughly one quarter of reports are
assigned to the January-conditioned component. Because gross reporting itself
has probability `r sprintf('%.2f',100*heaping_comparison$tenure_contamination[2])`%,
their prior product is
`r sprintf('%.2f',100*heaping_comparison$prior_heaped_tenure_report_prob[2])`%
of otherwise correctly classified employed-wave tenure reports.

Fit improves sharply: the log likelihood rises by
`r format(round(diff(heaping_comparison$loglik),1),big.mark=',')`, AIC falls by
`r format(round(-diff(heaping_comparison$AIC),1),big.mark=',')`, and the
boundary likelihood-ratio statistic is
`r format(round(heaping_lr$LR,1),big.mark=',')`. Entry changes from
`r sprintf('%.3f',100*heaping_comparison$entry_rate[1])`% to
`r sprintf('%.3f',100*heaping_comparison$entry_rate[2])`%, exit from
`r sprintf('%.3f',100*heaping_comparison$exit_rate[1])`% to
`r sprintf('%.3f',100*heaping_comparison$exit_rate[2])`%, and the job-reset
probability from `r sprintf('%.3f',100*heaping_comparison$job_change_prob[1])`%
to `r sprintf('%.3f',100*heaping_comparison$job_change_prob[2])`%. The main
transition conclusions are therefore stable, while the reporting likelihood
fits the calendar pattern much better.

The observed survey-weighted January share averages
`r sprintf('%.1f',100*heaping_observed$observed_january_share[2])`% across the
three waves. This is not directly comparable to (h): clean reports and the
ordinary marginal gross distribution can also imply January starts. Nor does
the model establish that January reports are errors; genuine seasonal hiring
or a coding convention could generate the same concentration. The parameter
should be described as a structured reporting component unless external
validation separates those mechanisms.

```{r calendar-heaping-recovery}
heap_recovery_show <- heaping_recovery
heap_recovery_show$parameter <- c("January heaping","Job reset",
  "Status error","Tenure error","Timegap error")
for (nm in c("truth","estimate","bias"))
  heap_recovery_show[[nm]] <- sprintf("%.4f",heap_recovery_show[[nm]])
names(heap_recovery_show) <- c("Parameter","Truth","Estimate","Bias")
knitr::kable(heap_recovery_show,booktabs=TRUE,
  caption="Calendar-heaping recovery check (6,000 simulated records)")
```

In the recovery exercise, the truth is (h=0.40). The conditional optimum is
0.385, and two dispersed joint starts converge to estimates agreeing within
`r sprintf('%.5f',diff(range(heaping_recovery_fits$tenure_heaping_prob)))`.
The small downward error in (h) does not materially distort the reset, status,
or timegap-error probabilities. This supports identification of the added
component, although repeated simulations would be required for bias and
coverage statements.

```{r calendar-heaping-inference}
heap_se_names <- c("entry_rate","exit_rate","status_misclassification",
  "tenure_contamination","timegap_contamination","job_change_prob",
  "gross_january_heaping_prob")
heap_se_labels <- c("Entry","Exit","Status error","Tenure error",
  "Timegap error","Job reset","January heaping, conditional on gross")
heap_se_show <- heaping_inference[match(heap_se_names,
  heaping_inference$quantity),c("estimate","se","lower","upper")]
heap_se_show <- 100*heap_se_show
heap_se_show <- data.frame(Quantity=heap_se_labels,
  Estimate=sprintf("%.3f",heap_se_show$estimate),
  `Analytical SE`=sprintf("%.3f",heap_se_show$se),
  `95% lower`=sprintf("%.3f",heap_se_show$lower),
  `95% upper`=sprintf("%.3f",heap_se_show$upper),check.names=FALSE)
knitr::kable(heap_se_show,booktabs=TRUE,
  caption="Observed-information and delta-method inference for January heaping (percent)")
```

The full 16-parameter observed-information matrix is positive definite and
full rank. Its condition number is 97.2, providing no numerical indication of
a locally unidentified heaping parameter. These are analytical
observed-information standard errors, not bootstrap standard errors.

### Correlated whole-year date revisions

The next nested component targets the second prominent calendar pattern. Let
$m_{t-1}$ be the previous gross tenure report, let $c=m_{t-1}+3\Delta$ be its
exact continuation after $\Delta$ panel intervals, and let $S_{t-1}$ be its
implied reported start month. A whole-year revision draws the current report
from the marginal duration law conditional on preserving $S_{t-1}$ while
excluding (c):

\[
 f^{year}(m_t\mid S_{t-1},c)=
 \frac{\mathbf{1}\{S(m_t)=S_{t-1},\ m_t\ne c\}f(m_t)}
 {P_f\{S(m)=S_{t-1}\}-f(c)}.
\]

For eligible consecutive gross reports within the same latent job, the model
uses the ordinary January-heaping gross law with probability $1-\omega$ and
this different-year law with probability $\omega$. It is normalized exactly
and nests the preceding model at $\omega=0$. This is distinct from the failed
persistence extension: persistence repeats the exact reported date, whereas
this component changes the year but preserves the month.

```{r calendar-revision-comparison}
revision_rates <- c("entry_rate","exit_rate","status_misclassification",
  "tenure_contamination","timegap_contamination","job_change_prob",
  "gross_january_heaping_prob","whole_year_revision_prob")
revision_show <- data.frame(
  Quantity=c("Entry","Exit","Status error","Tenure error","Timegap error",
    "Job reset","January heaping, conditional on gross",
    "Whole-year revision, conditional on eligible gross pair",
    "Log likelihood","AIC"),
  January_heaping=c(sprintf("%.3f",100*unlist(
    revision_comparison[1,revision_rates])),
    format(round(revision_comparison$loglik[1],1),big.mark=",",
      scientific=FALSE),format(round(revision_comparison$AIC[1],1),
      big.mark=",",scientific=FALSE)),
  Whole_year_revision=c(sprintf("%.3f",100*unlist(
    revision_comparison[2,revision_rates])),
    format(round(revision_comparison$loglik[2],1),big.mark=",",
      scientific=FALSE),format(round(revision_comparison$AIC[2],1),
      big.mark=",",scientific=FALSE)),check.names=FALSE)
names(revision_show)[2:3] <- c("January heaping",
  "January + whole-year revisions")
knitr::kable(revision_show,booktabs=TRUE,
  caption="Correlated whole-year date-revision extension (rates in percent)")
```

The preferred high start converges normally at
$\omega=`r sprintf('%.4f',revision_comparison$whole_year_revision_prob[2])`$.
The central start reaches
`r sprintf('%.4f',revision_starts$whole_year_revision_prob[revision_starts$start=='main'])`
but is
`r sprintf('%.1f',max(revision_starts$loglik)-revision_starts$loglik[revision_starts$start=='main'])`
log-likelihood points lower. The deliberately low 0.03 start stops in an
inferior local mode at its initial revision probability, more than
`r format(round(max(revision_starts$loglik)-min(revision_starts$loglik),1),big.mark=',')`
points below the preferred solution. It is therefore a numerical warning, not
a competing substantive estimate: multistart likelihood selection is needed.
Relative to January heaping alone, the preferred log likelihood changes by
`r format(round(diff(revision_comparison$loglik),1),big.mark=',')`, AIC changes
by `r format(round(diff(revision_comparison$AIC),1),big.mark=',')`, and the
boundary LR statistic is `r format(round(revision_lr$LR,1),big.mark=',')`.
Entry moves from `r sprintf('%.3f',100*revision_comparison$entry_rate[1])`% to
`r sprintf('%.3f',100*revision_comparison$entry_rate[2])`%, exit from
`r sprintf('%.3f',100*revision_comparison$exit_rate[1])`% to
`r sprintf('%.3f',100*revision_comparison$exit_rate[2])`%, and tenure
contamination from
`r sprintf('%.3f',100*revision_comparison$tenure_contamination[1])`% to
`r sprintf('%.3f',100*revision_comparison$tenure_contamination[2])`%.

The raw survey-weighted share of employed-report pairs whose implied start
month is unchanged but whose year differs ranges from
`r sprintf('%.2f',100*min(revision_observed$observed_whole_year_revision_share))`%
to
`r sprintf('%.2f',100*max(revision_observed$observed_whole_year_revision_share))`%.
This is not directly $\omega$: the observed share mixes clean clocks, job
changes, gross reports, and status error, whereas $\omega$ is conditional on
an eligible latent gross-report pair.

```{r calendar-revision-recovery}
revision_recovery_show <- revision_recovery
revision_recovery_show$parameter <- c("Whole-year revision","January heaping",
  "Job reset","Status error","Tenure error","Timegap error")
for (nm in c("truth","estimate","bias"))
  revision_recovery_show[[nm]] <- sprintf("%.4f",revision_recovery_show[[nm]])
names(revision_recovery_show) <- c("Parameter","Truth","Estimate","Bias")
knitr::kable(revision_recovery_show,booktabs=TRUE,
  caption="Whole-year revision recovery check (6,000 simulated records)")
```

With a generating value of $\omega=0.45$, the conditional likelihood optimum
is 0.456. Two dispersed joint starts recover 0.443 and 0.458. The best fit also
recovers January heaping within 0.0005 and tenure contamination within 0.0040
of their generating values, supporting separate identification of the two
calendar components in this diagnostic sample.

```{r calendar-revision-inference}
revision_se_names <- c("entry_rate","exit_rate","status_misclassification",
  "tenure_contamination","timegap_contamination","job_change_prob",
  "gross_january_heaping_prob","whole_year_revision_prob")
revision_se_labels <- c("Entry","Exit","Status error","Tenure error",
  "Timegap error","Job reset","January heaping, conditional on gross",
  "Whole-year revision, conditional on eligible gross pair")
revision_se_show <- revision_inference[match(revision_se_names,
  revision_inference$quantity),c("estimate","se","lower","upper")]
revision_se_show <- 100*revision_se_show
revision_se_show <- data.frame(Quantity=revision_se_labels,
  Estimate=sprintf("%.3f",revision_se_show$estimate),
  `Analytical SE`=sprintf("%.3f",revision_se_show$se),
  `95% lower`=sprintf("%.3f",revision_se_show$lower),
  `95% upper`=sprintf("%.3f",revision_se_show$upper),check.names=FALSE)
knitr::kable(revision_se_show,booktabs=TRUE,
  caption="Observed-information and delta-method inference for the whole-year revision model (percent)")
```

The full 17-parameter observed-information matrix is positive definite and
full rank. Its condition number is
`r sprintf('%.1f',revision_hessian$condition_number)`. The analytical standard
error of $\omega$ is
`r sprintf('%.3f',100*revision_inference$se[revision_inference$quantity=='whole_year_revision_prob'])`
percentage points. The nuisance-adjusted profile interval is
`r sprintf('%.2f',100*revision_profile_interval$profile_lower)`--
`r sprintf('%.2f',100*revision_profile_interval$profile_upper)`%, compared with
the analytical interval
`r sprintf('%.2f',100*revision_profile_interval$analytical_lower)`--
`r sprintf('%.2f',100*revision_profile_interval$analytical_upper)`%. Thus the
preferred local solution has regular local curvature; this does not by itself
establish that the reporting process is correctly specified.

### Parametric predictive checks and source of fit improvement

The fitted-parameter checks below compare observed Panel A moments with 40
simulated samples of 50,000 records. They assess replication of the observed
calendar patterns conditional on the fitted parameters; they do not integrate
over parameter uncertainty.

```{r calendar-revision-predictive}
predictive_metrics <- c("all_employed_reports_january_start",
  "consecutive_pairs_exact","consecutive_pairs_whole_year",
  "consecutive_pairs_other")
predictive_labels <- c("January start month, all employed reports",
  "Exact continuation, consecutive employed pairs",
  "Whole-year revision, consecutive employed pairs",
  "Other revision, consecutive employed pairs")
pred <- revision_predictive[revision_predictive$metric %in% predictive_metrics,]
pred$Measure <- predictive_labels[match(pred$metric,predictive_metrics)]
pred$Model <- ifelse(pred$model=="January heaping","January only",
  "January + year revision")
pred_show <- reshape(pred[,c("Measure","Model","predicted_mean")],
  idvar="Measure",timevar="Model",direction="wide")
names(pred_show) <- sub("predicted_mean\\.","",names(pred_show))
pred_show$Observed <- pred$observed[match(pred_show$Measure,pred$Measure)]
pred_show <- pred_show[match(predictive_labels,pred_show$Measure),]
pred_show <- pred_show[,c("Measure","January only",
  "January + year revision","Observed")]
for (nm in names(pred_show)[-1]) pred_show[[nm]] <-
  sprintf("%.2f",100*pred_show[[nm]])
knitr::kable(pred_show,booktabs=TRUE,row.names=FALSE,
  caption="Fitted-parameter predictive checks for observed calendar patterns (percent)")
```

The revision component moves the predicted whole-year-revision share in the
intended direction, from
`r sprintf('%.2f',100*pred$predicted_mean[pred$Model=='January only' & pred$metric=='consecutive_pairs_whole_year'])`%
to
`r sprintf('%.2f',100*pred$predicted_mean[pred$Model=='January + year revision' & pred$metric=='consecutive_pairs_whole_year'])`%,
but the observed share is
`r sprintf('%.2f',100*pred$observed[pred$metric=='consecutive_pairs_whole_year'][1])`%.
None of the four observed moments lies within the corresponding 95% simulation
envelope. In particular, predicted January starts remain near 13.6% compared
with 26.4% observed. A conditional January probability of about 24% among
gross reports cannot reproduce the aggregate month distribution when only
about 22% of tenure reports use that gross component. Genuine hiring
seasonality, calendar-specific clean reports, or both may therefore be
missing from the baseline start-month law.

```{r calendar-revision-decomposition}
decomp_show <- revision_decomposition
decomp_show$pattern <- c(
  "Exact continuation"="Exact continuation",
  "At least one exact continuation"="Exact continuation",
  "At least one whole-year revision"="At least one whole-year revision",
  "Fewer than two employed reports"="Fewer than two employed reports",
  "Other multiple employed reports"="Other multiple-employed pattern"
)[decomp_show$pattern]
decomp_show <- data.frame(Pattern=unname(decomp_show$pattern),
  `Population share (%)`=sprintf("%.2f",100*decomp_show$population_share),
  `Log-likelihood gain`=sprintf("%.1f",decomp_show$loglik_gain),
  `Share of net gain (%)`=sprintf("%.1f",100*decomp_show$gain_share),
  check.names=FALSE)
knitr::kable(decomp_show,booktabs=TRUE,
  caption="Decomposition of the whole-year extension's log-likelihood gain")
```

All of the positive fit improvement comes from the 6.16% of weighted records
with at least one observed whole-year date revision. Their gain exceeds the
net gain because the extension loses about 3,083 log-likelihood points on
other patterns. This concentration is substantively informative, but the
offsetting deterioration and failed predictive checks show that the current
correlated branch is too narrow. It requires both adjacent reports to be
gross. The following extension implements the first repair; start-month
seasonality remains a separate next step.

\clearpage

### Revisions of preceding clean reports

Let $\eta$ denote the probability that a gross current tenure report draws
from the same normalized whole-year revision law when its preceding observed
tenure report was clean. The existing gross-to-gross probability $\omega$
remains separate. Thus $\eta=0$ reproduces the preceding 17-parameter model
exactly, while the 18-parameter extension distinguishes which type of report
supplies the preceding date anchor. Missing tenure reports do not erase an
anchor within the same latent job, but a latent job change does.

```{r clean-anchor-revision-comparison}
clean_rates <- c("entry_rate","exit_rate","status_misclassification",
  "tenure_contamination","timegap_contamination","job_change_prob",
  "gross_january_heaping_prob","gross_anchor_revision_prob",
  "clean_anchor_revision_prob")
clean_show <- data.frame(
  Quantity=c("Entry","Exit","Status error","Tenure error","Timegap error",
    "Job reset","January heaping, conditional on gross",
    "Gross-anchor whole-year revision",
    "Clean-anchor whole-year revision","Log likelihood","AIC"),
  Nested=c(sprintf("%.3f",100*unlist(
    clean_revision_comparison[1,clean_rates])),
    format(round(clean_revision_comparison$loglik[1],1),big.mark=",",
      scientific=FALSE),format(round(clean_revision_comparison$AIC[1],1),
      big.mark=",",scientific=FALSE)),
  Extended=c(sprintf("%.3f",100*unlist(
    clean_revision_comparison[2,clean_rates])),
    format(round(clean_revision_comparison$loglik[2],1),big.mark=",",
      scientific=FALSE),format(round(clean_revision_comparison$AIC[2],1),
      big.mark=",",scientific=FALSE)),check.names=FALSE)
names(clean_show)[2:3] <- c("Gross-anchor revisions only",
  "Gross + clean-anchor revisions")
knitr::kable(clean_show,booktabs=TRUE,
  caption="Clean-anchor whole-year revision extension (rates in percent)")
```

The extension improves the log likelihood by
`r format(round(diff(clean_revision_comparison$loglik),1),big.mark=',')`
points and lowers AIC by
`r format(round(-diff(clean_revision_comparison$AIC),1),big.mark=',')`.
The boundary LR statistic is
`r format(round(clean_revision_lr$LR,1),big.mark=',')`.
Starts at 3%, 25%, and 65% all converge near
`r sprintf('%.2f',100*clean_revision_comparison$clean_anchor_revision_prob[2])`%
for $\eta$. Allowing clean anchors reduces $\omega$ from
`r sprintf('%.2f',100*clean_revision_comparison$gross_anchor_revision_prob[1])`%
to
`r sprintf('%.2f',100*clean_revision_comparison$gross_anchor_revision_prob[2])`%,
while entry, exit, status error, and the two marginal contamination rates move
by less than 0.04 percentage points. The new channel therefore reallocates
the reporting process without materially changing the employment-transition
interpretation.

\clearpage

```{r clean-anchor-revision-inference}
clean_se_names <- c("entry_rate","exit_rate","status_misclassification",
  "tenure_contamination","timegap_contamination","job_change_prob",
  "gross_january_heaping_prob","whole_year_revision_prob",
  "clean_anchor_revision_prob")
clean_se_labels <- c("Entry","Exit","Status error","Tenure error",
  "Timegap error","Job reset","January heaping, conditional on gross",
  "Gross-anchor whole-year revision","Clean-anchor whole-year revision")
clean_se_show <- clean_revision_inference[match(clean_se_names,
  clean_revision_inference$quantity),c("estimate","se","lower","upper")]
clean_se_show <- 100*clean_se_show
clean_se_show <- data.frame(Quantity=clean_se_labels,
  Estimate=sprintf("%.3f",clean_se_show$estimate),
  `Analytical SE`=sprintf("%.3f",clean_se_show$se),
  `95% lower`=sprintf("%.3f",clean_se_show$lower),
  `95% upper`=sprintf("%.3f",clean_se_show$upper),check.names=FALSE)
knitr::kable(clean_se_show,booktabs=TRUE,
  caption="Observed-information inference for the clean-anchor extension (percent)")
```

The 18-parameter observed-information matrix is positive definite and full
rank, with condition number
`r sprintf('%.1f',clean_revision_hessian$condition_number)`. The analytical
standard error of $\eta$ is
`r sprintf('%.3f',100*clean_revision_inference$se[clean_revision_inference$quantity=='clean_anchor_revision_prob'])`
percentage points. The delta-method correlation between $\eta$ and $\omega$
is
`r sprintf('%.3f',cov2cor(clean_revision_inference_full$vcov_quantities)['clean_anchor_revision_prob','whole_year_revision_prob'])`.
The channels are related but the local information does not indicate a nearly
singular reallocation.

```{r clean-anchor-revision-recovery}
clean_recovery_show <- clean_revision_recovery
clean_recovery_show$parameter <- c("Clean-anchor revision",
  "Gross-anchor revision","January heaping","Tenure error","Job reset",
  "Status error")
for (nm in c("truth","estimate","bias"))
  clean_recovery_show[[nm]] <- sprintf("%.4f",clean_recovery_show[[nm]])
names(clean_recovery_show) <- c("Parameter","Truth","Estimate","Bias")
knitr::kable(clean_recovery_show,booktabs=TRUE,
  caption="Clean-anchor revision recovery check (8,000 simulated records)")
```

Both recovery starts converge to the same likelihood and estimate $\eta$ at
about 0.325 when its generating value is 0.35. The corresponding estimate of
$\omega$ is about 0.244 rather than 0.204. This demonstrates joint
recoverability, while the offsetting finite-sample biases caution against
interpreting the two probabilities too finely without a larger repeated
recovery design.

```{r clean-anchor-revision-predictive}
clean_pred_metrics <- c("all_employed_reports_january_start",
  "consecutive_pairs_exact","consecutive_pairs_whole_year",
  "consecutive_pairs_other")
clean_pred_labels <- c("January start month, all employed reports",
  "Exact continuation, consecutive employed pairs",
  "Whole-year revision, consecutive employed pairs",
  "Other revision, consecutive employed pairs")
cp <- clean_revision_predictive[
  clean_revision_predictive$metric %in% clean_pred_metrics,]
cp$Measure <- clean_pred_labels[match(cp$metric,clean_pred_metrics)]
cp$Model <- ifelse(cp$model=="Gross-anchor revisions only",
  "Gross-anchor only","Gross + clean-anchor")
cp_show <- reshape(cp[,c("Measure","Model","predicted_mean")],
  idvar="Measure",timevar="Model",direction="wide")
names(cp_show) <- sub("predicted_mean\\.","",names(cp_show))
cp_show$Observed <- cp$observed[match(cp_show$Measure,cp$Measure)]
cp_show <- cp_show[match(clean_pred_labels,cp_show$Measure),
  c("Measure","Gross-anchor only","Gross + clean-anchor","Observed")]
for (nm in names(cp_show)[-1]) cp_show[[nm]] <-
  sprintf("%.2f",100*cp_show[[nm]])
knitr::kable(cp_show,booktabs=TRUE,row.names=FALSE,
  caption="Predictive checks after allowing clean-anchor revisions (percent)")
```

The targeted whole-year-revision prediction rises from 4.10% to 5.79%, closer
to the observed 7.01%, while other revisions fall from 35.58% to about 33.8%
against 28.29% observed. The checks nevertheless still fail: exact
continuations remain too low and January starts remain near 13.4% rather than
26.4%. The extension explains an important part of the date-revision pattern,
but it does not repair the missing baseline calendar distribution.

```{r clean-anchor-revision-decomposition}
clean_decomp_show <- data.frame(
  Pattern=clean_revision_decomposition$pattern,
  `Population share (%)`=sprintf("%.2f",
    100*clean_revision_decomposition$population_share),
  `Log-likelihood gain`=sprintf("%.1f",
    clean_revision_decomposition$loglik_gain),
  `Share of net gain (%)`=sprintf("%.1f",
    100*clean_revision_decomposition$gain_share),check.names=FALSE)
knitr::kable(clean_decomp_show,booktabs=TRUE,
  caption="Likelihood gain from allowing clean-anchor revisions")
```

As intended, the positive gain is concentrated among the 6.16% of weighted
records with a whole-year revision. Their gain is partly offset by lower fit
for exact continuations and other multiple-employed patterns. The next model
should therefore estimate a genuine start-month distribution shared by clean
and gross reports before adding further dependence or AR(2) dynamics.

\clearpage

### Flexible baseline start-month distribution

The next extension replaces the implicit uniform distribution of job-start
months with 12 probabilities, estimated through 11 logits relative to
December. The duration mass is multiplied by the probability of its implied
calendar start month and renormalized separately for each interview month.
The same distribution enters clean and gross marginal tenure reports and the
0--2 month support for latent entries and job resets. Uniform probabilities
reproduce the preceding 18-parameter likelihood exactly. The gross-only
January-heaping branch and the two whole-year revision probabilities remain
separate.

```{r start-month-baseline-comparison}
start_rates <- c("entry_rate","exit_rate","status_misclassification",
  "tenure_contamination","timegap_contamination","job_change_prob",
  "gross_january_heaping_prob","gross_anchor_revision_prob",
  "clean_anchor_revision_prob")
start_show <- data.frame(
  Quantity=c("Entry","Exit","Status error","Tenure error","Timegap error",
    "Job reset","January heaping, conditional on gross",
    "Gross-anchor whole-year revision",
    "Clean-anchor whole-year revision","Log likelihood","AIC"),
  Uniform=c(sprintf("%.3f",100*unlist(
    start_month_comparison[1,start_rates])),
    format(round(start_month_comparison$loglik[1],1),big.mark=",",
      scientific=FALSE),format(round(start_month_comparison$AIC[1],1),
      big.mark=",",scientific=FALSE)),
  Flexible=c(sprintf("%.3f",100*unlist(
    start_month_comparison[2,start_rates])),
    format(round(start_month_comparison$loglik[2],1),big.mark=",",
      scientific=FALSE),format(round(start_month_comparison$AIC[2],1),
      big.mark=",",scientific=FALSE)),check.names=FALSE)
names(start_show)[2:3] <- c("Uniform baseline","Flexible baseline")
knitr::kable(start_show,booktabs=TRUE,
  caption="Shared flexible baseline start-month distribution (rates in percent)")
```

The flexible distribution improves the log likelihood by
`r format(round(diff(start_month_comparison$loglik),1),big.mark=',')` points
and lowers AIC by
`r format(round(-diff(start_month_comparison$AIC),1),big.mark=',')` despite 11
additional parameters. The LR statistic for equal month probabilities is
`r format(round(start_month_lr$LR,1),big.mark=',')` on 11 degrees of freedom.
Entry, exit, status error, and tenure and timegap contamination each move by
less than 0.2 percentage points. The transition interpretation is therefore
stable. In contrast, the gross-specific January-heaping probability falls
from
`r sprintf('%.2f',100*start_month_comparison$gross_january_heaping_prob[1])`%
to approximately zero. January concentration belongs in the shared baseline
calendar distribution rather than in an error process used only by gross
reports.

\clearpage

```{r start-month-baseline-distribution}
month_show <- data.frame(Month=start_month_distribution$month,
  `Observed reports`=sprintf("%.2f",
    100*start_month_distribution$observed_report_share),
  `Estimated baseline`=sprintf("%.2f",
    100*start_month_distribution$estimated_baseline),
  Uniform=sprintf("%.2f",100*start_month_distribution$uniform),
  check.names=FALSE)
knitr::kable(month_show,booktabs=TRUE,
  caption="Observed and estimated job-start month distributions (percent)")
```

The estimated baseline puts
`r sprintf('%.2f',100*start_month_distribution$estimated_baseline[1])`% of
start dates in January, followed by March, February, and April. December has
only
`r sprintf('%.2f',100*start_month_distribution$estimated_baseline[12])`%.
This strongly non-uniform pattern explains why the earlier gross-only heaping
component could not reproduce the aggregate data: most January concentration
is common to clean and gross reports.

```{r start-month-baseline-predictive}
pred_metrics <- c("consecutive_pairs_exact","consecutive_pairs_whole_year",
  "consecutive_pairs_other","start_month_1")
pred_labels <- c("Exact continuation","Whole-year revision","Other revision",
  "January start month")
sp <- start_month_predictive[start_month_predictive$model=="Flexible baseline" &
  start_month_predictive$metric %in% pred_metrics,]
sp <- sp[match(pred_metrics,sp$metric),]
pred_show <- data.frame(Measure=pred_labels,
  Predicted=sprintf("%.2f",100*sp$predicted_mean),
  Observed=sprintf("%.2f",100*sp$observed),
  `Within 95% simulation envelope`=ifelse(sp$within_95,"Yes","No"),
  check.names=FALSE)
knitr::kable(pred_show,booktabs=TRUE,
  caption="Predictive checks for the flexible baseline (percent)")
```

The month distribution is repaired: 11 of 12 observed month shares lie inside
their fitted-parameter simulation envelopes. Whole-year revisions rise to
6.89%, close to 7.01% observed. However, exact continuations remain too low
(60.7% versus 64.7%) and other revisions too high (32.5% versus 28.3%). The
remaining misspecification is therefore serial dependence in reporting, not
baseline calendar seasonality.

\clearpage

```{r start-month-baseline-recovery}
best_recovery <- start_month_recovery_fits[
  which.max(start_month_recovery_fits$loglik),]
recovery_show <- data.frame(
  Quantity=c("January probability","Month-distribution RMSE",
    "Gross-anchor revision","Clean-anchor revision","Tenure error",
    "Status error"),
  Truth=c(start_month_recovery$truth[1],NA,0.065764500,0.104707950,
    0.218171810,0.030195474),
  Estimate=c(start_month_recovery$estimate[1],best_recovery$month_rmse,
    best_recovery$gross_anchor_revision_prob,
    best_recovery$clean_anchor_revision_prob,
    best_recovery$tenure_contamination,
    best_recovery$status_misclassification))
recovery_show$Truth <- ifelse(is.na(recovery_show$Truth),"--",
  sprintf("%.4f",recovery_show$Truth))
recovery_show$Estimate <- sprintf("%.4f",recovery_show$Estimate)
knitr::kable(recovery_show,booktabs=TRUE,
  caption="Joint recovery check (2,000 simulated records)")
```

Uniform and reversed conditional starts converge to essentially the same
likelihood, and the safeguarded joint optimizer converges before a capped
central refinement. The 12-month distribution is recovered with RMSE about
0.0044. The two revision channels show more small-sample reallocation, so
their sum and predictive implications remain more reliable than a fine
comparison of their separate probabilities. Full observed-information
inference for all 29 parameters remains to be completed; the near-zero
gross-only January-heaping estimate is a boundary quantity and should not be
given a regular Wald interpretation.

\clearpage

### Exact reported-start-date retention

The next nested extension lets a gross current tenure report retain the
preceding reported job-start date exactly, regardless of whether the preceding
report was clean or gross. Let $\rho$ denote this retention probability. If
the anchor is not retained, the report follows the existing whole-year
revision versus marginal-redraw mixture. Thus $\rho=0$ reproduces the flexible
baseline likelihood exactly and the two revision parameters retain their
conditional interpretation.

```{r exact-anchor-comparison}
anchor_rates <- c("entry_rate","exit_rate","status_misclassification",
  "tenure_contamination","timegap_contamination","job_change_prob",
  "gross_anchor_revision_prob","clean_anchor_revision_prob",
  "exact_anchor_retention_prob")
anchor_show <- data.frame(
  Quantity=c("Entry","Exit","Status error","Tenure error","Timegap error",
    "Job reset","Gross-anchor whole-year revision",
    "Clean-anchor whole-year revision","Exact-anchor retention",
    "Log likelihood","AIC"),
  Baseline=c(sprintf("%.3f",100*unlist(
    exact_anchor_comparison[1,anchor_rates])),
    format(round(exact_anchor_comparison$loglik[1],1),big.mark=",",
      scientific=FALSE),format(round(exact_anchor_comparison$AIC[1],1),
      big.mark=",",scientific=FALSE)),
  Retention=c(sprintf("%.3f",100*unlist(
    exact_anchor_comparison[3,anchor_rates])),
    format(round(exact_anchor_comparison$loglik[3],1),big.mark=",",
      scientific=FALSE),format(round(exact_anchor_comparison$AIC[3],1),
      big.mark=",",scientific=FALSE)),check.names=FALSE)
names(anchor_show)[2:3] <- c("Flexible baseline","Exact retention")
knitr::kable(anchor_show,booktabs=TRUE,
  caption="Exact reported-start-date retention (rates in percent)")
```

The polished joint estimate is
`r sprintf('%.2f',100*exact_anchor_comparison$exact_anchor_retention_prob[3])`%.
The log likelihood rises by
`r format(round(exact_anchor_comparison$loglik[3]-
  exact_anchor_comparison$loglik[1],1),big.mark=',')` points; the corresponding
one-parameter LR statistic is
`r format(round(exact_anchor_lr$LR,1),big.mark=',')`. Entry and exit rates move
only from 2.28% and 4.14% to 2.31% and 4.14%. The substantive transition result
is therefore stable. Tenure contamination rises from 21.82% to
`r sprintf('%.2f',100*exact_anchor_comparison$tenure_contamination[3])`% because
the enlarged gross-report process can now include both retained and redrawn
dates.

```{r exact-anchor-refinement}
refine_keep <- exact_anchor_refinement$stage %in%
  c("continued","low","high","promoted","polished")
refine_show <- exact_anchor_refinement[refine_keep,
  c("stage","loglik","gain_over_incumbent","exact_anchor_retention")]
refine_show$gain_over_incumbent <- refine_show$loglik-
  exact_anchor_refinement$loglik[exact_anchor_refinement$stage=="continued"]
refine_show$stage <- c("First central block","Low-retention start",
  "High-retention start","Continued joint fit","Conditional polish")
refine_show$loglik <- format(round(refine_show$loglik,1),big.mark=",",
  scientific=FALSE)
refine_show$gain_over_incumbent <- sprintf("%+.2f",
  refine_show$gain_over_incumbent)
refine_show$exact_anchor_retention <- sprintf("%.2f",
  100*refine_show$exact_anchor_retention)
names(refine_show) <- c("Fit","Log likelihood","Gain","Retention (%)")
rownames(refine_show) <- NULL
knitr::kable(refine_show,booktabs=TRUE,
  caption="Longer refinement of the exact-retention model")
```

The continued joint refinement raises the log likelihood by
`r sprintf('%.2f',exact_anchor_refinement$loglik[
  exact_anchor_refinement$stage=='polished']-
  exact_anchor_refinement$loglik[exact_anchor_refinement$stage=='continued'])`
points relative to the first central block. Low- and high-retention screens
are hundreds of likelihood points worse. The final scalar
polish adds 0.001 point, so the point estimates are practically stable even
though the capped joint blocks retain optimizer convergence code 1. Full
observed-information inference for the 30 parameters remains outstanding.

```{r exact-anchor-predictive}
anchor_metrics <- c("exact_continuation","whole_year_revision",
  "other_revision")
anchor_labels <- c("Exact continuation","Whole-year revision",
  "Other revision")
ap <- exact_anchor_predictive[
  exact_anchor_predictive$metric %in% anchor_metrics,]
ap <- ap[order(match(ap$metric,anchor_metrics),ap$model),]
ap_show <- data.frame(Measure=rep(anchor_labels,each=2),
  Model=ap$model,Predicted=sprintf("%.2f",100*ap$predicted_mean),
  Observed=sprintf("%.2f",100*ap$observed),check.names=FALSE)
knitr::kable(ap_show,booktabs=TRUE,
  caption="Predictive implications of exact-anchor retention (percent)")
```

Exact continuations rise from
`r sprintf('%.1f',100*ap$predicted_mean[ap$model=='Flexible baseline' & ap$metric=='exact_continuation'])`%
to
`r sprintf('%.1f',100*ap$predicted_mean[ap$model=='Exact-anchor retention' & ap$metric=='exact_continuation'])`%,
toward the observed 64.7%, and other revisions fall from
`r sprintf('%.1f',100*ap$predicted_mean[ap$model=='Flexible baseline' & ap$metric=='other_revision'])`%
to
`r sprintf('%.1f',100*ap$predicted_mean[ap$model=='Exact-anchor retention' & ap$metric=='other_revision'])`%,
toward 28.3%. Whole-year revisions
remain essentially exact at 7.0%. The extension therefore targets the intended
serial feature but does not close the full gap. Likelihood gains are concentrated
in records containing a mixture of continuation patterns and are offset by
losses among single-pattern records. A richer local date-revision kernel remains
warranted.

```{r exact-anchor-recovery}
anchor_recovery_show <- data.frame(
  Quantity=c("Generating retention","Mean recovered retention",
    "Recovery RMSE","Minimum estimate","Maximum estimate",
    "Mean LR against zero"),
  Value=c(exact_anchor_recovery$truth,
    exact_anchor_recovery$mean_estimate,exact_anchor_recovery$rmse,
    exact_anchor_recovery$minimum,exact_anchor_recovery$maximum,
    exact_anchor_recovery$mean_LR_zero))
anchor_recovery_show$Value <- sprintf("%.4f",anchor_recovery_show$Value)
knitr::kable(anchor_recovery_show,booktabs=TRUE,
  caption="Conditional recovery of exact-anchor retention (20 samples of 5,000)")
```

Holding nuisance parameters at their generating values, the new probability
is recovered with small mean bias and an RMSE of about three percentage points.
All 20 replications produce a positive interior estimate and strongly favor
retention over the nested zero value. This establishes conditional
identification, but is intentionally an easier check than joint recovery.

```{r exact-anchor-joint-recovery}
joint_selected <- exact_anchor_joint_fits[exact_anchor_joint_fits$selected,]
transition_recovery <- data.frame(
  Sample=paste("Replication",joint_selected$replication),
  `True entry`=100*joint_selected$true_entry_rate,
  `Estimated entry`=100*joint_selected$entry_rate,
  `True exit`=100*joint_selected$true_exit_rate,
  `Estimated exit`=100*joint_selected$exit_rate,
  check.names=FALSE)
for (j in 2:ncol(transition_recovery))
  transition_recovery[[j]] <- sprintf("%.3f",transition_recovery[[j]])
knitr::kable(transition_recovery,booktabs=TRUE,
  caption="Full 30-parameter recovery: transition rates (percent)")

joint_parameters <- c("exact_anchor_retention","eps","pi",
  "clean_anchor_revision","gross_anchor_revision")
joint_labels <- c("Exact-anchor retention","Tenure contamination",
  "Status misclassification","Clean-anchor revision",
  "Gross-anchor revision")
jr <- exact_anchor_joint_recovery[
  match(joint_parameters,exact_anchor_joint_recovery$parameter),]
joint_recovery_show <- data.frame(Parameter=joint_labels,
  Truth=sprintf("%.2f",100*jr$truth),
  Mean=sprintf("%.2f",100*jr$mean_estimate),
  Bias=sprintf("%+.2f",100*jr$mean_bias),
  RMSE=sprintf("%.2f",100*jr$rmse),check.names=FALSE)
knitr::kable(joint_recovery_show,booktabs=TRUE,
  caption="Full 30-parameter recovery: reporting probabilities (percentage points)")
```

All three simulated samples can be re-estimated from near and dispersed starts.
The selected pooled entry- and exit-rate RMSEs are
`r sprintf('%.3f',100*sqrt(mean(joint_selected$entry_rate_bias^2)))` and
`r sprintf('%.3f',100*sqrt(mean(joint_selected$exit_rate_bias^2)))` percentage
points. The third replication is less favorable than the first two, with
entry and exit errors of
`r sprintf('%+.3f',100*joint_selected$entry_rate_bias[3])` and
`r sprintf('%+.3f',100*joint_selected$exit_rate_bias[3])` points, respectively.
The aggregate transition result is nevertheless much more robust than
individual duration-hazard bins. By
contrast, the reporting decomposition is weakly separated at this sample size:
the retention RMSE is
`r sprintf('%.2f',100*jr$rmse[jr$parameter=='exact_anchor_retention'])` points,
and clean- and gross-anchor revision probabilities partly offset one another.
The exercise shows that the full likelihood is estimable, but three
replications do not establish
precise finite-sample identification of every reporting component.

```{r exact-anchor-opg-inference}
opg_names <- c("entry_rate","exit_rate","status_misclassification",
  "tenure_contamination","timegap_contamination","job_change_prob",
  "gross_january_heaping_prob","gross_anchor_revision_prob",
  "clean_anchor_revision_prob","exact_anchor_retention_prob")
opg_labels <- c("Entry","Exit","Status error","Tenure error",
  "Timegap error","Job reset","Gross-only January heaping",
  "Gross-anchor revision","Clean-anchor revision","Exact-anchor retention")
oi <- exact_anchor_opg[match(opg_names,exact_anchor_opg$quantity),]
opg_show <- data.frame(Quantity=opg_labels,
  Estimate=sprintf("%.3f",100*oi$estimate),
  `Analytical SE`=sprintf("%.3f",100*oi$se),
  `95% interval`=sprintf("[%.3f, %.3f]",100*oi$lower,100*oi$upper),
  check.names=FALSE)
knitr::kable(opg_show,booktabs=TRUE,
  caption="OPG analytical inference for the exact-retention model (percent)")
```

The 30-parameter score matrix is full rank, but its condition number is
`r format(signif(exact_anchor_opg_diagnostics$condition_number,3),scientific=TRUE)`,
consistent with weak local separation. The exact-retention standard error is
0.574 percentage points under this analytical approximation. The most
important natural-scale correlations are 0.885 between exact retention and
overall tenure contamination, and -0.540 between the clean- and gross-anchor
revision probabilities. These correlations agree with the simulation evidence
that reporting channels can offset one another. Gross-only January heaping is
effectively on the boundary: its regular interval crosses zero by a wide
margin, so its Wald standard error should be treated only as a diagnostic.
The OPG calculation uses 61 likelihood evaluations. A resumable central
observed-Hessian calculation has also been implemented; its first 48 of 1,801
points are cached, but the exact calculation is not yet complete.

\clearpage

### Local reported-start-date revisions

The next nested extension addresses near-anchor changes directly. Conditional
on a gross current tenure report, the preceding reported start date is first
retained exactly with probability $\rho$. If it is not retained, the report
receives a non-zero local revision with probability $\kappa$; otherwise it
passes to the existing whole-year-revision versus marginal-redraw mixture. The
local revision is a normalized discrete-Laplace kernel over one to six months
in either direction, with fixed three-month scale and truncation at zero
tenure. Consequently, $\kappa=0$ reproduces the exact-anchor model exactly and
zero-month changes are not counted twice.

```{r local-anchor-comparison}
local_rates <- c("entry_rate","exit_rate","status_misclassification",
  "tenure_contamination","timegap_contamination","job_change_prob",
  "gross_anchor_revision_prob","clean_anchor_revision_prob",
  "exact_anchor_retention_prob","local_anchor_revision_prob")
local_show <- data.frame(
  Quantity=c("Entry","Exit","Status error","Tenure error","Timegap error",
    "Job reset","Gross-anchor whole-year revision",
    "Clean-anchor whole-year revision","Exact-anchor retention",
    "Local revision conditional on no exact retention","Log likelihood",
    "AIC"),
  `Exact retention`=c(sprintf("%.3f",100*unlist(
    local_anchor_comparison[1,local_rates])),
    format(round(local_anchor_comparison$loglik[1],1),big.mark=",",
      scientific=FALSE),format(round(local_anchor_comparison$AIC[1],1),
      big.mark=",",scientific=FALSE)),
  `Local revision`=c(sprintf("%.3f",100*unlist(
    local_anchor_comparison[3,local_rates])),
    format(round(local_anchor_comparison$loglik[3],1),big.mark=",",
      scientific=FALSE),format(round(local_anchor_comparison$AIC[3],1),
      big.mark=",",scientific=FALSE)),check.names=FALSE)
knitr::kable(local_show,booktabs=TRUE,
  caption="Local reported-start-date revision extension (rates in percent)")
```

The conditional local-revision estimate is
`r sprintf('%.2f',100*local_anchor_comparison$local_anchor_revision_prob[3])`%.
Because exact retention is tried first, its corresponding unconditional share
among eligible gross reports is
`r sprintf('%.2f',100*(1-local_anchor_comparison$exact_anchor_retention_prob[3])*
  local_anchor_comparison$local_anchor_revision_prob[3])`%. The log likelihood
increases by
`r format(round(local_anchor_comparison$loglik[3]-
  local_anchor_comparison$loglik[1],1),big.mark=',')` points, giving a boundary
LR statistic of `r format(round(local_anchor_lr$LR,1),big.mark=',')` and a much
lower AIC. The conditional profile peaks near 0.35. Reporting-parameter fits
started near 0.02, the conditional optimum, and 0.65; the low and high screens
are respectively
`r format(round(max(local_anchor_refinement$loglik)-
  local_anchor_refinement$loglik[local_anchor_refinement$stage=='low'],1),
  big.mark=',')` and
`r format(round(max(local_anchor_refinement$loglik)-
  local_anchor_refinement$loglik[local_anchor_refinement$stage=='high'],1),
  big.mark=',')` likelihood points below the best result. A central-difference
reporting block improves the main forward result and a separate transition-
hazard block adds only
`r sprintf('%.2f',local_anchor_refinement$loglik[
  local_anchor_refinement$stage=='transition']-
  local_anchor_refinement$loglik[local_anchor_refinement$stage=='central'])`
points. The blocks are deliberately short and retain convergence code 1, so
this is strong multistart evidence but not a claim of final full convergence.

Entry changes from 2.308% to
`r sprintf('%.3f',100*local_anchor_comparison$entry_rate[3])`% and exit from
4.142% to `r sprintf('%.3f',100*local_anchor_comparison$exit_rate[3])`%.
Status error also remains near 2.96%. The new channel therefore changes the
measurement process while leaving the paper's transition interpretation
essentially intact.

\clearpage

```{r local-anchor-predictive}
local_metrics <- c("exact_continuation","local_revision_1_to_6_months",
  "whole_year_revision","other_revision")
local_labels <- c("Exact continuation","Local revision (1 to 6 months)",
  "Whole-year revision","Other revision")
lp <- local_anchor_predictive[
  local_anchor_predictive$metric %in% local_metrics,]
lp <- lp[order(match(lp$metric,local_metrics),lp$model),]
lp_show <- data.frame(Measure=rep(local_labels,each=2),Model=lp$model,
  Predicted=sprintf("%.2f",100*lp$predicted_mean),
  Observed=sprintf("%.2f",100*lp$observed),check.names=FALSE)
knitr::kable(lp_show,booktabs=TRUE,
  caption="Predictive implications of local anchor revisions (percent)")
```

The old exact-anchor model predicts only
`r sprintf('%.1f',100*lp$predicted_mean[lp$model=='Exact-anchor model' &
  lp$metric=='local_revision_1_to_6_months'])`% local revisions, compared with
`r sprintf('%.1f',100*lp$observed[lp$metric==
  'local_revision_1_to_6_months'][1])`% observed. The extension raises the
prediction to
`r sprintf('%.1f',100*lp$predicted_mean[lp$model=='Local anchor revisions' &
  lp$metric=='local_revision_1_to_6_months'])`%. Local-only records contribute
`r format(round(local_anchor_decomposition$loglik_gain[
  local_anchor_decomposition$pattern=='Local revision only']),big.mark=',')`
likelihood points and mixed-pattern records contribute another
`r format(round(local_anchor_decomposition$loglik_gain[
  local_anchor_decomposition$pattern=='Mixed']),big.mark=',')`; losses on
other and whole-year-only records partly offset these gains. This confirms
that the extension improves the feature it was designed to capture rather
than obtaining its fit from unrelated observations.

The check is not yet fully satisfactory. Exact continuations remain
underpredicted (about 61.7% versus 64.7%), local revisions remain slightly
underpredicted (10.8% versus 11.4%), and residual revisions remain
overpredicted (22.2% versus 16.9%). Whole-year revisions now fall below the
observed share. The fixed local kernel is therefore a useful first repair, not
a complete reporting model. Holding all other parameters fixed, the local
probability has a curvature-based standard error of
`r sprintf('%.3f',100*local_anchor_curvature$se_probability)` percentage
points and a 95% interval of
[`r sprintf('%.2f',100*local_anchor_curvature$lower_probability)`%,
`r sprintf('%.2f',100*local_anchor_curvature$upper_probability)`%]. This is
conditional rather than full nuisance-adjusted inference and should not be
used as the final paper standard error.

\clearpage

### Shared person-level duration-reporting reliability

The next extension allows tenure and timegap reports from the same person to
share a latent propensity for contamination. Each person belongs with equal
prior probability to a reliable or unreliable reporting class. Relative to a
common midpoint, the reliable class shifts both contamination logits down by
$\delta$ and the unreliable class shifts them up by $\delta$. Thus
$\delta=0$ reproduces the preceding independent-reporting likelihood exactly.
The fixed equal class shares keep this first extension parsimonious and avoid
an additional mixing-proportion parameter and label switching.

```{r shared-reliability-comparison}
sr <- shared_reliability_comparison[c(1,3),]
shared_show <- data.frame(
  Quantity=c("Entry","Exit","Status error","Job reset",
    "Exact-anchor retention","Local revision",
    "Reliability-logit dispersion","Log likelihood","AIC"),
  Independent=c(sprintf("%.3f",100*c(sr$entry_rate[1],sr$exit_rate[1],
    sr$status_misclassification[1],sr$job_change_prob[1],
    sr$exact_anchor_retention_prob[1],sr$local_anchor_revision_prob[1])),
    "0.000",format(round(sr$loglik[1],1),big.mark=",",scientific=FALSE),
    format(round(sr$AIC[1],1),big.mark=",",scientific=FALSE)),
  `Shared reliability`=c(sprintf("%.3f",100*c(sr$entry_rate[2],
    sr$exit_rate[2],sr$status_misclassification[2],sr$job_change_prob[2],
    sr$exact_anchor_retention_prob[2],sr$local_anchor_revision_prob[2])),
    sprintf("%.3f",sr$duration_reliability_shift[2]),
    format(round(sr$loglik[2],1),big.mark=",",scientific=FALSE),
    format(round(sr$AIC[2],1),big.mark=",",scientific=FALSE)),
  check.names=FALSE)
knitr::kable(shared_show,booktabs=TRUE,
  caption="Shared duration-reporting reliability (rates in percent)")
```

The likelihood rises by
`r format(round(diff(sr$loglik),1),big.mark=',')` points with one additional
parameter, while entry remains `r sprintf('%.3f',100*sr$entry_rate[2])`% and
exit remains `r sprintf('%.3f',100*sr$exit_rate[2])`%. The reporting extension
therefore leaves the substantive transition estimates essentially unchanged.
The exact equality of the two likelihoods at $\delta=0$ was verified
numerically. The LR statistic is
`r format(round(shared_reliability_lr$LR,1),big.mark=',')`, but its usual
one-degree-of-freedom chi-squared reference distribution is not valid because
the null is a homogeneous-mixture boundary. Final inference requires a
parametric bootstrap.

```{r shared-reliability-components}
component_show <- data.frame(
  Class=c("Reliable","Unreliable","Equal-prior marginal mean"),
  `Prior share`=c(sprintf("%.1f",100*
    shared_reliability_components$prior_probability),"--"),
  `Tenure contamination`=sprintf("%.2f",100*c(
    shared_reliability_components$tenure_contamination,
    mean(shared_reliability_components$tenure_contamination))),
  `Timegap contamination`=sprintf("%.2f",100*c(
    shared_reliability_components$timegap_contamination,
    mean(shared_reliability_components$timegap_contamination))),
  check.names=FALSE)
knitr::kable(component_show,booktabs=TRUE,
  caption="Implied duration-contamination probabilities by latent reliability class (percent)")
```

The reliable class has tenure and timegap contamination probabilities of
`r sprintf('%.2f',100*shared_reliability_components$tenure_contamination[1])`%
and
`r sprintf('%.2f',100*shared_reliability_components$timegap_contamination[1])`%,
compared with
`r sprintf('%.2f',100*shared_reliability_components$tenure_contamination[2])`%
and
`r sprintf('%.2f',100*shared_reliability_components$timegap_contamination[2])`%
in the unreliable class. The survey-weighted mean posterior probability of
the unreliable class is
`r sprintf('%.1f',100*shared_reliability_posterior$posterior_unreliable[shared_reliability_posterior$statistic=='weighted mean'])`%,
with a median of
`r sprintf('%.1f',100*shared_reliability_posterior$posterior_unreliable[shared_reliability_posterior$statistic=='median'])`%.
The midpoint contamination parameters reported by the optimizer are not
population-average probabilities in this mixture; the class-specific values
and equal-prior marginal means above are the interpretable quantities.

All three deliberately dispersed joint starts improve on the independent
model. The best low and high starts remain respectively
`r format(round(max(shared_reliability_refinement$loglik)-shared_reliability_refinement$loglik[shared_reliability_refinement$stage=='low'],1),big.mark=',')`
and
`r format(round(max(shared_reliability_refinement$loglik)-shared_reliability_refinement$loglik[shared_reliability_refinement$stage=='high'],1),big.mark=',')`
likelihood points below the preferred basin. The short joint blocks terminate
at their iteration caps, however, so the result is a strong provisional point
estimate rather than a final convergence claim.

```{r shared-reliability-predictive}
shared_metrics <- c("exact_continuation","local_revision_1_to_6_months",
  "other_revision","timegap_inconsistent",
  "two_opportunities_no_irregularity",
  "two_opportunities_one_irregularity",
  "two_opportunities_two_irregularities")
shared_labels <- c("Exact continuation","Local revision (1 to 6 months)",
  "Other revision","Timegap inconsistency",
  "Two opportunities: no irregularity",
  "Two opportunities: one irregularity",
  "Two opportunities: two irregularities")
sp <- shared_reliability_predictive[
  shared_reliability_predictive$metric %in% shared_metrics,]
sp$label <- shared_labels[match(sp$metric,shared_metrics)]
sp_wide <- reshape(sp[,c("label","model","predicted_mean")],
  idvar="label",timevar="model",direction="wide")
sp_obs <- sp[!duplicated(sp$label),c("label","observed")]
sp_wide <- merge(sp_obs,sp_wide,by="label",sort=FALSE)
sp_wide <- sp_wide[match(shared_labels,sp_wide$label),]
shared_pred_show <- data.frame(
  Measure=sp_wide$label,Observed=sprintf("%.2f",100*sp_wide$observed),
  Independent=sprintf("%.2f",100*sp_wide$`predicted_mean.Independent reliability`),
  `Shared reliability`=sprintf("%.2f",100*sp_wide$`predicted_mean.Shared reliability`),
  check.names=FALSE)
knitr::kable(shared_pred_show,booktabs=TRUE,
  caption="Predictive implications of shared reporting reliability (percent)")
```

Shared reliability improves the local-revision prediction from 10.77% to
11.13% against 11.43% observed and moves the repeated-opportunity distribution
in the right direction: the predicted shares with no and exactly one
irregularity change from 60.36% and 24.07% to 62.88% and 19.39%, against
68.05% and 16.43% observed. It is not a complete repair. The model raises the
two-irregularity share from 15.58% to 17.73% against 15.52% observed, and the
exact, residual-revision, and timegap moments remain outside their simulation
intervals.

The likelihood decomposition confirms that the improvement is concentrated
where duration information is informative: records with no irregularity,
tenure deviations, and timegap inconsistencies account for 65.6%, 28.9%, and
5.8% of the gain, respectively. The average posterior unreliable probability
is 28.0% for regular repeated-state records, 87.5% for tenure deviations, and
92.6% for timegap inconsistencies.

There is an important identification limitation. Across only three waves, a
record cannot contain both a repeated employed spell and a repeated
nonemployment spell. The common shift across tenure and timegap is therefore
not identified from observing both clocks repeatedly for the same person.
Instead, it is a parsimonious cross-clock restriction learned indirectly from
serial clustering within each clock. The following specification check allows
separate tenure- and timegap-reliability dispersions and tests whether the
common-shift restriction is supported.

\clearpage

#### Relaxing the common-shift restriction

The restriction test replaces the single $\delta$ with separate nonnegative
tenure and timegap dispersions, while retaining the same person-level class
and equal prior shares. Setting the two dispersions equal reproduces the
common-shift likelihood exactly; this was verified to numerical precision.

```{r separate-reliability-comparison}
sep <- separate_reliability_comparison
separate_show <- data.frame(
  Quantity=c("Entry","Exit","Status error","Job reset",
    "Exact-anchor retention","Local revision","Tenure dispersion",
    "Timegap dispersion","Log likelihood","AIC"),
  `Common dispersion`=c(sprintf("%.3f",100*c(sep$entry_rate[1],
    sep$exit_rate[1],sep$status_misclassification[1],
    sep$job_change_prob[1],sep$exact_anchor_retention_prob[1],
    sep$local_anchor_revision_prob[1])),
    sprintf("%.3f",c(sep$tenure_reliability_shift[1],
      sep$timegap_reliability_shift[1])),
    format(round(sep$loglik[1],1),big.mark=",",scientific=FALSE),
    format(round(sep$AIC[1],1),big.mark=",",scientific=FALSE)),
  `Separate dispersions`=c(sprintf("%.3f",100*c(sep$entry_rate[2],
    sep$exit_rate[2],sep$status_misclassification[2],
    sep$job_change_prob[2],sep$exact_anchor_retention_prob[2],
    sep$local_anchor_revision_prob[2])),
    sprintf("%.3f",c(sep$tenure_reliability_shift[2],
      sep$timegap_reliability_shift[2])),
    format(round(sep$loglik[2],1),big.mark=",",scientific=FALSE),
    format(round(sep$AIC[2],1),big.mark=",",scientific=FALSE)),
  check.names=FALSE)
knitr::kable(separate_show,booktabs=TRUE,
  caption="Relaxing equal tenure and timegap reliability dispersions (rates in percent)")
```

The estimated tenure dispersion is
`r sprintf('%.3f',sep$tenure_reliability_shift[2])`, compared with
`r sprintf('%.3f',sep$timegap_reliability_shift[2])` for timegap. The log
likelihood rises by
`r format(round(diff(sep$loglik),1),big.mark=',')` points and AIC falls by
`r format(round(-diff(sep$AIC),1),big.mark=',')`. The equality-restriction LR
statistic is `r format(round(separate_reliability_lr$LR,1),big.mark=',')` on
one degree of freedom. This is a like-for-like comparison: the unrestricted
fit was continued jointly until a further step made no improvement, the
restricted fit was initialized by projecting those nuisance estimates onto a
common shift, its common shift was polished conditionally, and a final full
restricted step also made no improvement. Both retained fits have convergence
code zero. The earlier unrefined LR statistic was much larger because it
incorrectly attributed unrelated nuisance-parameter refinement to the single
dispersion restriction.

Entry and exit remain
`r sprintf('%.3f',100*sep$entry_rate[2])`% and
`r sprintf('%.3f',100*sep$exit_rate[2])`%. Relaxing the restriction therefore
changes the duration-reporting decomposition without changing the paper's
substantive employment-transition estimates.

```{r separate-reliability-components}
separate_component_show <- data.frame(
  Class=c("Reliable","Unreliable","Equal-prior marginal mean"),
  `Prior share`=c(sprintf("%.1f",100*
    separate_reliability_components$prior_probability),"--"),
  `Tenure contamination`=sprintf("%.2f",100*c(
    separate_reliability_components$tenure_contamination,
    mean(separate_reliability_components$tenure_contamination))),
  `Timegap contamination`=sprintf("%.2f",100*c(
    separate_reliability_components$timegap_contamination,
    mean(separate_reliability_components$timegap_contamination))),
  check.names=FALSE)
knitr::kable(separate_component_show,booktabs=TRUE,
  caption="Contamination probabilities under separate reliability dispersions (percent)")
```

The unrestricted model sharpens the tenure separation: tenure contamination
is `r sprintf('%.2f',100*separate_reliability_components$tenure_contamination[1])`%
in the reliable class and
`r sprintf('%.2f',100*separate_reliability_components$tenure_contamination[2])`%
in the unreliable class. The corresponding timegap probabilities are
`r sprintf('%.2f',100*separate_reliability_components$timegap_contamination[1])`%
and
`r sprintf('%.2f',100*separate_reliability_components$timegap_contamination[2])`%.
The survey-weighted
mean posterior unreliable probability is
`r sprintf('%.1f',100*separate_reliability_posterior$posterior_unreliable[separate_reliability_posterior$statistic=='weighted mean'])`%.

```{r separate-reliability-inference}
inference_labels <- c(
  entry_rate="Entry rate",exit_rate="Exit rate",
  status_misclassification="Status error",
  reliable_tenure="Tenure contamination: reliable",
  unreliable_tenure="Tenure contamination: unreliable",
  reliable_timegap="Timegap contamination: reliable",
  unreliable_timegap="Timegap contamination: unreliable",
  tenure_reliability_shift="Tenure dispersion",
  timegap_reliability_shift="Timegap dispersion",
  shift_difference="Timegap minus tenure dispersion")
sir <- separate_reliability_inference[
  separate_reliability_inference$quantity %in% names(inference_labels),]
sir <- sir[match(names(inference_labels),sir$quantity),]
probability <- !sir$quantity %in% c("tenure_reliability_shift",
  "timegap_reliability_shift","shift_difference")
scale <- ifelse(probability,100,1)
inference_show <- data.frame(
  Quantity=unname(inference_labels[sir$quantity]),
  Estimate=sprintf("%.3f",scale*sir$estimate),
  `OPG standard error`=sprintf("(%.3f)",scale*sir$se),check.names=FALSE)
knitr::kable(inference_show,booktabs=TRUE,
  caption="Provisional full-nuisance OPG inference (rates in percent)")
```

The 33-parameter OPG matrix is full rank and positive definite, but its
condition number is
`r format(round(separate_reliability_information$condition_number),big.mark=',')`.
The fourth exit-hazard perturbation required a tenfold smaller finite-
difference step. Most importantly, the OPG Wald interval for the difference
between the two dispersions includes zero even though the directly optimized
LR statistic is large. The OPG is an outer-product approximation rather than
the exact observed Hessian, and the mixture likelihood is evidently not well
approximated by one common local quadratic for this contrast. The LR comparison
is therefore the preferred evidence about the equality restriction; the OPG
errors are useful provisional uncertainty measures for the headline rates but
not a substitute for a profile or bootstrap of the dispersion difference.

```{r separate-reliability-profile}
srpi <- separate_reliability_profile_interval[1,]
profile_show <- data.frame(
  Quantity=c("Timegap minus tenure dispersion","95% profile interval",
    "LR test of equal dispersions","LR p-value"),
  Result=c(sprintf("%.4f",srpi$estimate),
    sprintf("[%.4f, %.4f]",srpi$profile_lower,srpi$profile_upper),
    sprintf("%.2f",srpi$equality_LR),
    format(srpi$equality_p_value,digits=3,scientific=TRUE)))
knitr::kable(profile_show,booktabs=TRUE,
  caption="Nuisance-adjusted profile inference for the dispersion difference")
```

The direct profile fixes the timegap-minus-tenure dispersion difference and
reoptimizes the other 32 parameters. Each outer bracket was screened from a
nearby solution and from the common-dispersion nuisance estimates; all
retained fits converged, and the nearby starts gave the higher likelihood on
both sides. Tightly optimized points near both confidence limits
give a 95% profile interval from
`r sprintf('%.4f',srpi$profile_lower)` to
`r sprintf('%.4f',srpi$profile_upper)`, which excludes zero by a wide margin.
The profile therefore agrees with the fully optimized LR test and shows that
the broad OPG Wald interval is a numerical/local-quadratic failure for this
specific contrast. It does not invalidate the provisional OPG uncertainty
measures for the headline transition rates.

A 10,000-person simulation-recovery check held the remaining nuisance
parameters at their generating values and estimated the two midpoints and two
dispersions from three widely separated starts. All starts converged to the
same likelihood. Tenure midpoint and dispersion biases were
`r sprintf('%+.3f',100*(separate_reliability_recovery$eps-
  separate_reliability_comparison$tenure_contamination_midpoint[2]))`
percentage points and
`r sprintf('%+.3f',separate_reliability_recovery$tenure_reliability_shift-
  separate_reliability_comparison$tenure_reliability_shift[2])`;
timegap was noisier, with corresponding biases of
`r sprintf('%+.3f',100*(separate_reliability_recovery$eps_d-
  separate_reliability_comparison$timegap_contamination_midpoint[2]))`
points and
`r sprintf('%+.3f',separate_reliability_recovery$timegap_reliability_shift-
  separate_reliability_comparison$timegap_reliability_shift[2])`.
The posterior reliability score classified the simulated latent groups with
AUC `r sprintf('%.3f',separate_reliability_recovery$AUC)`. This is encouraging
but remains a one-replication, conditional recovery diagnostic rather than a
complete finite-sample validation of all 33 parameters.

```{r separate-reliability-predictive}
sep_metrics <- c("exact_continuation","local_revision_1_to_6_months",
  "other_revision","timegap_inconsistent",
  "two_opportunities_no_irregularity",
  "two_opportunities_one_irregularity",
  "two_opportunities_two_irregularities")
sep_labels <- c("Exact continuation","Local revision (1 to 6 months)",
  "Other revision","Timegap inconsistency",
  "Two opportunities: no irregularity",
  "Two opportunities: one irregularity",
  "Two opportunities: two irregularities")
spp <- separate_reliability_predictive[
  separate_reliability_predictive$metric %in% sep_metrics,]
spp$label <- sep_labels[match(spp$metric,sep_metrics)]
spp_wide <- reshape(spp[,c("label","model","predicted_mean")],
  idvar="label",timevar="model",direction="wide")
spp_obs <- spp[!duplicated(spp$label),c("label","observed")]
spp_wide <- merge(spp_obs,spp_wide,by="label",sort=FALSE)
spp_wide <- spp_wide[match(sep_labels,spp_wide$label),]
separate_pred_show <- data.frame(
  Measure=spp_wide$label,Observed=sprintf("%.2f",100*spp_wide$observed),
  `Common dispersion`=sprintf("%.2f",100*
    spp_wide$`predicted_mean.Common dispersion`),
  `Separate dispersions`=sprintf("%.2f",100*
    spp_wide$`predicted_mean.Separate dispersions`),check.names=FALSE)
knitr::kable(separate_pred_show,booktabs=TRUE,
  caption="Predictive implications of separate reliability dispersions (percent)")
```

Separate dispersions improve exact continuations from 63.8% to 64.1% against
64.7% observed. They change the predicted share with two regular duration
opportunities from 65.0% to 64.8% against 68.1%, and increase predicted
timegap inconsistencies from 16.6% to 17.1% against 15.5%. The local-revision
prediction also moves away from its observed value, and all reported moments
remain outside the simulation intervals. The extension is preferred by the
observed likelihood, but is not an adequate reporting model on these coarse
predictive checks.

The likelihood decomposition makes the tradeoff explicit. Regular
repeated-state records lose about 591 likelihood points relative to the
fully refined common-dispersion model, tenure deviations lose about 559, and
records without repeated-state duration information lose about 9. These are
more than offset by a gain of about 1,231 among timegap-inconsistent records.
Thus the likelihood improvement comes from fitting the distribution within
the timegap-inconsistent group, not from reducing its aggregate predicted
frequency.

# 6. Script and output register

**`diagnostics/audit_job_start_dates.R`.** Implemented. Reconstructs
reported start months and audits heaping, revisions, persistence, and matching
sensitivity. It is a base-R script, accepts an A/B/C panel path, and writes
panel-specific and combined outputs.

**`EM-tenure/R/job_change_eps.R`.** Implemented. Enumerates same-job and
new-job branches inside full latent employment spells, exactly nests the old
emission at (q=0), uses log-sum-exp, and exposes posterior job-change counts.

**`EM-tenure/estimate_job_change_tenure_contamination.R`.** Implemented.
Profiles (q), jointly estimates it with all nuisance parameters, rejects an
invalid or inferior fit, and preserves the preferred Table 7 fit as the nested
comparison. It does not bootstrap.

**`EM-tenure/estimate_job_change_matching_robustness.R`.** Implemented.
Re-estimates the identical 15-parameter model on Panels B and C from the Panel
A optimum and writes separate robustness outputs without replacing the main
fit.

**`EM-tenure/validate_job_change_model.R`.** Implemented. Runs dispersed
starts, tight refinement, a nuisance-adjusted profile, multistart full-model
recovery simulations, and a large-sample score diagnostic. The recovery check
currently fails and identifies the monthly-mass repair as the next priority.

**`EM-tenure/R/tenure_monthly_eps.R`.** Implemented. Replaces continuous
tenure densities with monthly probability masses, advances clean clocks by
three months, and uses a discrete 0--2 month support for entries and resets.

**`EM-tenure/validate_monthly_tenure_model.R`.** Implemented. Checks mass
normalization, evaluates large-sample scores, and re-estimates simulated data
from dispersed starts. The first full-likelihood recovery exercise is close to
the generating parameters.

**`EM-tenure/validate_monthly_tenure_recovery_grid.R`.** Implemented. Runs a
resumable recovery grid at multiple sample sizes, repeats dispersed-start
checks, and reports bias, RMSE, hazard recovery, and likelihood scores.

**`EM-tenure/estimate_monthly_job_change_tenure_contamination.R`.**
Implemented. Preserves reported zero-month tenure, estimates the no-reset and
reset models on Panel A, and compares three reset starting values.

**`EM-tenure/estimate_monthly_job_change_matching_robustness.R`.**
Implemented. Re-estimates the corrected monthly model on Panels B and C,
checks the monthly grid, screens low, main, and high reset starts, and writes
separate cached robustness outputs without replacing Panel A.

**`EM-tenure/estimate_monthly_persistent_tenure_reporting.R`.** Implemented.
Adds a within-job retain-versus-redraw process for consecutive gross tenure
reports, checks exact nesting, screens three persistence starts, and retains
the exact boundary fit when all positive approximations are inferior.

**`EM-tenure/validate_monthly_persistent_tenure_reporting.R`.** Implemented.
Simulates positive persistence, evaluates a conditional profile, and performs
two full-likelihood recovery fits from dispersed starts.

**`EM-tenure/estimate_monthly_calendar_heaping.R`.** Implemented. Retains the
nominal interview quarter, adds a normalized January-conditioned component to
gross tenure reports, verifies exact nesting, and screens three Panel A starts.

**`EM-tenure/validate_monthly_calendar_heaping.R`.** Implemented. Simulates a
positive January-heaping probability, evaluates its conditional likelihood,
and checks joint recovery from low and high starts.

**`EM-tenure/infer_monthly_calendar_heaping.R`.** Implemented. Evaluates the
full numerical observed-information matrix in resumable parallel chunks and
uses the delta method for the reported rates and calendar-heaping quantities.

**`EM-tenure/estimate_monthly_calendar_revision.R`.** Implemented. Adds a
normalized, correlated whole-year revision branch to eligible consecutive
gross reports, verifies exact nesting, and screens three Panel A starts.

**`EM-tenure/validate_monthly_calendar_revision.R`.** Implemented. Verifies
conditional identification and joint recovery of whole-year revision and
January-heaping probabilities from dispersed starts.

**`EM-tenure/infer_monthly_calendar_revision.R`.** Implemented. Computes the
full 17-parameter observed-information matrix in resumable parallel chunks and
uses the delta method for all reported rates.

**`EM-tenure/profile_monthly_calendar_revision.R`.** Implemented. Reoptimizes
all nuisance parameters from warm and alternative starts on both sides of the
whole-year revision estimate and interpolates the 95% profile interval.

**`EM-tenure/diagnose_monthly_calendar_revision.R`.** Implemented. Simulates
calendar moments at fitted parameters and decomposes the exact observed-data
likelihood gain by mutually exclusive observed patterns.

**`EM-tenure/estimate_monthly_clean_anchor_revision.R`.** Implemented. Adds a
separate whole-year revision probability for a current gross report following
a clean reported date, verifies exact nesting, and screens three starts.

**`EM-tenure/validate_monthly_clean_anchor_revision.R`.** Implemented. Checks
the conditional likelihood and joint recovery of clean-anchor, gross-anchor,
and January-heaping probabilities from two dispersed starts.

**`EM-tenure/diagnose_monthly_clean_anchor_revision.R`.** Implemented. Compares
fitted calendar moments with the nested model and decomposes the row-level
likelihood gain by observable report patterns.

**`EM-tenure/estimate_monthly_start_month_baseline.R`.** Implemented. Adds 11
month logits shared by clean and gross marginal reports, verifies exact nesting
at equal probabilities, screens three conditional starts, and performs staged
joint Panel A refinement.

**`EM-tenure/refine_monthly_start_month_baseline.R`.** Implemented. Continues
the full-sample central-difference refinement without repeating the conditional
mode search.

**`EM-tenure/diagnose_monthly_start_month_baseline.R`.** Implemented. Compares
the full fitted month distribution and consecutive-report patterns with the
nested model and decomposes the likelihood gain by observed start month.

**`EM-tenure/validate_monthly_start_month_baseline.R`.** Implemented. Checks
distant month-distribution starts and joint recovery, including a safeguarded
one-sided gradient near the zero-heaping boundary.

**`EM-tenure/estimate_monthly_exact_anchor_retention.R`.** Implemented. Adds a
single retain-versus-revise/redraw channel after either a clean or gross
reported anchor, verifies exact nesting, profiles the new parameter widely,
and performs staged joint refinement plus a conditional polish.

**`EM-tenure/diagnose_monthly_exact_anchor_retention.R`.** Implemented.
Compares exact, whole-year, and other consecutive-report changes with fitted
simulations and decomposes the observed likelihood gain by report pattern.

**`EM-tenure/validate_monthly_exact_anchor_retention.R`.** Implemented. Runs
20 conditional recovery samples at the empirical generating value and records
bias, RMSE, range, and the LR statistic against zero.

**`EM-tenure/refine_monthly_exact_anchor_retention.R`.** Implemented. Resumes
the saved 30-parameter fit, compares continued, low-retention, and
high-retention starts, promotes the best joint fit, and finishes with an exact
conditional polish without discarding cached work.

**`EM-tenure/validate_monthly_exact_anchor_joint_recovery.R`.** Implemented.
Simulates the complete model and re-estimates all 30 parameters from near and
dispersed starts, with resumable replication caches and separate summaries of
transition, hazard, calendar-baseline, and reporting-process recovery.

**`EM-tenure/infer_monthly_exact_anchor_retention_opg.R`.** Implemented. Uses
61 finite-difference likelihood evaluations and record-level scores to obtain
a full-rank OPG information approximation, delta-method errors, boundary
flags, and correlations among natural-scale reporting probabilities.

**`EM-tenure/infer_monthly_exact_anchor_retention.R`.** Implemented as a
resumable exact central observed-Hessian calculation. It requires 1,801 full
likelihood evaluations; 48 are currently cached, so this remains a
confirmatory calculation rather than the reported inference result.

**`EM-tenure/estimate_monthly_local_anchor_revision.R`.** Implemented. Adds a
normalized one-to-six-month revision kernel conditional on not retaining the
preceding anchor exactly. It verifies exact nesting, profiles the new
probability, screens low and high starts, refines the competing reporting
parameters, checks the transition-hazard block, and caches every completed
stage. The calendar fitter now accepts named free-parameter blocks while
retaining its full-fit default.

**`EM-tenure/diagnose_monthly_local_anchor_revision.R`.** Implemented. Compares
exact, local, whole-year, and residual revisions in observed and simulated
data, decomposes the likelihood gain by report pattern, and reports a clearly
labelled conditional-curvature approximation for the new probability.

**`EM-tenure/estimate_monthly_shared_duration_reliability.R`.** Implemented.
Adds an equal-weight person-level reliability mixture shared by tenure and
timegap contamination, verifies exact nesting at zero dispersion, profiles
the new parameter, and performs resumable dispersed-start joint refinement.

**`EM-tenure/diagnose_monthly_shared_duration_reliability.R`.** Implemented.
Compares observed and fitted duration-irregularity patterns, decomposes the
row-level likelihood gain, and summarizes posterior unreliable-class
probabilities without interpreting class membership as an observed trait.

**`EM-tenure/estimate_monthly_separate_duration_reliability.R`.** Implemented.
Allows tenure and timegap contamination to have separate reliability
dispersions, verifies exact nesting when they are equal, maps the conditional
surface, and screens common and asymmetric starts before a conditional polish.

**`EM-tenure/diagnose_monthly_separate_duration_reliability.R`.** Implemented.
Compares fitted duration-pattern distributions under common and separate
dispersions and decomposes their row-level likelihood difference.

**Separate-reliability refinements.** Implemented in:

```
EM-tenure/refine_monthly_separate_duration_reliability.R
EM-tenure/refine_monthly_separate_reliability_blocks.R
EM-tenure/refine_monthly_separate_reliability_joint.R
```

Continue the four focal parameters, test reporting, transition-hazard, and
full blocks, and finish in interruption-safe joint checkpoints.

**Common-reliability refinements.** Implemented in:

```
EM-tenure/refine_monthly_common_duration_reliability.R
EM-tenure/polish_monthly_common_duration_reliability.R
```

Project the converged nuisance estimates into the equality-restricted model,
polish its common shift, and verify that a full restricted step cannot improve
the comparison likelihood.

**`EM-tenure/validate_monthly_separate_duration_reliability.R`.** Implemented.
Runs cached multiple-start conditional recovery simulations for the two
midpoints and two dispersions and checks posterior latent-class separation.

**`EM-tenure/infer_monthly_separate_duration_reliability_opg.R`.** Implemented.
Computes full-nuisance OPG/delta-method errors from cached record-level scores,
with adaptive perturbations and explicit rank, conditioning, and score
diagnostics.

**Separate-reliability profile.** Implemented in:

```
EM-tenure/profile_monthly_separate_duration_reliability.R
EM-tenure/refine_monthly_separate_duration_reliability_profile_limits.R
```

Reparameterize the clock-specific dispersions by their average
and fixed difference, screen both profile sides from distinct nuisance starts,
and tightly refine the two 95% likelihood-ratio crossings.

**[Analytical-inference script](../EM-tenure/infer_monthly_job_change_tenure_contamination.R).** Implemented.
Computes the full observed-information matrix in parallel, propagates it to
reported rates by the delta method, and reoptimizes all nuisance parameters for
the reset-probability profile.

**[Profile-refinement script](../EM-tenure/refine_monthly_profile_point.R).**
Implemented. Refines a single
fixed reset-probability point from the nearest converged saved profile fit.

**`EM-tenure/tests/testthat/test-job-change-eps.R`.** Implemented. Covers
unit behavior, exact emission and complete-likelihood nesting, finite
likelihoods, strict validation, posterior reset classification, and recovery
of (q) in a simulated emission model.

Generated diagnostics belong under
`diagnostics/output/job_start_dates/`; fitted-model outputs belong under
`EM-tenure/output/results/job_change/`. Both are ignored working artifacts.

# 7. Acceptance checks for the first extension

The first empirical result should not be added to Table 7 until all of the
following are complete. Checks already passed are marked accordingly:

- exact equality to the existing likelihood at (q=0) --- **passed**;
- finite likelihoods at interior (q) values and valid posterior probabilities
  --- **passed**;
- optimizer convergence from the nested fit --- **passed**;
- no deterioration relative to the nested model --- **passed**;
- recovery of (q) in a simplified emission simulation --- **passed**;
- dispersed starts and tighter tolerances --- **passed**, with an inferior
  low-reset local mode that makes multistart selection necessary;
- recovery of (q) and nuisance parameters in full-likelihood simulations ---
  **failed** under the current continuous-density approximation;
- discrete monthly masses sum to one and zero-month reports are retained ---
  **passed**;
- discrete-month full-likelihood recovery from dispersed starts ---
  **passed** in six repeated fits at 10,000 and 40,000 observations, with bias
  and RMSE contracting at the larger sample;
- discrete-month Panel A convergence from low, main, and high reset starts ---
  **passed**, with closely agreeing estimates and log likelihoods;
- matching-panel sensitivity --- **passed** for both the continuous diagnostic
  and the corrected discrete model; all B/C starts converge, while reset and
  contamination estimates decline monotonically with stricter matching; and
- full-rank positive-definite observed information and delta-method errors ---
  **passed**; and
- nuisance-adjusted profile-likelihood inference --- **passed**, with all
  retained points converged and a profile interval nearly identical to the
  analytical interval; and
- positive persistence recoverability --- **passed provisionally** in one
  12,000-record simulation; both joint starts identify substantial positive
  persistence, although additional replications are required to characterize
  its finite-sample bias; and
- January-heaping normalization, nesting, recovery, and Panel A multistart
  convergence --- **passed**; all three empirical starts estimate about 24%
  conditional heaping, and a 6,000-record simulation closely recovers a 40%
  generating probability; and
- January-heaping observed-information inference --- **passed**; the full
  16-parameter information matrix is positive definite and full rank; and
- whole-year revision normalization, nesting, recovery, and Panel A multistart
  convergence --- **passed**; the recovery exercise separates the revision
  and January-heaping components, and the empirical results are reported
  above;
- whole-year revision observed-information and nuisance-profile inference ---
  **passed**; the information matrix is full rank and positive definite, and
  the two interval estimates agree closely; and
- whole-year revision predictive adequacy --- **failed**; the extension moves
  the targeted moment in the correct direction but materially underpredicts
  both January starts and whole-year revisions, while its positive likelihood
  gain is concentrated in 6.16% of weighted records;
- clean-anchor revision nesting, multistart convergence, and joint recovery
  --- **passed provisionally**; three empirical starts agree and two simulated
  starts recover both revision channels, although an 8,000-record exercise
  shows offsetting finite-sample bias between them;
- clean-anchor observed-information inference --- **passed**; the full
  18-parameter matrix is positive definite and full rank; and
- clean-anchor predictive adequacy --- **failed but improved**; the predicted
  whole-year revision share moves from 4.10% to 5.79% against 7.01% observed,
  while January starts and exact continuations remain poorly reproduced;
- shared baseline start-month normalization and exact nesting --- **passed**;
  equal probabilities reproduce the 18-parameter model exactly and the
  calendar-weighted duration masses sum to one for every interview month;
- baseline start-month multistart and recovery --- **passed provisionally**;
  three empirical conditional starts and distant simulated starts meet, and a
  2,000-record joint recovery gives month-distribution RMSE about 0.0044;
- baseline start-month predictive adequacy --- **passed for calendar months,
  failed for serial dependence**; 11 of 12 month shares and the whole-year
  revision rate are reproduced, but exact continuations remain too low and
  other revisions too high; and
- full 29-parameter observed-information inference --- **pending**; the
  gross-only January-heaping estimate is effectively on the boundary and
  should not receive a regular Wald interpretation;
- exact-anchor retention nesting and conditional identification --- **passed**;
  zero reproduces the flexible baseline likelihood exactly, the empirical
  profile has a clear interior maximum, and 20 conditional recovery samples
  recover the generating probability with small mean bias;
- exact-anchor predictive adequacy --- **failed but improved**; exact
  continuations rise from 60.6% to 61.8% against 64.7% observed and other
  revisions fall from 32.4% to 31.4% against 28.3% observed;
- longer 30-parameter multistart refinement --- **passed provisionally**;
  low- and high-retention screens are decisively inferior, a second joint
  block adds only 1.71 likelihood points, and the final conditional polish adds
  0.001 point, although the capped joint blocks retain convergence code 1;
- full 30-parameter recovery --- **passed for estimation and headline
  transition rates, inconclusive for the reporting decomposition**; three
  5,000-record samples give entry- and exit-rate RMSEs of about 0.12 and 0.11
  percentage points, while
  retention and the two revision probabilities remain noisy and offset one
  another;
- full-rank OPG analytical inference --- **passed provisionally**; the exact
  retention SE is 0.574 percentage points, but the information matrix is
  ill-conditioned, retention correlates 0.885 with tenure contamination, and
  gross-only January heaping is a boundary quantity; and
- exact central observed-information inference --- **pending**, with 48 of
  1,801 likelihood points cached;
- local-revision normalization and exact nesting --- **passed**; the
  one-to-six-month kernel renormalizes at zero tenure, excludes zero from its
  support, and setting its probability to zero reproduces the exact-anchor
  log likelihood exactly;
- local-revision conditional identification and multistart screening ---
  **passed provisionally**; the profile has a clear interior maximum, low and
  high starts are inferior, and central reporting-block refinement improves
  the preferred basin, although the deliberately capped blocks retain
  convergence code 1; and
- local-revision predictive adequacy --- **failed but substantially
  improved**; the predicted one-to-six-month share rises from 3.0% to 10.8%
  against 11.4% observed, and the likelihood gain is concentrated in records
  with local changes, but exact and residual-revision moments remain outside
  the simulation intervals;
- shared-reliability exact nesting and dispersed starts --- **passed
  provisionally**; zero dispersion reproduces the independent likelihood,
  all starts improve on it, and the preferred basin is clearly superior,
  although the short joint blocks terminate at their iteration caps;
- shared-reliability predictive adequacy --- **failed but informative**; the
  local-revision and zero/one-irregularity predictions improve, but the model
  overpredicts two irregularities and leaves several duration moments outside
  their simulation intervals; and
- cross-clock reliability interpretation --- **limited by the three-wave
  design**; within-clock clustering identifies the dispersions, but the data
  do not directly reveal repeated tenure and repeated timegap clocks for the
  same person;
- separate-dispersion joint convergence --- **passed**; broad reporting and
  transition blocks materially improve the initial fit, two cached joint
  continuations add 78.9 and 49.4 likelihood points, and a third returns the
  identical parameter vector with convergence code zero;
- fair common-dispersion refinement --- **passed**; projecting the converged
  nuisance block into the restricted model, polishing the common shift, and
  checking the full restricted score reduces the equality-restriction LR to
  144.6 without altering the transition rates;
- separate-dispersion recovery --- **passed provisionally**; one 10,000-person
  conditional recovery sample gives the same optimum from three starts and
  good latent-class discrimination, but more replications and full-nuisance
  recovery are still required;
- separate-dispersion OPG inference --- **passed numerically but limited for
  the restriction contrast**; the 33-parameter matrix is positive definite
  and full rank, but has condition number about 55 million, one hazard
  direction needs a smaller perturbation, and its Wald approximation conflicts
  with the directly optimized LR test; and
- separate-dispersion nuisance-adjusted profile inference --- **passed**; all
  retained points converge, two-start screens select the higher-likelihood
  warm branch on both sides, and the 95% profile interval for the
  timegap-minus-tenure dispersion is about 0.1100 to 0.1520, excluding the
  equality restriction and resolving the OPG/LR conflict in favor of the
  direct likelihood evidence; and
- separate-dispersion predictive adequacy --- **failed**; exact continuations
  improve modestly, but local revisions, residual revisions, timegap
  inconsistencies, and zero/one/two-irregularity shares remain outside the
  simulation intervals.

# 8. Development log

| Date | Change | Evidence or decision |
|---|---|---|
| `r format(Sys.Date(), '%Y-%m-%d')` | Added start-date reconstruction and A/B/C audit | Current-job start dates are persistent but frequently revised; January and whole-year patterns are prominent; strict job resets are visible. |
| `r format(Sys.Date(), '%Y-%m-%d')` | Selected the first model extension | Add a one-parameter employed-to-employed job-change mixture before enriching the reporting-error or AR(2) processes. |
| `r format(Sys.Date(), '%Y-%m-%d')` | Implemented and estimated the job-change mixture | The joint fit converged at (q=`r sprintf('%.4f', job_comparison$job_change_prob[2])`); fit improves sharply and estimated gross tenure contamination declines. |
| `r format(Sys.Date(), '%Y-%m-%d')` | Re-estimated on Panels B and C | The reset probability is stable in B and somewhat lower in C, while all three contamination estimates decline with stricter matching. |
| `r format(Sys.Date(), '%Y-%m-%d')` | Completed numerical validation | The preferred mode is stable and tightly profiled, but full-model recovery fails because the continuous likelihood mixes duration components with incompatible units. |
| `r format(Sys.Date(), '%Y-%m-%d')` | Replaced continuous densities with monthly masses | The discrete likelihood normalizes, substantially improves recovery, and estimates a Panel A reset probability near 0.97%. |
| `r format(Sys.Date(), '%Y-%m-%d')` | Repeated discrete-model recovery | Six fits at two sample sizes converge; bias and RMSE contract sharply at 40,000 observations and dispersed starts do not find a better mode. |
| `r format(Sys.Date(), '%Y-%m-%d')` | Completed discrete-model inference | The observed information is full rank and positive definite; analytical and nuisance-adjusted profile intervals agree closely. |
| `r format(Sys.Date(), '%Y-%m-%d')` | Re-estimated the corrected model on Panels B and C | All six multistart fits converge; reset and contamination estimates decline as matching becomes stricter, but the reset process remains present. |
| `r format(Sys.Date(), '%Y-%m-%d')` | Estimated persistent gross start-date reporting | The extension nests exactly and can recover positive persistence in simulation, but Panel A places the MLE at zero and leaves all substantive rates unchanged. |
| `r format(Sys.Date(), '%Y-%m-%d')` | Estimated January-heaped gross reporting | The normalized one-parameter extension passes recovery checks, all three Panel A starts agree near 24%, fit improves sharply, and the main transition estimates change little. |
| `r format(Sys.Date(), '%Y-%m-%d')` | Added analytical inference for January heaping | The 16-parameter observed-information matrix is full rank and positive definite; the main rates and heaping parameter are precise. |
| `r format(Sys.Date(), '%Y-%m-%d')` | Estimated correlated whole-year date revisions | The normalized extension passes recovery and multistart checks; its empirical implications are reported above. |
| `r format(Sys.Date(), '%Y-%m-%d')` | Completed revision-model inference and predictive diagnostics | Local information and profile inference are regular, but the fitted model does not reproduce the calendar moments and shifts likelihood away from non-target patterns. |
| `r format(Sys.Date(), '%Y-%m-%d')` | Allowed gross reports to revise preceding clean dates | Three starts converge near a 9.6% clean-anchor revision rate; fit and the targeted predictive moment improve sharply, but missing baseline start-month seasonality remains. |
| `r format(Sys.Date(), '%Y-%m-%d')` | Estimated a shared flexible start-month baseline | The likelihood improves by about 42,982 points, January receives 26.46% baseline mass, gross-only January heaping collapses to zero, and transition estimates remain stable. |
| `r format(Sys.Date(), '%Y-%m-%d')` | Validated the start-month extension | Distant starts meet and joint recovery succeeds after a one-sided gradient safeguard; calendar predictions pass, while remaining failures isolate serial report dependence. |
| `r format(Sys.Date(), '%Y-%m-%d')` | Added exact reported-start-date retention | The nested channel is strongly supported and conditionally recoverable; it improves exact-continuation predictions without materially changing entry or exit rates, but does not remove the full serial-reporting gap. |
| `r format(Sys.Date(), '%Y-%m-%d')` | Refined and jointly validated exact retention | Longer multistart work stabilizes the empirical point estimate near 20%; three full re-estimations recover pooled transitions reasonably well but reveal weak finite-sample separation among retention and revision probabilities. |
| `r format(Sys.Date(), '%Y-%m-%d')` | Added analytical inference diagnostics | The efficient OPG approximation is full rank and gives tight errors for headline transitions, while its condition number, reporting-channel correlations, and boundary heaping estimate document the remaining limitations. |
| `r format(Sys.Date(), '%Y-%m-%d')` | Estimated normalized local start-date revisions | The conditional probability is about 33.8%, multistart reporting blocks favor the same basin, and the predicted local-revision share moves from 3.0% to 10.8% against 11.4% observed without materially changing entry or exit rates. |
| `r format(Sys.Date(), '%Y-%m-%d')` | Added shared person-level duration-reporting reliability | The one-parameter equal-weight mixture strongly improves fit and duration-pattern clustering while leaving transition estimates stable, but its common cross-clock dispersion is a maintained restriction and predictive gaps remain. |
| `r format(Sys.Date(), '%Y-%m-%d')` | Relaxed equal tenure and timegap reliability dispersion | Symmetric restricted/unrestricted refinement converges at dispersions 2.09 and 2.22 with LR 144.6; transition rates remain stable, conditional recovery succeeds, and full-nuisance OPG errors are informative for headline rates but unreliable for the restriction contrast. |
| `r format(Sys.Date(), '%Y-%m-%d')` | Profiled the clock-specific dispersion difference | Multistart nuisance-adjusted profiling gives a 95% interval of about 0.1100 to 0.1520 for timegap minus tenure dispersion, decisively excluding equality and resolving the conflict with the ill-conditioned OPG Wald interval. |

\clearpage

# 9. Four-wave AR(1) bridge to the AR(2) extension

The next substantive extension is employment-history dependence, rather than
further refinement of the separation between duration-reliability classes.
The four-wave bridge retains the preferred three-wave measurement model:
discrete-month tenure, piecewise duration hazards, current-job resets,
calendar and local reporting revisions, joint timegap contamination, and
equal prior weights for the two reliability classes. No additional employment
heterogeneity class has been introduced.

`EM-tenure/R/four_wave_eps.R` sums over all 16 four-wave latent employment
histories. `EM-tenure/estimate_monthly_four_wave_ar1.R` prepares Panel A,
caches incremental optimisation blocks, and constructs a waves 1--3
comparison using exactly the people and weights in the four-wave sample.
The initial employment probability remains free. The likelihood is estimated
directly, not with the old three-wave EM sufficient-statistic updates.

## Sample and estimates

**Subsequent validation warning:** Section 10 identifies a timegap likelihood
normalization defect, repaired in Section 11. The estimates below remain
saved, numerically converged pre-repair results. Section 12 reports the
separate empirical refit under the repaired likelihood when validated.

There are 317,437 records with four observed employment statuses and 316,889
with the required state-appropriate duration reports (548 exclusions).
The likelihood collapses these to 101,916 distinct observation cells without
changing their weighted contribution. Zero-month tenure is retained;
unreported-state durations remain missing. Never-worked reports follow the
same recoding convention as in the three-wave model. Interview months remain
nominal quarter-end months, not recovered exact interview dates.

```{r four-wave-bridge-results, echo=FALSE, results='asis'}
bridge_dir <- file.path("..", "EM-tenure", "output", "results",
  "job_change_monthly", "four_wave_ar1")
if (!dir.exists(bridge_dir)) bridge_dir <- file.path("EM-tenure", "output",
  "results", "job_change_monthly", "four_wave_ar1")
bridge_path <- file.path(bridge_dir, "stage_summary_latest.csv")
if (file.exists(file.path(bridge_dir, "convergence_summary_latest.csv")))
  bridge_path <- file.path(bridge_dir, "convergence_summary_latest.csv")
if (file.exists(bridge_path)) {
  bridge <- read.csv(bridge_path)
  selected <- if ("independently_converged" %in% names(bridge))
    seq_len(nrow(bridge)) else unique(c(1L, which.max(bridge$loglik)))
  bridge_fields <- c("entry_rate", "exit_rate", "status_misclassification",
    "initial_employment", "loglik")
  bridge_table <- data.frame(Measure = c("Entry rate", "Exit rate",
    "Misclassification rate", "Initial employment", "Log likelihood"))
  bridge_titles <- c(preliminary="Preliminary", exact1="Refined",
    start_minus="Start (-)", start_plus="Start (+)", start="Inherited")
  for (j in selected) {
    title <- if (bridge$stage[j] %in% names(bridge_titles))
      unname(bridge_titles[bridge$stage[j]]) else bridge$stage[j]
    values <- as.numeric(unlist(bridge[j, bridge_fields], use.names=FALSE))
    bridge_table[[title]] <- c(sprintf("%.5f", values[1:4]),
      formatC(values[5], format="f", digits=2, big.mark=","))
  }
  print(knitr::kable(bridge_table, digits=5,
    caption="Four-wave AR(1): numerical refinement and starting-point checks"))
} else cat("Four-wave optimisation checkpoints are not yet consolidated.\n")
```

The preliminary column is an unconverged checkpoint. All three completed
refinement runs have optimizer code zero and final maximum projected central
scores between $1.5\times10^{-6}$ and $3.0\times10^{-6}$ per observation,
below the declared $10^{-5}$ criterion.
The optimizer's scaled tolerance is chosen to imply this unscaled criterion;
the latter is checked separately after termination. Start (+) has the highest
log likelihood and is selected in `converged_comparison_latest.rds`. The
three likelihoods differ by at most 0.056; entry, exit, and status-error
rates differ by less than 0.00004 percentage points. This supports numerical
stability of these rates, not precise identification of every reporting
parameter.

Entry and exit rates pool the three quarterly intervals, weighting each
individual-interval by the survey weight and posterior probability of the
relevant origin state. Misclassification is the common employment-status
error probability, not a duration-contamination probability. The reliability
mixture's tenure and timegap midpoint parameters should not be interpreted
as mean contamination probabilities.

Relative to the preliminary four-wave point, quarterly entry rises from
2.446% to 2.603%, exit rises slightly from 4.033% to 4.060%, status
misclassification falls from 2.822% to 2.765%, and initial employment is
essentially unchanged (48.685% versus 48.705%). The log likelihood improves
by about 665.8. This is a numerical improvement within the **same model and
sample**, not a test of a new specification. The preliminary fit understated
entry more than exit; the broad finding of substantial misclassification
remains. Likelihood levels on different three-wave and four-wave samples
must not be compared as an LR test. The optional same-person three-wave
re-estimation is omitted at the author's request.

## Checks and outstanding work

The continuation now has a separately validated compiled evaluator
(`four_wave_fast.R` and `src/four_wave_monthly.cpp`). It uses the reference R
probability tables and caches repeated spell and reporting-pattern calculations;
no likelihood terms or model restrictions are changed. On the saved full sample,
the log-likelihood difference is less than $10^{-7}$ and maximum posterior
differences are below $10^{-13}$. Perturbed-parameter, zero-status-error,
disabled-revision, and serial/parallel comparisons also pass. A benchmark using
eight threads takes about 1.3 seconds per full likelihood evaluation, rather
than roughly a minute for the earlier R implementation.

`refine_four_wave_ar1_fast.R` jointly refines all 33 parameters, retains
improving checkpoints, and checks a central-difference projected score after
the optimizer stops. The diagonal of the saved three-wave information matrix
is used only for numerical parameter scaling; it is not a substitute for
four-wave standard errors and requires no three-wave re-estimation.
The same-person three-wave re-estimation has been dropped from the work plan
at the author's request.

For the final continuation, category-averaged and missing-duration risks are
calculated with exact piecewise-exponential integrals. The integral splits at
the duration-hazard knots and their quarter-shifted counterparts. Fifty tests
against independent numerical integration and limiting cases pass. At the
saved preliminary point this changes the total log likelihood by about
$-0.000037$ and the largest posterior probability by less than $2\times10^{-8}$;
its purpose is to remove amplified quadrature noise from the numerical score,
not to change the model. The three-wave estimator is not modified. Central
finite differences are used for the final joint refinement.

- Tests pass for all 16 histories, posterior normalization, the zero-status-error
  limit, three-wave joint-timegap equivalence, and four-wave timegap probability
  normalization. Preparation tests cover zero tenure, never-worked recoding,
  missing flags, and exclusion of invalid state-appropriate durations.
- Sharing and compressing repeated job-segment calculations reproduces the
  saved full-sample likelihood and posterior probabilities exactly. This is
  a computational change, not a model change.
- Accepted bounded optimisation steps are flagged as non-converged. Early
  checkpoints with a zero code are corrected when consolidated; step
  acceptance is not evidence that the likelihood has been maximized.
- All three four-wave runs pass optimizer and central-score convergence
  checks. The full-sample likelihood and posterior agree with the R reference
  at the fitted parameters. Starting-point checks are saved separately in
  `multistart_summary_latest.csv`; all fit objects and diagnostics are retained.
- The two nearby starts perturb all 33 transformed parameters by
  $+0.05\sin(j)$ and $-0.05\sin(j)$, respectively, from the saved
  `start_multistart_exact.rds` point. Continuations resume saved checkpoints
  without applying the perturbation again.
- Nearby-start agreement is a numerical stability check, not proof of a global
  optimum or strong identification of the reliability-class separation.
- The first four-wave simulation gate fails (Section 10); Section 11 records
  the repair and renewed validation. Analytical standard errors and the
  duration-model AR(2) extension have not been estimated.

# 10. Four-wave simulation-recovery gate

## Changes and scope

- Extended the existing sequential simulator to four waves, retaining its
  three-wave defaults and adding the exact piecewise-risk option.
- Generated three independent samples of 20,000 records at the selected
  Panel A four-wave parameters (seeds 260904--260906). These are simulations,
  not additional empirical panels or bootstrap replications.
- Checked the mean record-level score at the known truth before fitting.
  Its Monte Carlo standard error is the sample standard deviation of the
  score divided by the square root of the simulated sample size.
- Added exhaustive probability-mass and timegap-clock checks. No empirical
  fit or likelihood formula was changed, and no simulation recovery fit was
  run after these gates failed.

## Results

The short-duration entry-hazard score is negative in all three samples
(individual standardized scores -5.18, -5.26, and -5.50). The pooled mean is
-0.01225 with Monte Carlo standard error 0.001331: a standardized discrepancy
of -9.20. This fails the conservative screening threshold of five Monte Carlo
standard errors. It establishes a generator/likelihood mismatch, not a
precisely quantified bias in the empirical entry rate.

```{r four-wave-recovery-score, echo=FALSE, results='asis'}
recovery_dir <- file.path("..", "EM-tenure", "output", "results",
  "job_change_monthly", "four_wave_ar1", "recovery")
if (!dir.exists(recovery_dir)) recovery_dir <- sub("^../", "", recovery_dir)
score_file <- file.path(recovery_dir, "sequential_score_summary_latest.csv")
if (file.exists(score_file)) {
  score <- read.csv(score_file)
  fields <- c("log_hd1", "log_hd3", "pi", "alpha")
  score <- score[match(fields, score$parameter), ]
  shown <- data.frame(Parameter=c("Entry log-hazard: 0--3 months",
    "Entry log-hazard: 1--3 years", "Status-error logit",
    "Initial-employment logit"),
    `Mean score`=score$mean_score, `Monte Carlo SE`=score$mcse,
    `Standardized score`=score$z_score, check.names=FALSE)
  print(knitr::kable(shown, digits=c(0,6,6,2),
    caption="Generator/likelihood score check: 60,000 simulated records"))
}
```

An independent finite-support check exposes a defect in the likelihood itself.
The illustrative nested case uses constant tenure hazards of 20 per year,
uniform start months, no job resets or serial tenure-revision channels,
initial employment 0.5, tenure contamination 0.2, and timegap hazards
either constant at 0.4 or equal to (4, 1, 0.5, 0.3, 0.2). It enumerates all
four-wave observed status/duration combinations with tenure 0--24 months,
including all seven timegap categories: 1,048,576 cells.

```{r four-wave-recovery-mass, echo=FALSE, results='asis'}
mass_file <- file.path(recovery_dir, "probability_mass_m24.csv")
if (file.exists(mass_file)) {
  mass <- read.csv(mass_file)[c(7,6,4,8), ]
  shown <- data.frame(`Duration-dependent entry`=ifelse(mass$varying_hazard,"Yes","No"),
    `Status error`=mass$status_error,
    `Timegap error`=ifelse(mass$timegap_error<1e-7,"Approximately zero","0.30"),
    `Total probability`=sprintf("%.9f",mass$probability_mass),check.names=FALSE)
  print(knitr::kable(shown,caption="Exhaustive normalization check in a nested example"))
}
```

When duration dependence, status errors, and timegap contamination are all
present, total mass is **0.999327180679**, rather than one. Increasing the
tenure cutoff from 12 to 24 months leaves the discrepancy essentially
unchanged. A conservative omitted-tail bound at 24 months is below
$1.7\times10^{-10}$; it cannot explain the missing mass of about 0.000673.
The R reference and compiled implementation give the same result at the
12-month cutoff. Constant-hazard and near-zero-error controls sum to one
within numerical/tail tolerance. This is a counterexample to normalization,
not an estimate of the defect's size at the empirical parameters.

## Diagnosis and interpretation

- **Future-report-dependent fallback:** the four-wave E-step uses a joint
  timegap likelihood only when an entire latent nonemployment spell is
  observed as nonemployment. Otherwise it falls back to adjacent-pair terms.
  Thus a fourth reported employment status can change the likelihood factor
  for the first three timegap reports. In the illustrative case, the joint
  three-report distribution and pairwise fallback differ by 0.2286 in total
  variation. Individually normalized components do not guarantee a normalized
  full likelihood when transition risks also depend on the reports.
- **A separate simulator mismatch:** the simulator retains a continuous
  nonemployment clock and ages it across waves. The likelihood instead
  multiplies category-to-category transition probabilities, effectively
  discarding within-category information from earlier reports. Exact
  interval-intersection probabilities and that category-Markov approximation
  differ even with negligible timegap contamination (total variation 0.02143
  in the same illustrative hazard case).
- These findings are about probability construction, not merely weak
  identification of the two reliability classes. They do not establish how
  much the empirical entry, exit, or misclassification rates will change.
  Earlier numerical convergence and implementation-equivalence checks still
  hold, but the four-wave results are not yet validated for interpretation.
  Earlier normalization checks covered individual emissions and posterior
  weights, not the sum of the complete joint probability over observable
  records; those are different tests.

The repair should use one joint timegap-spell likelihood for every observed
subset, integrating missing or contaminated reports and retaining the clean
clock's interval restrictions across all waves. The simulator must implement
that same probability law. Re-run normalization and score checks first; only
then run the small parameter-recovery fits. This repair was deferred at the
initial validation stage and subsequently authorized; see Section 11.

Reproduction scripts are `validate_four_wave_simulation.R`,
`audit_four_wave_probability_mass.R`, and `summarise_four_wave_recovery.R`.
For the mass audit, use `FOUR_WAVE_MASS_MAX_MONTH=24` for the tight-tail check;
use 12 with `FOUR_WAVE_MASS_REFERENCE=true` for the independent R comparison.
Set `FOUR_WAVE_TIMEGAP_CLOCK=legacy_pairwise` to reproduce these historical
failures. The consolidated `recovery/recovery_status_latest.rds` records the
failed gates, zero recovery fits, and unchanged empirical estimates.

# 11. Repaired continuous timegap clock and recovery

## Changes

- Replaced the four-wave joint/pairwise fallback with one joint timegap
  likelihood for every latent nonemployment spell and every observed subset.
- Each clean report restricts the same spell-start duration. Missing reports
  are integrated out, and all contamination patterns are summed. Information
  from earlier categories is retained rather than discarded at each wave.
- Applied the same calculation in the R reference and compiled evaluator.
  The existing sequential simulator already retains and ages this clock;
  its four-wave exact-risk option now matches the repaired likelihood.
- Kept the initial timegap distribution, transition-risk formulas, tenure
  reporting model, and three-wave estimator unchanged. In particular, this
  repair does not introduce a new within-quarter distribution for timegap at
  a nonemployment spell's onset. That remains a separate modelling question.
- Preserved legacy evaluation explicitly as `legacy_pairwise`; repaired
  four-wave evaluation defaults to `continuous_joint`. Historical fit objects
  and failed diagnostics are not overwritten.

For a clean report in category interval $[a_t,b_t)$ at offset $t$, the
spell-start clock satisfies $D_0\in[a_t-t\Delta,b_t-t\Delta)$. Its joint
clean-report probability is

$$
P\left\{\max(0,\max_t(a_t-t\Delta))\le D_0
<\min_t(b_t-t\Delta)\right\},\qquad \Delta=0.25.
$$

An empty intersection has probability zero. Gross reports instead contribute
their marginal category probabilities; unobserved reports impose no
restriction. Mixture weights retain the existing clock-specific reliability
classes. This enforces clean-clock consistency across all observed waves
without requiring every status report to be correct.

## Probability and score checks

All eight exhaustive probability cases now sum to one within numerical
tolerance, including the former failing case. Summing out any report gives
the same joint distribution for the remaining reports. Fifty-six new unit
checks cover observed subsets, boundary contamination probabilities, interval
intersections, and missing early reports. The repaired full-sample compiled
likelihood agrees with R within $4.6\times10^{-8}$; maximum posterior
differences are below $1.3\times10^{-14}$.

Repeating the same three 20,000-person simulation samples removes the score
failure: the short-duration entry-hazard discrepancy falls from -9.20 to
+1.86 Monte Carlo standard errors. No pooled score among the 33 parameters
exceeds 1.87 in absolute value. These checks support the repaired probability
law; they do not by themselves establish finite-sample recovery or precision.

```{r four-wave-clock-repair-checks, echo=FALSE, results='asis'}
repaired_dir <- file.path(dirname(recovery_dir),"recovery_continuous_clock")
new_score_file <- file.path(repaired_dir,"sequential_score_summary_latest.csv")
if (file.exists(new_score_file)) {
  old <- read.csv(file.path(recovery_dir,"sequential_score_summary_latest.csv"))
  new <- read.csv(new_score_file)
  keep <- c("log_hd1","log_hd3","pi","alpha")
  comparison <- data.frame(Parameter=c("Entry log-hazard: 0--3 months",
    "Entry log-hazard: 1--3 years","Status-error logit","Initial-employment logit"),
    Legacy=old$z_score[match(keep,old$parameter)],
    Repaired=new$z_score[match(keep,new$parameter)])
  print(knitr::kable(comparison,digits=2,
    caption="Standardized score at the generating parameters: before and after repair"))
}
```

## Small full-parameter recovery exercise

`estimate_four_wave_recovery.R` estimates all 33 parameters on two independent
5,000-person samples, using two nearby starting points for the first sample
and one for the second (seeds 261004 and 261005). It uses the same bounded
likelihood optimization and a separate projected central-score tolerance of
$10^{-5}$. The saved empirical scaling vector is only a numerical
preconditioner. No nuisance parameter is fixed at its generating value.

```{r four-wave-clock-recovery-fits, echo=FALSE, results='asis'}
fits_file <- file.path(repaired_dir,"fit_summary_latest.csv")
if (file.exists(fits_file)) {
  recovery_fits <- read.csv(fits_file)
  for (rep in unique(recovery_fits$replication)) {
    fits <- recovery_fits[recovery_fits$replication==rep,]
    tab <- data.frame(Measure=c("Entry rate","Exit rate","Misclassification rate",
      "Initial employment"),
      Generating=c(fits$entry_truth[1],fits$exit_truth[1],
        fits$pi_truth[1],fits$alpha_truth[1]))
    for (j in seq_len(nrow(fits))) tab[[fits$start[j]]] <-
      c(fits$entry_estimate[j],fits$exit_estimate[j],fits$pi_estimate[j],fits$alpha_estimate[j])
    print(knitr::kable(tab,digits=5,caption=paste("Simulation recovery: sample",rep)))
  }
  print(knitr::kable(recovery_fits[c("replication","start","n","converged","score","loglik")],
    digits=c(0,0,0,0,8,3),caption="Recovery convergence diagnostics"))
} else cat("Full-parameter recovery fits are not yet consolidated.\n")
```

The generating transition rates use the true parameters and their posterior
risk weights on the same simulated sample, so differences isolate parameter
estimation rather than differences in sample composition. Full hazard
coefficients are saved in `coefficient_recovery_latest.csv` and shown below;
fit objects also
retain all reporting parameters and numerical diagnostics. This is a small
implementation/recovery check, not a large Monte Carlo bias study or a
bootstrap standard-error exercise. Empirical Panel A rates are not updated
by these simulated fits.

All three fits converge with maximum projected central scores between
$2.54\times10^{-6}$ and $2.93\times10^{-6}$. The two starts for sample 1 differ
by less than 0.001 log-likelihood units and 0.00013 percentage points in the
main rates. Across the two samples, entry, exit, and status misclassification
are each within 0.16 percentage points of their generating values. Initial
employment differs by -1.10 and +0.16 percentage points, respectively.
The high short-duration entry hazard and lower long-duration hazards remain
visible in the recovered coefficients. These results support the repaired
implementation and approximate recovery of economically relevant quantities;
two samples do not establish finite-sample unbiasedness, precise reliability
separation, or a globally unique optimum.

```{r four-wave-clock-recovery-hazards, echo=FALSE, results='asis'}
coef_file <- file.path(repaired_dir,"coefficient_recovery_latest.csv")
if (file.exists(coef_file) && file.exists(fits_file)) {
  coef <- read.csv(coef_file)
  fits <- read.csv(fits_file)
  keep <- c(paste0("entry_hazard_",1:5),paste0("exit_hazard_",1:5))
  bands <- c("0--3 months","3--12 months","1--3 years","3--5 years","5+ years")
  tab <- data.frame(Hazard=c(paste("Entry:",bands),paste("Exit:",bands)),
    Generating=coef$truth[match(keep,coef$parameter)])
  for (rep in unique(fits$replication)) {
    candidates <- fits[fits$replication==rep & fits$converged,]
    if (!nrow(candidates)) next
    label <- candidates$start[which.max(candidates$loglik)]
    selected <- coef[coef$replication==rep & coef$start==label,]
    tab[[paste("Sample",rep)]] <- selected$estimate[match(keep,selected$parameter)]
  }
  print(knitr::kable(tab,digits=4,
    caption="Duration-hazard recovery (annual hazards, not quarterly transition probabilities)"))
}
```

New outputs are under `four_wave_ar1/recovery_continuous_clock`; the failed
legacy checks remain under `four_wave_ar1/recovery`. The empirical continuation
defaults to a new `clock_joint1` label and rejects stale-source checkpoints.

# 12. Repaired four-wave AR(1): empirical Panel A refit

## Changes and scope

- Re-estimate all 33 parameters under the repaired continuous-clock likelihood
  on the same 316,889 Panel A records (101,916 collapsed likelihood cells).
- Start at the best converged pre-repair fit, preserve the old fit, and use a
  separate output label. Require optimizer convergence and an independent
  projected central score no greater than $10^{-5}$ per unit of sample weight.
- Keep the duration hazards, tenure reporting specification, fixed equal
  reliability-class shares, and free initial employment probability unchanged.
  Do not re-estimate the three-wave comparator or introduce AR(2) in this step.
- Compare quarterly entry and exit probabilities using the same pooled
  posterior-origin-risk and survey weights. Misclassification refers to
  employment-status error, not duration contamination.

```{r four-wave-clock-empirical, echo=FALSE, results='asis'}
empirical_dir <- file.path(dirname(recovery_dir),"continuous_clock_empirical")
empirical_file <- file.path(empirical_dir,"comparison_latest.rds")
if (file.exists(empirical_file)) {
  empirical <- readRDS(empirical_file)
  stopifnot(isTRUE(empirical$best$converged),
    isTRUE(empirical$best$independent_score_pass))
  cmp <- empirical$comparison
  tab <- data.frame(Measure=c("Entry rate","Exit rate","Misclassification rate",
    "Initial employment"), `Pre-repair`=100*cmp$legacy,
    Repaired=100*cmp$repaired, `Change (pp)`=cmp$change_percentage_points,
    check.names=FALSE)
  print(knitr::kable(tab,digits=3,
    caption="Panel A four-wave AR(1): main rates (percent)"))
  cat(sprintf("\nRepaired log likelihood: %.3f; maximum projected central score: %.3g.\n",
    empirical$best$loglik,empirical$best$gradient_max))
  cat(sprintf("\nRe-optimization improves the repaired objective by %.3f relative to the old parameter vector.\n",
    empirical$repaired_objective_gain))
} else cat("The empirical refit has not yet been consolidated; no revised rates are reported here.\n")
```

```{r four-wave-clock-empirical-hazards, echo=FALSE, results='asis'}
if (file.exists(empirical_file)) {
  print(knitr::kable(empirical$hazards,digits=4,
    col.names=c("Hazard","Duration","Pre-repair","Repaired"),
    caption="Panel A duration coefficients: annual hazards"))
}
```

## Findings and interpretation

- The repaired empirical fit converges with optimizer code zero and a
  separately recomputed maximum projected central score of
  $5.53\times10^{-6}$. All 33 parameters were optimized from the best
  converged legacy starting vector. This is one empirical starting point,
  not a new dispersed-start exercise.
- Entry rises from 2.603% to 2.693% per quarter; exit rises from 4.060% to
  4.092%. Employment-status misclassification falls from 2.765% to 2.646%,
  and initial employment rises from 48.705% to 49.108%.
- Thus the repair changes the estimates modestly rather than overturning
  their main interpretation. Estimated status misclassification remains
  material, while exit probability conditional on employment remains above
  entry probability conditional on nonemployment. Those probabilities have
  different populations at risk and should not be directly subtracted to
  infer the change in employment.
- The largest duration-hazard adjustment is in the shortest nonemployment
  band: its annual entry hazard rises from about 0.509 to 0.563. Longer-duration
  entry hazards remain substantially lower; exit hazards change comparatively
  little. These are annual hazard coefficients, not the weighted quarterly
  entry probabilities in the main comparison.

The old likelihood is retained for historical auditing, not as a nested-model
benchmark: changing a defective probability construction means that its
difference from the repaired likelihood is not an LR statistic. The gain
reported above compares two parameter vectors under the *same repaired*
objective. Neither convergence nor the recovery checks establish strong
identification of the two reliability dispersions or a globally unique optimum.
No analytical or bootstrapped standard errors are estimated in this step.

At the new fitted parameter vector, the independent full-sample R evaluator
agrees with the compiled likelihood within $4.52\times10^{-8}$ and with its
posteriors within $1.24\times10^{-14}$. The repaired log likelihood is
-2,894,475.050, a gain of 414.246 over evaluating the old parameter vector
under the repaired law. The focused regression suite passes 249 checks.

Execution note: a trailing parsing error was reported after the optimizer,
independent score check, fit save, and rate printing had completed; the runner
had been edited while active. The current runner parses cleanly. A fresh
process loaded the saved fit and successfully performed the independent
full-sample checks and consolidation above. The error did not require a new
estimation or affect the saved parameter estimates.

Run `refine_four_wave_ar1_fast.R` with label `clock_joint1`, using
`fit_fast_start_plus_latest.rds` as the start, then run
`summarise_four_wave_clock_empirical.R`. The collector verifies the full-sample
R reference at the new parameters and stores rates, annual hazards, all
transformed coefficients, posteriors, convergence diagnostics, and source/input
fingerprints in `four_wave_ar1/continuous_clock_empirical`. These files do not
replace the legacy results or the separate simulation-recovery outputs.

# 13. Four-wave duration model with two AR(2) return effects

A compact candidate paper Table 9, including history-specific entry and exit
probabilities, is provided in `four_wave_duration_results.Rmd`. Rebuild its
LaTeX and unrounded CSV snapshot with
`scripts/build_four_wave_duration_paper_table.R`; this reads the saved fits
without re-estimation. Existing paper Tables 1--8 are unchanged. Analytical
uncertainty and the questionnaire-motivated sensitivity checks remain pending.

## Changes and identifying restrictions

- Add one entry and one exit coefficient to the repaired Panel A model;
  re-estimate all 35 parameters jointly. Keep duration-reporting mechanisms,
  duration reliability classes, and the estimation sample unchanged.
- For transitions 2-to-3 and 3-to-4, shift the log odds of a transition when
  the current latent state differs from the preceding latent state. Positive
  coefficients mean an increased probability of returning to the prior state.
- Retain free initial employment and the AR(1) law for transition 1-to-2.
  This is an explicit initial-pair restriction, not an assertion of stationarity
  or an estimate of unobserved pre-panel history. A free initial-pair model is
  deferred to a sensitivity check.
- Use forward numerical gradients during optimization and an independent
  central score for acceptance. Every accepted fit must have optimizer code
  zero and maximum projected central score at most $10^{-5}$. Use a main
  start and two perturbed starts, with all nuisance parameters free.

Writing $S_t=1$ for employment and $r_{01,t},r_{10,t}$ for the existing
duration-dependent AR(1) probabilities, for $t=2,3$:

$$
\begin{aligned}
\operatorname{logit}P(S_{t+1}=1\mid S_t=0,\ldots)
 &=\operatorname{logit}(r_{01,t})+b_N\,1\{S_{t-1}=1\},\\
\operatorname{logit}P(S_{t+1}=0\mid S_t=1,\ldots)
 &=\operatorname{logit}(r_{10,t})+b_E\,1\{S_{t-1}=0\}.
\end{aligned}
$$

Thus $\exp(b_N)$ is the entry-odds multiplier after a recent exit and
$\exp(b_E)$ is the exit-odds multiplier after a recent entry, holding the
baseline duration-dependent probability fixed. These are **odds ratios, not
hazard ratios**. Apply them after the existing integration over category or
unobserved wrong-state durations. Do not alter the marginal duration-reporting
distributions or impose a new stationary duration distribution.

The implementation reuses the validated AR(1) likelihood through the exact
identity $L_2=L_1\sum_hP_1(h\mid\text{reports})P_2(h)/P_1(h)$. The AR(1)
likelihood and posterior are recomputed at every new nuisance vector: neither
is fixed during estimation. The independent R reference instead constructs
the AR(2) history prior directly before combining reporting probabilities.
Class and job-change posterior aggregates must also be recomputed under AR(2);
they cannot simply be copied from the AR(1) evaluator.

## Validation and recovery

Zero effects reproduce the full-sample repaired AR(1) likelihood and
posteriors. Nonzero effects agree with the independent direct-prior R
calculation. Four exhaustive finite-support joint-probability checks sum to
one, within the tenure-tail allowance. Three 10,000-person simulations check
all 35 likelihood scores at known nonzero coefficients. The largest pooled
absolute discrepancy is 2.18 Monte Carlo standard errors; the entry and exit
AR(2) discrepancies are 0.83 and -0.01. Twenty-four added unit checks cover nesting,
reference agreement, the seeded simulator, and which histories receive each
effect. These checks concern implementation, not precision of empirical effects.

```{r four-wave-ar2-recovery, echo=FALSE, results='asis'}
ar2_dir <- file.path(dirname(dirname(recovery_dir)),"four_wave_ar2")
ar2_recovery_file <- file.path(ar2_dir,"recovery/recovery_latest.rds")
if (file.exists(ar2_recovery_file)) {
  ar2_recovery <- readRDS(ar2_recovery_file)
  print(knitr::kable(ar2_recovery$summary[c("replication","start","converged",
    "ar2_entry_truth","ar2_entry","ar2_exit_truth","ar2_exit")],digits=4,row.names=FALSE,
    caption="AR(2) recovery with all 35 parameters free: two 5,000-person samples"))
} else cat("The full-parameter recovery exercise has not yet been consolidated.\n")
```

## Empirical comparison

All three full-parameter recovery fits converge, with independent projected
central scores below $4.4\times10^{-6}$. The two starts on sample 1 differ by
less than 0.0004 likelihood units and 0.00024 in either AR(2) coefficient.
The generating entry coefficient is 1.099; estimates are 1.245 and 0.964
across the two samples. The generating exit coefficient is 0.693; estimates
are 0.472 and 0.620. These are encouraging implementation checks, not a
large-sample bias or coverage study. After projecting out all nuisance-score
directions in the 30,000-person score simulation, the two-by-two information
matrix has eigenvalues 70.78 and 56.65: the two return effects have separately
varying scores in this experiment.

```{r four-wave-ar2-empirical, echo=FALSE, results='asis'}
ar2_file <- file.path(ar2_dir,"empirical/comparison_latest.rds")
if (file.exists(ar2_file)) {
  ar2_results <- readRDS(ar2_file)
  rows <- ar2_results$summary[match(c("AR1",ar2_results$best_label),ar2_results$summary$model),]
  labels <- c("Entry rate (%)","Exit rate (%)","Misclassification rate (%)",
    "Initial employment (%)","Log likelihood","N")
  column <- function(x) c(sprintf("%.3f",100*unlist(x[c("entry","exit","misclassification","initial_employment")])),
    formatC(x$loglik,format="f",digits=3,big.mark=","),format(x$n,big.mark=","))
  print(knitr::kable(data.frame(Measure=labels,AR1=column(rows[1,]),AR2=column(rows[2,])),
    row.names=FALSE,caption="Repaired four-wave duration model: AR(1) and AR(2), Panel A"))
  b <- unname(unlist(rows[2,c("ar2_entry","ar2_exit")]))
  print(knitr::kable(data.frame(Measure=c("Entry return effect (log odds)",
    "Exit return effect (log odds)","Entry return odds ratio","Exit return odds ratio"),
    AR1=c(0,0,1,1),AR2=c(b,exp(b))),digits=3,row.names=FALSE,
    caption="Return effects: both coefficients are fixed at zero in AR(1)"))
  cat(sprintf("\nLog-likelihood gain: %.3f; twice the gain: %.3f.\n",
    ar2_results$likelihood_gain,ar2_results$lr_statistic))
  print(knitr::kable(ar2_results$summary[-1,c("model","loglik","score","ar2_entry","ar2_exit")],
    digits=c(0,3,8,4,4),row.names=FALSE,caption="AR(2) starting-point and convergence checks"))
} else cat("No validated empirical AR(2) comparison has been consolidated yet.\n")
```

All main rates pool the three transitions using survey weights and posterior
origin-risk weights; only the last two transition probabilities contain an
AR(2) term. Initial employment is the estimated initial probability. The
misclassification parameter concerns employment status, not duration reports.
Unlike the historical pre-repair comparison, these two models are nested under
one repaired probability law. Any formal LR inference still requires regularity
and identification assumptions; the likelihood gain alone is not a validation
of causal state dependence.

```{r four-wave-ar2-implications, echo=FALSE, results='asis'}
if (file.exists(ar2_file)) {
  for (direction in c("Entry","Exit")) {
    rows <- ar2_results$implications[ar2_results$implications$transition==direction,]
    measures <- c("raw_reversal","not_confirmed","genuine_reversed","genuine_persistent")
    tab <- data.frame(Measure=c("Raw reversal","Not confirmed","Genuine, reversed","Genuine, persistent"),
      AR1=100*as.numeric(rows[rows$model=="AR1",measures]),
      AR2=100*as.numeric(rows[rows$model=="AR2",measures]))
    table_tex <- knitr::kable(tab,format="latex",booktabs=TRUE,digits=2,longtable=FALSE,
      label=paste0("four-wave-ar2-",tolower(direction),"-implications"),
      caption=paste("Posterior decomposition of apparent",tolower(direction),"transitions (\\%)"))
    table_tex <- sub("\\begin{table}","\\begin{table}[H]",table_tex,fixed=TRUE)
    cat("\n```{=latex}\n",table_tex,"\n```\n",sep="")
  }
}
```

This decomposition pools transitions 1-to-2 and 2-to-3, which have an observed
follow-up wave. A transition is confirmed only when the latent origin and
destination match its apparent direction. The three model-based categories partition
apparent transitions into not confirmed, genuine but reversed next wave, and
genuine and persistent next wave. All four reports inform the posterior; this
is a retrospective decomposition, not a forecast using just two reports.
"Not confirmed" does not distinguish misreporting from record-matching errors.

```{r four-wave-ar2-hazards, echo=FALSE, results='asis'}
if (file.exists(ar2_file)) print(knitr::kable(ar2_results$hazards,digits=4,
  caption="Baseline annual duration hazards: the AR(2) odds adjustment is additional"))
```

## Findings and interpretation

- All three empirical starts converge with independent projected central
  scores between $3.68\times10^{-6}$ and $7.70\times10^{-6}$. Their entry,
  exit, and status-misclassification rates differ by less than 0.00007
  percentage points, and their return coefficients by less than 0.00027.
  The highest-likelihood run is `plus`; its log likelihood exceeds repaired
  AR(1) by 2,196.404 (twice the gain is 4,392.807). The small likelihood
  spread across starts is 0.112, not evidence that every nuisance parameter
  is strongly identified or that a global optimum has been proved.
- The entry return coefficient is 1.299 and the exit return coefficient
  1.931: odds multipliers of 3.666 and 6.895, respectively. These duration
  controls do not absorb the tendency to return to the preceding state.
  Current-job tenure is not the same as uninterrupted employment duration
  when workers can change jobs while remaining employed; this is one reason
  surviving history effects need not contradict duration dependence.
- Average quarterly entry rises from 2.693% to 3.206%, and exit from 4.092%
  to 4.775%. Status misclassification falls modestly, from 2.646% to 2.441%;
  initial employment barely changes (49.108% versus 49.102%). The main
  interpretation is more genuine turnover, not a large revision to initial
  employment or the disappearance of misclassification.
- Among apparent entries, the share inferred to be genuine and reversed
  next wave rises from 6.48% to 13.99%; among apparent exits it rises from
  5.54% to 10.06%. Persistent-transition shares change much less. Nevertheless,
  55.80% of apparent entries and 45.88% of apparent exits are not confirmed
  in their reported direction under AR(2). The model therefore attributes
  more short-lived movement to genuine transitions while retaining an
  important role for misclassification, from either reporting or matching
  errors. These are shares of all apparent transitions, not a decomposition
  conditional on observing a reversal.
- After projecting out all 33 nuisance-score directions, the empirical
  two-effect information matrix has eigenvalues 1,226.75 and 1,634.83
  (condition number 1.33), unchanged across three projection tolerances.
  This supports local separation of the two return effects; it does not
  remove the earlier caveat about weak duration-reliability separation.
  No parameter reaches its numerical bound in the selected fit.

The full-sample direct-prior R and reweighted compiled likelihoods agree to
displayed machine precision; the maximum posterior difference is
$3.84\times10^{-15}$. Analytical return-coefficient scores agree with the
numerical scores within $4.3\times10^{-11}$. The weighted posterior mean status
misclassification is 2.44142%, consistent with the common fitted probability.
The focused suite passes 273 checks, including the audit table chunks.

## Remaining checks

- Check sensitivity to the initial-pair restriction before a structural
  interpretation. Keep broader reliability-mixture refinements secondary.
- Retain the separate caveat from Section 11: the timegap clock at a newly
  entered nonemployment spell is still drawn from the existing marginal law,
  rather than restricted to an onset within the last quarter. The AR(2)
  extension does not repair or validate that substantive modelling assumption.
- Interpret any surviving AR(2) effects as conditional associations, potentially
  reflecting residual duration misspecification, omitted heterogeneity, or
  reporting assumptions. No analytical or bootstrap standard errors are
  computed in this first extension.

Reproduce with `validate_four_wave_ar2.R`, `recover_four_wave_ar2.R`,
`estimate_four_wave_ar2.R`, and `summarise_four_wave_ar2.R`, in that order.
Outputs reside in the four-wave AR(2) results directory; repaired AR(1) and
historical pre-repair objects are preserved.

# 14. Subsequent extensions

The flexible baseline repairs the calendar distribution, exact retention and
local revisions capture part of the serial dependence, and separate
clock-specific reliability dispersions capture additional clustering of
duration irregularities. Following the decision to limit further work on
reliability-class separation, the remaining work should proceed in this order:

The repaired Panel A AR(1) refit is reported in Section 12. Its AR(2)
extension, recovery, nearby-start checks, local-identification diagnostic,
and first reversal comparison are completed in Section 13. The same-person
three-wave re-estimation remains omitted at the author's request.

1. check sensitivity to the initial-pair rule and the clean timegap
   distribution at a new nonemployment spell, changing one assumption at a time;
2. if a structural interpretation of the return effects is required, assess
   finer short-duration controls and remaining observed or unobserved
   heterogeneity rather than treating AR(2) coefficients as causal effects;
3. compute approximate analytical uncertainty once the preferred specification
   is settled, treating weak nuisance directions explicitly; and
4. retain uncertainty about reliability-class separation as a limitation.
   Additional class-share and bootstrap sensitivity work
   is secondary rather than a prerequisite for the employment extension.
